\section{Conclusion}\label{sec:concl}
We have surveyed the two document-formed user interfaces in the \pn project: Jupyter
Notebooks and Active Documents and developed a joint perspective on them that allows us to
propose a joint generalization oft the functionalities combines their respetive
advantages. We feel that embedding computations in arbitrary places in traditional
mathematical documents forms a natural generalization of the two existing user
interfaces. In particular the appearance as ``enhanced mathematical documents'' may make
the VRE more natural to mathematicians who have little experience with symbolic
computation tools, as they naturally have experiences with paper articles and textbooks
from their studies.

The survey and outlook provided by this report will be used as a basis for the discussion
of an integrated user interface and in the \pn project (WP4). As integration of knowledge
and computation (and interoperability between the various system involved) is a central
theme of WP6, this will require interaction between the work packages.

\subsection*{Acknowledgements} The authors are grateful to Min RK from Simula for
discussions on the inner workings of Jupyter and help with initial experiments with
setting up a a MMT kernel for Jupyter. Florian Rabe, Dan Alistarh, and Tom Wiesing have
worked on an initial integration of MMT with Jupyter that shaped and validated the vision
for an integrated user interface for the \pn VRE formulated in this report. Discussions
with Nicolas M. Thiéry have refined the ideas presented here.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:
