\addtocounter{wpno}{1}

\begin{Workpackage}{\thewpno}
\WPTitle{\wpname{\thewpno}}
\WPStart{Month 1}
\WPParticipant{PS}{48}
\WPParticipant{SA}{48}

\begin{WPObjectives}
  Objectives: The objectives of \theWP{} are to undertake all project
  management activities, including:
  \begin{itemize}
  \item monitoring the overall progress of the project and the use of
    resources;
  \item ensuring the timely production of deliverables and other
    project outputs;
  \item reporting to the European Commission on financial matters;
  \item preparing for and attending the annual project review
    meetings; and
  \item managing the project Advisory Board.
  \end{itemize}

  % The objective of  is to undertake all project management
  % activities, including setting up joint infrastructure, organizing
  % meetings, and producing overview reports.
\end{WPObjectives}

\begin{WPDescription}
  This workpackage will perform all the activities related to
  monitoring of progress towards the project milestones shown on
  Page~\ref{sect:milestones} and the deliverables listed on
  Page~\ref{sect:deliverables}, assuring the quality of the
  deliverables, ensuring the collation and distribution of the
  required reports, questionnaires and deliverables including the
  annual reports to the European Commission, arranging project
  management meetings, tracking the project budget in terms of
  expenditure and person-months, obtaining financial certificates as
  required, convening project management meetings, ensuring that
  important project documents such as the project contract and the
  consortium agreement are properly maintained and amended as
  necessary, ensuring that contractual details are complied with,
  monitoring compliance with the grant agreement, preparing for the
  annual review meetings, and reviewing research results against the
  aims and objectives of the project. It also involves managing and
  supporting the project Advisory Board, including supporting
  attendance at project meetings, convening Advisory Board meetings,
  and obtaining feedback on the project direction and results.

\end{WPDescription}

\begin{WPDeliverables}
\WPDeliverable{1}{mgt:tickets}{Create tickets for all relevant tasks / deliverables}

\TODO{Rewrite everything below using the WPDeliverable macro}
\begin{itemize}
\item
\ref{mgt:mailinglists}
(Month 1): 
Internal and external mailing lists.
\item 
\ref{mgt:swrepository}
(Month 1): 
Internal software repository.\TODO{Needed?}
\item
\ref{mgt:swrepository}
(Month 1): 
Internal software repository.\TODO{Needed?}
\item
\ref{mgt:periodic-rep-1}
(Month 12): 
Project Periodic Report (first year).
\item
\ref{mgt:periodic-rep-2}
 (Month 24): 
Project Periodic Report (second year).
\item
\ref{mgt:periodic-rep-3}
(Month 36): 
Project Periodic Report (third year).
\item
\ref{mgt:periodic-rep-4}
(Month 48): 
Project Periodic Report (fourth year).
\item
\ref{mgt:final-mgt-rep}
(Month 48): 
Project Final Report
\end{itemize}
\end{WPDeliverables}
\end{Workpackage}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: 
