\begin{sitedescription}{JU}
Jacobs University is a private Anglo-Saxon style research university.  It opened in 2001
and has an international student body (1320 students from 115 nations as of 2011).  The
KWARC (KnoWledge Adaptation and Reasoning for Content~\cite{KWARC:online}) Group headed by
{\emph{Prof.\ Dr.\ Michael Kohlhase}} specialises in knowledge management for STEM.
Formal logic, natural language semantics, and semantic web technology provide the
foundations for the research of the group.

\subsubsection*{Curriculum vitae}
% Curriculum of the personnel at this institution
\input{CVs/Michael.Kohlhase}
\input{CVs/Florian.Rabe}
\input{CVs/Christian.Maeder}
\input{CVs/Mihnea.Iancu}

\subsubsection*{Relevant previous experience:}

The KWARC group is the lead implementor of the OMDoc (Open Mathematical Document) format
for representing mathematical knowledge \cite{Kohlhase:OMDoc1.2} and redeveloped its
formal core in the OMDoc/MMT format~\cite{RabKoh:WSMSML13}. The latter has been
implemented in the MMT system~\cite{MMTSVN:on,RabKoh:WSMSML13} which provides efficient
implementations of the computational primitives such as type checking, flattening, and
presentation at a logic/foundation-independent level.  The group has developed services
powered by such semantically rich representations, different paths to obtaining them, as
well as platforms that integrate both aspects.  \emph{Services} include the adaptive
context-sensitive presentation framework provided by the MMT API and the semantic search
engine MathWebSearch\cite{KohSuc:asemf06,ProKoh:mwssofse12}. 

Semantic services can be integrated into the documents generated from OMDoc/MMT
representations, making them into ``active documents'', i.e. documents that are
interactive and adaptive to the user and situation.  For \emph{obtaining} rich content,
the group investigates assisted manual editing \cite{JucKoh:sidesc10:biblatex} as well as
automatic annotation using linguistic techniques \cite{GinJucAnc:alsaacl09}.  Finally,
KWARC has developed the \textsf{MathHub.info} portal a community-based library and
knowledge management system for flexiformal libraries, which can be used for semantic
publishing and eLearning~ \cite{KohDavGin:psewads11,MathHub:on,IanJucKoh:sdm14}.

The \textsf{OMDoc/MMT} knowledge representation format and the \textsf{MathHub.info}
system will an important basis for the developments Work Packages 4 and 6.

Michael Kohlhase has initiated and led the CALCULEMUS! IHP-Research and Training Network
and participated in the FP6 IST MoWGLI (Mathematics on the Web: Get it by Logic and
Interfaces) project, the FP6 CSA Once-CS (Open Network of Centres of Excellence in Complex
Systems), The FP7 EDC project WebALT (Web Advanced Learning Technologies).

\subsubsection*{Specific expertise:}
\begin{compactitem}
\item Modelling formal structures of mathematical knowledge in a web-scalable way.
\item Transforming large collections of legacy scientific publications to semantically
  structured markup.
\item Designing user interfaces for authoring and interacting with mathematical knowledge.
\end{compactitem}

\site{JU} lead \WPtref{dksbases} and tasks related to active and structured documents as
user interfaces (\WPtref{UI}) to the math VRE to be developed in \TheProject. It will run
an infrastructure for authoring and interacting with such documents and a search engine for
the \DKS base. 
\end{sitedescription}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: 

% LocalWords:  site-jacu.tex clange sitedescription emph compactitem pn semmath RabKoh
% LocalWords:  prosuming-flexiformal KohSuc asemf06 GinJucAnc alsaacl09 StaKoh ProKoh
% LocalWords:  tlcspx10 KohDavGin psewads11 ednote Radboud Bia ystok CALCULEMUS textsf
% LocalWords:  textbf keypubs OntoLangMathSemWeb uwb Deyan Ginev Stamerjohanns mwssofse12
% LocalWords:  searchability krasm Iacob KohKoh skmfe08 HilKohSta copmem06 ako IanJucKoh
%  LocalWords:  subsubsection sdm14 Centres vspace TOWRITE Modelling WPtref dksbases
