\begin{sitedescription}{LEEDS}

% Our person months should be 54 
% Dissemination: 24, HPC: 12, Social: 18

% PIC: 999976881
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there
The University of Leeds is a leading Research University in the
United Kingdom that was ranked 101st in the World in the most recent
2018 QS World University Rankings. In the 2014 UK Research Excellence Framework (REF),
Leeds was ranked in 10th place for research power, a measure of the number of people producing the highest quality research.
The assessment identified that over 80\% of research at Leeds has a top quality rating of either “world leading” or “internationally excellent”.

Dr Mike Croucher is a leading expert in Research Software Engineering (RSE) based in the Departments of Computer Science and IT Services.
He has an EPSRC RSE fellowship and is Head of Research Computing at The University of Leeds which includes High Performance Computing, Research Software Engineering and Data Analytics.
Mike is a Fellow of the Software Sustainability Institute and has over 13 years experience in Research Software.

The University of Leeds will focus on tasks on HPC (\taskref{hpc}{hpc-jupyter}),
dissemination (\taskref{dissem}{project-intro}).
\TheProject outputs. Most of the research tasks will be accomplished by the two researchers hired on the project, under the supervision of Mike Croucher who has replaced Neil lawrence as leadPI in the course of the project. The total amount ot PM goes up to 72.

\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

%\input{CVs/Neil.Lawrence.tex}
\input{CVs/Mike.Croucher.tex}
\begin{participant}[type=res,PM=1]{NN}
  We will support one month of an administrators time for assistance
  with workshop organisation and project administration across the
  whole duration of \TheProject.
\end{participant}


%\input{CVs/First.Last.tex}

\end{sitedescription}


\begin{draft}
\vspace{1cm}
\begin{verbatim}
- [X] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [X] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [X] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [X] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
  
- [X] Do all our tasks include us in the list of sites involved?
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS

%  LocalWords:  sitedescription SITraN Hallamshire taskref hpc hpc-jupyter dissem Fusi pp
%  LocalWords:  subsubsection organisation Lippert Stegle Hensman Luengo modelling Biol
%  LocalWords:  Computat Organisers IPython Bioconductor tigre gprege Jupyter vspace
