\section{Towards a Mathematical Virtual Research Environment}

From their earliest days, computers have been used in pure
mathematics, either to make tables, to prove theorems (famously the
four colour theorem) or, as with the astronomer's telescope, to
explore new theories. Computer aided experiments, and the use of
databases relying on computer calculations such as the Small Groups
Library in GAP, the Modular Atlas in group and representation theory,
or the LMFDB, are now part of the standard toolbox of the pure
mathematician, and certain areas of mathematics completely depend on
it. Computers are also increasingly used to support collaborative work
and education.

The last decades witnessed the emergence of a wide ecosystem of
open-source tools to support research in pure mathematics. This ranges
from specialized to general purpose computational tools such as \GAP,
\PariGP, \Linbox, \MPIR, \Sage, or \Singular, via online databases
like the \LMFDB. Not counting of course online services like the
Wikipedia, \Arxiv, or MathOverflow.

A great opportunity is the rapid emergence of key technologies, and in
particular the \Jupyter (previously \IPython) platform for interactive
and exploratory computing which targets all areas of science.

\subsection{Virtual Research Environments}

Promoting collaborations at all scales is key to further improve the
productivity of researchers in pure mathematics and applications. An
important next step in that direction is that of \emph{Virtual
Research Environments} (VRE), that is online services enabling
groups of researchers, typically widely dispersed, to work
collaboratively on a per project basis.
% , and supporting the entire
% life-cycle of computational work in mathematical research, from
% \emph{initial exploration} to \emph{publication}, \emph{teaching}, and
% \emph{outreach}.

\begin{oldpart}{MK: this should be shortened, also we must mention Mathematica, Maple,
    what else? What are their common parts?}
  Early VRE's include \SMC, a collaborative online environment where students, teachers
  and researchers can create, customize, and share a project. This project essentially
  consists of a virtual machine, with a simple web-based user interface, and ready-to-use
  software for interactive computations (e.g.\ \Sage) and authoring (e.g.\ \LaTeX), with
  facilities for real-time communication through chat, video, and shared editing of
  documents, programs and worksheets.  For education purposes, course material can be
  provided as worksheets, assignments can be distributed, collected, and returned as well.

  Technically speaking, \SMC is a specific open-source cloud-based Virtual Research and
  Teaching Environment for mathematics developed since 2013 under the lead of William
  Stein, with funding from the NSF, and Google's Education Grant program.  \ednote{NT:
    Mention SageMath, Inc.?}  \ednote{SL: 2016: SageMath, Inc.\ accepted in Google startup
    program, see https://twitter.com/wstein389/status/708439137200836608}

  It presently hosts over 250,000 projects and has over 12,000 weekly active users. This
  fast adoption by a wide variety of users demonstrates the relevance and the long-term
  impact this kind of collaborative environments can have.
\end{oldpart}

\subsection{Motivations}

The \ODK project grew out of a reflection on the needs of the (pure) mathematics community
in terms of computational software and databases. The highly successful development in the
last decades of open-source systems has proven the viability and power of collaborative
open-source development models, by users and for users, even for delivering general
purpose systems targeting a large public (researchers, teachers, engineers, amateurs,
\ldots).

Yet some critical long term investments, in particular on the technical side, are in order
to boost the productivity and lower the entry barrier:

\begin{itemize}
\item Streamline access, distribution, portability on a wide range of
  platforms, including High Performance Computers or cloud services.
\item Improve user interfaces, in particular in the promising area of
  collaborative workspaces as those provided by \SMC.
\item Lower barriers between research communities and promote
  dissemination. For example make it easy for a specialist of
  scientific computing to use tools from pure mathematics, and
  reciprocally.
\item Bring together the developers communities to promote tighter
  collaboration and symbiosis, accelerate joint development, and share
  best practices.
\item Outsource as much of the development as possible to larger
  communities to focus the work forces on their core specialty: the
  implementation of mathematical algorithms and databases.
\end{itemize}

Many people in the community have been working really hard on the
above items but crucially lack manpower or funding.

\subsection{\ODK}

The European H2020 call
\href{http://ec.europa.eu/research/participants/portal/desktop/en/opportunities/h2020/topics/2144-einfra-9-2015.html}{EINFRA-9:
  e-Infrastructure for Virtual Research Environment} was a natural fit
to seek for manpower and funding: putting the emphasis on Virtual
Research Environments nicely wraps up all the above needs in a single
mission.

A consortium was built by gathering core European developers of the
aforementioned systems for pure mathematics, and reaching toward the
numerical community, and in particular the \Jupyter community, to work
together on joint needs.
%
Together they answered the call with the \ODK proposal ``Open
Digital Research Environment Toolkit for the Advancement of
Mathematics''~\cite{OpenDreamKit:on} to help address the following aims in close
collaboration with the community:

\begin{itemize}
\item Further improve the productivity of researchers in pure mathematics and applications
  by further promoting collaborations on \emph{Data}, \emph{Knowledge}, and
  \emph{Software}.
\item Make it easy for teams of researchers of any size to set up custom, collaborative
  \emph{Virtual Research Environments} tailored to their specific needs, resources, and
  workflows.
\item Support the entire life-cycle of computational work in mathematical research, from
  \emph{initial exploration} to \emph{publication}, \emph{teaching}, and \emph{outreach}.
\end{itemize}

The acceptance of the proposal in May 2015 was a strong sign of
recognition, at the highest level of funding agencies, of the values
of open science and the strength and maturity of the ecosystem.

The \ODK projects~\cite{ODKproposal:on} will run for four years, starting from September
2015. It will provide substantial resources to the open source computational mathematics
ecosystem, and in particular popular tools such as LinBox, MPIR, SageMath, GAP, Pari/GP,
LMFDB, Singular, MathHub, and the IPython/Jupyter interactive computing environment.

The project involves about 50 people spread over 15 sites in Europe,
with a total budget of about 7.6 million euros. The largest portion of
that will be devoted to employing an average of 11 researchers and
developers working full time on the project. Additionally, the
participants will contribute the equivalent of six other people
working full time.

By definition this project will be mostly funding actions in Europe; however those actions
will be carried out, as usual, in close collaborations with the worldwide community
(potential users of the VRE as well as developers outside the \ODK consortium).

\subsection{Work plan}

\ednote{NT: this can be improved; e.g. Michael please have a go at
  WP6's item}

In practice, the \ODK work plan consists in 58 concrete tasks split in
seven work packages:
\begin{description}
\item[WP1] Project Management
\item[WP2: Community Building, Training, Dissemination, Exploitation,
  and Outreach]: organization of many development and training
  workshops, writing of tutorials and documentation, ...
\item[WP3: Component Architecture] work on portability -- especially
  on the Windows platform -- modularity, packaging, distribution,
  deployment, standardization and interoperability between components.
\item[WP4: User Interfaces] e.g. work on uniform \Jupyter notebook
  interface for all interactive computational components, improvements
  to\Jupyter, 3D visualization, documentation tools, ...
\item[WP5: High Performance Mathematical Computing] e.g. work within
  and between ODK's components to improve performance, and in
  particular better exploit multicore / parallel architectures.
\item[WP6: Data/Knowledge/Software-Bases] e.g identification and
  extensions of ontologies and standards to facilitate safe and
  efficient storage, reuse, interoperation and sharing of rich
  mathematical data whilst taking into account of provenance and
  citability; data archiving and sharing in a semantically sound way
  component architecture; Integration between computational software and databases;
\item[WP7: Social Aspects] research on social aspects of collaborative
  software/data/knowledge development in mathematics to inform the 
\end{description}
The completion of each task is validated by the release or publication
of deliverables.

% The tasks, deliverables, and progress thereupon are
%tracked publicly on the

\subsection{On open proposal writing and management}

An innovative aspect of the \ODK project is that its preparation and
management happens, as much as is practical and without infringing on
privacy, in the open. For example, most documents, including the
proposal itself, are version controlled on public repositories and
progress on tasks and deliverables is tracked using public issues (see
the \href{http://opendreamkit.org/}{OpenDreamKit web site}). This has
proven a strong feature to collaborate tightly with the community and
get early feedback.

%Computational experiments have led to new conjectures which have had a deep impact on the future development of
%mathematics. An outstanding example is the Birch and Swinnerton-Dyer conjecture (one of the Clay Millennium Problems).
%Databases relying on computer calculations such as the Small Groups Library in GAP, the Modular Atlas in group and rep-
%resentation theory, or the LMFDB, provide indispensable tools for researchers. A constructive way of understanding proofs
%%of deep theorems yields algorithmic tools to deal with highly abstract concepts. These tools make the concepts available
%to a broader class of researchers, with many potential applications. A prominent example from algebraic geometry is the
%desingularisation theorem of Hironaka, for which Hironaka won the Fields Medal, and its algorithmisation by Villamayor.


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
