\begin{participant}[type=leadPI,PM=13,gender=male]{Paul-Olivier Dehaye}
  is a Swiss National Science Foundation Assistant Professor at the
  University of Zurich. After his Phd at Stanford (2006), he has also worked in Oxford, at
  the Institut des Hautes Etudes Scientifiques and at ETH Zurich. He currently has 13
  papers published in international peer-reviewed journals. He is currently supervising
  three PhD students and one post-doc.

  His main research is at the intersection of Number Theory and Combinatorics, and in
  particular in Random Matrix Theory conjectures. He has additional interests in FLOSS,
  semantic tools, massive online education and crowdsourcing, all with the view of
  enabling larger scale mathematical and scientific collaborations. He is also member of
  the program committee of CICM 2015 (Conference on Intelligent Computer Mathematics).

  He is a contributor to the \Sage, \LMFDB and \OpenEdX projects, and has organised two
  conferences relating to these projects. The first was held in 2013 in Edinburgh, and
  organised jointly with Nicolas Thiery. Its official title was \emph{Online databases:
    from L-functions to combinatorics}, and it served as a precursor to some aspects of
  this grant, by bringing the \SageCombinat and \LMFDB communities together.  The second
  was held in June 2014 in Zurich and organised jointly with Stanford. It aimed at
  building a community around the open source python-based MOOC platform \OpenEdX, and
  opened a series of conferences now held twice annually.

  Dehaye has also taught for two years now a python course using \OpenEdX, which aims to
  bring first year students to the level of potential contributor to \Sage. This course
  also has a project-based component. It is now run locally for a small audience, but
  could be scaled up in various ways.
\end{participant}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:


