The design of OMDoc/MMT predates \pn, and we have reported on it previously~\cite{Kohlhase:OMDoc1.2,RabKoh:WSMSML13,DehKohKon:iop16,KohMuePfe:kbimss17}.
Therefore, we focus on the design of ULO here, which we report in Section~\ref{sec:ulo}.

Then, in Section~\ref{sec:isabelle}, we report on \taskref{dksbases}{isabelle}, which exports the Isabelle knowledge base into OMDoc/MMT and ULO.  The latter includes the generation of ULO data from concrete libraries in RDF format.  The resulting Isabelle Theorem dataset is massive, e.g., resulting in $>10^7$ RDF triples, requiring over $10$ CPU-hours to generate.
Contrary to the RDF formar, the OMDoc format does not abstract from the knowledge base, making it even larger. (In fact, we had to redesign MMT to allow for streaming compressed OMDoc files because it became infeasible to process them otherwise.)
In parallel work, we have developed a similar export of the Coq library ~\cite{MueRabSac:cltg19}, which is even bigger, but which we do not present in detail here.

Finally, in Section~\ref{sec:uloappl}, we demonstrate how to leverage these lightweight, high-level representations in practice.
As an example application, we set up a relational query engine based on Virtuoso.
It answers complex queries instantaneously, and even simple queries allow obtaining information that was previously impossible or expensive to extract.
Example queries include asking for all theorems of any library whose proof uses induction on $\mathbb{N}$,
%all theorems by a Michael Kohlhase or Makarius Wenzel, or 
%all theorems with incomplete Proofs,
or all authors of theorems ordered by how many of the proofs are incomplete,
%the average time it takes to check the proof assistant to check their theorems,
or all dependency paths through a particular library ordered by cumulative check time (which would enable optimized regression testing).
% \ednote{FR: make sure we mention motivating examples here that we can handle; then refer back to them when giving examples in Sect.~\ref{sec:appl}}

\subsection{The Upper Library Ontology}\label{sec:ulo}
\input{ulo}

\subsection{Exporting the Isabelle Knowledge Base}\label{sec:isabelle-export}
\input{isabelle}

\subsection{Applications}\label{sec:uloappl}
\input{ulo_appl}

%%% Local Variables:
%%% mode: latex
%%% mode: visual-line
%%% fill-column: 5000
%%% TeX-master: "report"
%%% End:

%  LocalWords:  ednote sec:uloappl mathbb Makarius optimized sec:appl ulo_appl RabKoh:WSMSML13,DehKohKon:iop16,KohMuePfe:kbimss17 taskref dksbases MueRabSac:cltg19
