\begin{participant}{Hans Fangohr}
  Hans Fangohr is Professor of Computational Modelling at the University of
  Southampton. He has studied Physics with specialisation in Computer Science and Applied
  Mathematics, gained his PhD in High Performance Computing (2002) and has since work in
  an Engineering department, carrying out interdisciplinary work based on computational
  science.

  He heads the University's interdisciplinary Computational Modelling Group
  (\url{http://cmg.soton.ac.uk}), and has more than 100 publications on applied computer
  simulation in magnetism, superconductivity, biochemistry, astrophysics and aircraft
  design, as well as development of computational methods.


  In 2013, he has attracted \EUR{5}m from the UK's Engineering and Physical Sciences
  Research Council (EPSRC) together with additional moneys from industry and his
  University to fund the \EUR{12}m Centre for Doctoral Training in Next Generation
  Computational Modelling (ngcm.soton.ac.uk) in the UK. This flagship activity will train
  about 75 PhD students (10 to 15 starting every year, beginning in September 2014) in the
  state-of-the-art and best-practice in computational modelling, the programming of
  existing and emerging parallel hardware and to apply these skills and tools to PhD
  research projects across a range of topics from Science and Engineering. The centre has
  chosen IPython as a key tool to deliver this teaching, document and communicate
  computational exploration and drive reproducible computation to push for excellent
  computational science.

  Hans Fangohr has led the development of the Open Source Nmag software
  (http://nmag.soton.ac.uk), which prodives a finite-element micromagnetic simulation
  suite to a community of material scientists, engineers and physicists who research
  magnetic nanostructures in academia and industry. He has designed the package in 2005 so
  that it has an IPython-compatible Python interface, to make the workflow of using the
  simulation package as accessible as possible to scientists without substantial
  computational background. He has extensive experience in micromagnetic simulation use
  and development.

  Hans has deep interested in excellence and innovation in learning and teaching. He has
  been awarded the prestigious Vice Chancellor’s teaching award ($\pounds 1000$) three
  times (in 2006, 2010, 2013) for initiating and realising three separate innovations in
  the university's teaching delivery, and has been voted ``best lecturer'' and ``funniest
  lecturer'' of the year by the students. Other Universities in the UK and elsewhere have
  adopted his teaching methods and materials. He has attracted grants to further develop
  learning and teaching activities, and given invited talks at international meetings on
  efficient learning and teaching of computational methods.
 
  Hans Fangohr is chairing the UK's national Scientific Advisory Committee for High
  Performance Computing.
\end{participant}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

