\paragraph{Ursula Martin}

% months=3
%
% Fair evaluation of the number of months you will be spending on this
% specific project along the four years

% salary=XXX
%
% Approximate monthly gross salary (in term of total cost for the
% employer). If you are uncomfortable having this information in a
% public file, you can alternatively send the information to Nicolas,
% or to your institution leader if the latter will be willing to fill
% in himself the budget forms on the eu portal.

% The above information will be used to evaluate the cost of the
% project for the institutions. You may remove the above comments once
% you have filled in the months= and salary= lines.

Dr. Michael Kohlhase is full professor for Computer Science at Jacobs University Bremen
and an associate adjunct professor at Carnegie Mellon University.

He studied pure mathematics at the Universities of T"ubingen and Bonn (1983 - 1989) and
continued with computer science, in particular higher-order unification and automated
theorem proving (Ph.D. 1994, Saarland University). 

His current research interests include knowledge representation for mathematics,
inference-based techniques for natural language processing, and computer-supported
education. He has pursued these interests during extended visits to Carnegie Mellon
University, SRI International, and the Universities of Amsterdam, Edinburgh, and Auckland.

Michael Kohlhase is recipient of the dissertation award of the Association of German
Artificial Intelligence Institutes (AKI; 1995) and of a Heisenberg stipend of the German
Research Council (DFG 2000-2003). He was a member of the Special Research Action 378
(Resource-Adaptive Cognitive Processes), leading projects on both automated theorem
proving and computational linguistics. Michael Kohlhase was trustee of the Conference on
Automated Deduction (CADE), Mathematical Knowledge Management (MKM), and the CALCULEMUS
conference, he is a member of the W3C Math working group, president of the OpenMath
Society, and the general secretary of the Conference on Intelligence Computer Mathematics
(CICM).

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../../proposal"
%%% End:


 


