\paragraph{Ursula Martin.}

% months=2.4
%
% Fair evaluation of the number of months you will be spending on this
% specific project along the four years

% salary=XXX
%
Professor Ursula Martin has  recently joined the University of Oxford, where
she holds a Professorship, in conjunction with a Senior EPSRC Fellowship, on a
joint arrangement between  the Department of Computer Science and the
Mathematical Institute. Her current  research concerns social and computational
techniques for creating mathematics, building on a significant track record at
the interface of mathematics and computing. Prior to this she worked at  Queen
Mary University of London, where as Vice Principal for Science and Engineering
she led strategic change, and was active in knowledge transfer
activities and developing young staff. 

Her work is characterized by strongly interdisciplinary collaboration in new
problem domains at the interface of mathematics and computer science,
identifying novel interactions between theory and practice, with real-world
problems inspiring scientific advance. Major achievements include results
linking randomness and symmetry, new unifying explanations of the power of
computational logic, and new practical techniques for using computational logic
and algebra in industry.

The work to be undertaken in the Work Package 5 (Social Aspects) fits very well
into her current project, which concerns crowdsourced mathematics: the overarching goal is
to understand and extend the human and computer creation of mathematics. 
It is mostly funded by her
2014 EPSRC Advanced Fellowship (EPSRC awards only one or two of these annually
in Computer Science) is a partnership of industry, government and international
academia.  
