% ---------------------------------------------------------------------------
%  Section 2: Impact
% ---------------------------------------------------------------------------

\section{Impact}
\label{sec:impact}

\TODO{Orsay's grant services will help here in December}

\subsection{Expected Impacts}

\eucommentary{Please be specific, and provide only information that applies
to the proposal and its objectives. Wherever possible, use quantified
indicators and targets.\\
Describe how your project will contribute to:\\
-- the expected impacts set out in the work programme, under the relevant topic
(including key performance indicators/metrics for monitoring results and impacts);\\
-- improving innovation capacity and the integration of new knowledge
(strengthening the competitiveness and growth of companies by developing
innovations meeting the needs of European and global markets; and, where
relevant, by delivering such innovations to the markets;\\
-- any other environmental and socially important impacts (if not already
covered above).\\
Describe any barriers/obstacles, and any framework conditions (such as
regulation and standards), that may determine whether and to what extent
the expected impacts will be achieved. (This should not include any risk
factors concerning implementation, as covered in section 3.2.)}

\draftpage

\subsection{Measures to Maximise Impact}

\subsubsection{Dissemination and Exploitation of Results}
\label{subsubsect:dissemination}

\eucommentary{-- Provide a draft 'plan for the dissemination and exploitation
of the project's results'. The plan, which should be proportionate to the
scale of the project, should contain measures to be implemented both during
and after the project.\\
Dissemination and exploitation measures should address the full range
of potential users and uses including research, commercial, investment,
social, environmental, policy making, setting standards, skills and
educational training.\\
The approach to innovation should be as comprehensive as possible,
and must be tailored to the specific technical, market and organisational
issues to be addressed\\
-- Explain how the proposed measures will help to achieve the expected impact of the
project . Provide a draft business plan for financial sustainability as stated in the Part
E of the Specific features for Research Infrastructures of the Horizon 2020 European
Research Infrastructures (including e-Infrastructures) Work Programme 2014-2015.\\
-- Where relevant, include information on how the participants will
manage the research data generated and/or collected during the
project, in particular addressing the following issues:
What types of data will the project generate/collect? What
standards will be used? How will this data be exploited and/or
shared/made accessible for verification and re-use (If data cannot
be made available, explain why)? How will this data be curated and preserved?\\ \\
-- Include information about any open source software used or developed by the
project.\\
You will need an appropriate consortium agreement to manage (amongst other things)
the ownership and access to key knowledge (IPR, data etc.). Where relevant,
these will allow you, collectively and individually, to pursue market opportunities
arising from the project's results.\\
The appropriate structure of the consortium to support exploitation is addressed
in section 3.3. \\ \\
-- Outline the strategy for knowledge management and protection. Include measures to
provide open access (free on-line access, such as the ``green'' or ``gold'' model) to
peer-reviewed scientific publications which might result from the project.\\
Open access publishing (also called 'gold' open access) means that an article is
immediately provided in open access mode by the scientific publisher. The associated costs
are usually shifted away from readers, and instead (for example) to the university or
research institute to which the researcher is affiliated, or to the funding agency supporting
the research.\\
Self-archiving (also called ``green'' open access) means that the published article or the
final peer-reviewed manuscript is archived by the researcher - or a representative - in an
online repository before, after or alongside its publication. Access to this article is often -
but not necessarily - delayed (``embargo period''), as some scientific publishers may wish to
recoup their investment by selling subscriptions and charging pay-per-download/view fees
during an exclusivity period.}


\paragraph{Long term sustainability}

The success of large specialized software like \Pari, \Singular or
\GAP in the last decades has shown the viability of the academic open
source development model for such. For a long time, it was bitterly
debated whether this model would have any chance to scale to general
purpose systems for pure mathematics. The rapid take off of Sage in
the last 10 years has proven the viability of the ``developed by users
for users'' model: despite its large community of about 150 active
developers, it's running on a tiny specific budget, with most
activities being funded indirectly by research grants that require
specific development.

This was made possibly by reusing existing components whenever
possible (e.g. hundreds of specialized open source math libraries, or
the \Python programming language with its developers tools and huge
library), and spinning off software development (e.g. the \Cython
compiler) to larger communities whenever possible.

\TODO{This piece of argument is tricky to setup!!!}

Yet, long term critical non mathematical features like portability,
modularization, packaging, user interfaces, large data, parallelism,
or outreach toward related software, have been lagging behind. Indeed
they can hardly be implemented as a side product of research projects,
and \textbf{need to be assigned to full time developers}. Regular
funding is also needed to better structure the computational
mathematics community in Europe and support its upcoming major
widening through training, development workshops, exchanges, ...

The purpose of this grant is to initiate this process. The principle
is that, with the growth of the user base, a tiny number of
institutions or companies will hire a full-time developer because they
critically need it to support their in-house research or development.
\TODO{Examples: LRI? Full time devs supported by research grant, like
  for Linbox? Others?}

The number of such required full time developers will be made even
tinier because most of the efforts now will be focused toward
outsourcing or spinning off more components to reduce the recurrent
needs.

For example, this project will save much recurrent efforts to the
mathematics community by outsourcing the development of the user
interface to IPython. This grant will provide the required temporary
boost to make IPython stand to the stringent needs of the community.
Later on, thanks to its large user base, both in academia and
industry, IPython will continue to thrive without specific funding or
major contributions from the mathematics community.

Another big focus of this project will be on the study of open source
development models for mathematical software and how they can be made
more productive, in particular by better processes and collaboration
between components, which will also reduce the number of required full
time developers.

\draftpage

\subsubsection{Communication activities}
\label{subsubsect:communication}

\eucommentary{Describe the proposed communication measures for promoting the
project and its findings during the period of the grant. Where appropriate
these measures should include social media and public events with user
participation. Measures should be proportionate to the scale of the project,
with clear objectives. They should be tailored to the needs of various audiences,
including groups beyond the project's own community. Where relevant, include
measures for public/societal engagement on issues related to the project.}
