\documentclass[a4paper,11pt]{article}

\newcommand{\XX}{\textbf{MathVRE}\xspace}
\newcommand{\TheProject}{\XX}

\input{preamble}
\input{participants}

\begin{document}

\begin{titlepage}

\begin{center}
{\Large \textbf{COVER PAGE}}
\end{center}

\begin{tabular}{llr}
\textbf{Title of Proposal:} & \textbf{\TheProject{}: 
  Open Virtual Research Environment Kit for Pure Mathematics and Applications} & \\[2ex] % \includeimage[scale=0.5]{logo} \\
\textbf{Date of preparation:} & \textbf{\today} & \comment{}{$
$Revision: 0.0$ $}\\[2ex]
\textbf{List of participants} && \\[2ex]


\end{tabular}

\begin{center}
\begin{tabular}{|l|p{3in}|l|l|}\hline
Participant no & Participant organisation name & Country\\

\hline
1 (Coordinator) & \longparticipant{1} & \country{1}  \\ \hline
2 & \longparticipant{2} & \country{2}  \\ \hline
3 & \longparticipant{3} & \country{3}  \\ \hline
4 & \longparticipant{4} & \country{4}  \\ \hline
5 & \longparticipant{5} & \country{5}  \\ \hline
6 & \longparticipant{6} & \country{6}  \\ \hline
7 & \longparticipant{7} & \country{7}  \\ \hline
8 & \longparticipant{8} & \country{8}  \\ \hline
9 & \longparticipant{9} & \country{9}  \\ \hline
10 & \longparticipant{10} & \country{10}  \\ \hline
11 & \longparticipant{11} & \country{11} \\ \hline
12 & \longparticipant{12} & \country{12} \\ \hline
13 & \longparticipant{13} & \country{13} \\ \hline
14 & \longparticipant{14} & \country{14} \\ \hline
15 & \longparticipant{15} & \country{15} \\ \hline
16 & \longparticipant{16} & \country{16} \\ \hline
\end{tabular}
\end{center}

\tableofcontents

\eucommentary{Please follow the structure of this template when
  preparing your proposal. It has been designed to ensure that the
  important aspects of your planned work are presented in a way that
  will enable the experts to make an effective assessment against the
  evaluation criteria. Sections 1, 2 and 3 each correspond to an
  evaluation criterion for a full proposal.\\
  Please be aware that proposals will be evaluated as they were
  submitted, rather than on their potential if certain changes were to
  be made. This means that only proposals that successfully address
  all the required aspects will have a chance of being funded. There
  will be no possibility for significant changes to content, budget
  and consortium composition during grant preparation.\\
  Page limit: The cover page, and sections 1, 2 and 3, together should
  not be longer than 70 pages. All tables in these sections must be
  included within this limit. The minimum font size allowed is 11
  points. The page size is A4, and all margins (top, bottom, left,
  right) should be at least 15 mm (not including any footers or
  headers).  If you attempt to upload a proposal longer than the
  specified limit, before the deadline you will receive an automatic
  warning, and will be advised to shorten and re-upload the
  proposal. After the deadline, any excess pages will be overprinted
  with a ‘watermark’, indicating to evaluators that these pages must
  be disregarded.\\
  Please do not consider the page limit as a target! It is in your
  interest to keep your text as concise as possible, since experts
  rarely view unnecessarily long proposals in a positive light.}
\end{titlepage}

\newpage

\include{outline}

% ---------------------------------------------------------------------------
%  Section 1: Excellence
% ---------------------------------------------------------------------------

\section{Excellence}

\TheProject focuses on delivering easy to setup Virtual Research
Environments, customizable to meet the diverse needs of collaborative
research in pure mathematics and applications, and built out of a
sustainable ecosystem of % both domain-specific and general purpose
community-developed open software, databases, and services.

The remarkable emergence and success in the last decade(s) of a large
ecosystem of community-developed computational software like
e.g.\ \GAP, \Linbox, \PariGP, \Sage, or \Singular and of interactive
scientific computing environments like \IPython showcase the viability
of community development models for such components.

The recent successes of the Virtual Research and Teaching Environment
\SMC (hosting more than 10k users and 100k projects after just
one year) and of the online number theory database LMFDB showcase the
strong need for \textbf{integrated solutions} enabling large-scale
collaboration on \textbf{software}, \textbf{knowledge}, and \textbf{data}.

Yet setting up infrastructures like those mentioned above currently requires
massive ad-hoc efforts. Challenges include portability, compatibility,
performance, usability, reproducibility, not to mention the many
social aspects involved in communities and ecosystems thereof.

\TODO{Better explain the specific needs in maths, and why we focus on
  VRE building blocks rather than VRE themselves, in the DIY spirit.}

A specific challenge in mathematics comes from the vast yet tightly
connected array of concepts involved. Different groups of researchers
may have radically different needs, workflows, and resources, calling
for a highly modular and customizable VRE infrastructure.

\TheProject will attack all those challenges upfront, while
consolidating Europe's leading position in this field.

% The development of effective, flexible and easy-to-use VRE technology
% is especially important for mathematicians, because their research
% groups tend to be small and can be isolated.

% \TODO{Reference this last claim? Maybe expand a little more,
% or justify later}

\subsection{Objectives}
\label{sect:objectives}

\eucommentary{1-2 pages}
\eucommentary{\emph{Describe the specific objectives for the project,
which should be clear, measurable, realistic and achievable within the
duration of the project. Objectives should be consistent with the expected
exploitation and impact of the project (see section 2).}}

The specific aims of \TheProject are:


\begin{enumerate}[\textbf{Aim} 1:]
\item \label{aim:collaboration} Improve the productivity of
  researchers in pure mathematics and applications by promoting
  collaborations based on mathematical \textbf{software},
  \textbf{data}, and \textbf{knowledge}.
\item \label{aim:vre} Make it easy for teams of
  researchers of any size to setup custom collaborative Virtual Research
  Environments tailored to their needs and workflows, supporting the
  entire life-cycle of computational work in mathematical research,
  from initial exploration to publication, teaching and outreach.
  % and bridge the gaps between
  % code, published results, and educational material.
\item \label{aim:sharing} Identify and promote best practices in
  computational mathematical research, including making results easily
  reproducible, making software sustainable, reusable and easily
  accessible and sharing data in a semantically sound way.
  \TODO{insert the \emph{ecosystem} keyword}
\item \label{aim:impact} Maximize sustainability as well as impact in
  mathematics, neighbor fields, and scientific computing.
\end{enumerate}



Our research will cover a wide variety of aspects, ranging from
software development models, user interfaces \TODO{virtual
  environments?}, deployment frameworks and novel collaborative tools,
component architecture, design, and standardization of software
\TODO{system?} and databases, to links to publication, data archival
and reproducibility of experiments, development models and tools, and
social aspects.


The concrete objectives of \TheProject are:
\begin{enumerate}[\textbf{Objective} 1:]
\item\label{objective:framework} To develop and standardise
  an architecture allowing a range of mathematical and
  data and software components to be combined 
  with off-the-shelf non-mathematical infrastructure to produce
  specialised VREs for different research communities. The
  architecture will take the form of standards documents and APIs
  equipped, where appropriate, with formal or informal mathematical
  semantics to ensure interactions are mathematically sound. This primarily
  addresses aim \ref{aim:vre}, thereby contributing to aims
  \ref{aim:collaboration} and~\ref{aim:sharing}. \TODO{This is a bit long}

\item\label{objectives:core} To develop open source core components
  for such VREs where existing software is not available. These
  components should support VREs running on a variety of
  platforms, including standard e-infrastructures. This primarily
  addresses Aim~\ref{aim:vre}, thereby contributing to Aim
  \ref{aim:collaboration} and~\ref{aim:sharing}.
  


\TODO{Urgent: talk to EPCC re e-infrastructure standards}

\TODO{SL Write detailed description}

\item \label{objective:community}Community building

  Bring together the communities (IPython, \Sage, \Singular)

\TODO{NT expand and write detailed description}
\item \label{objective:updates} Update a range of existing open source
  mathematical software systems for seamless deployment and efficient
  execution within the VRE architecture of objective~\ref{objective:framework}.
  This fulfills part of Aim~\ref{aim:vre}.

\TODO{SL: write detailed description}

  % Our tools will span the entire life-cycle of a research idea, . They will. This
  % project is based on existing, proven open source technologies
  % developed by our team over the last decade that have been widely
  % adopted in academia and industry.

\item \label{objective:sustainable} Foster a sustainable ecosystem of
  interoperable open source components developed by overlapping
  communities, in particular by identifying and promoting software
  development best practices, and outsourcing development to larger
  communities whenever suitable. This fulfills part of
  Aim~\ref{aim:sharing} and~\ref{aim:impact}.


\item \label{objective:social}Explore the social aspects: how do researchers collaborate in
  Mathematics? What can be the role of Virtual Research Environments?
  This addresses part of Aim~\ref{aim:sharing} and~\ref{aim:vre}.
\TODO{UM: write detailed description}
\TODO{SL: semantics from mathematical text?}

\item \label{objective:data}Identify and extend ontologies and standards to
  allow easy, safe and efficient storage, reuse, interoperation and sharing of rich mathematical
  data taking account of provenance and citability. This fulfills
  parts of Aim~\ref{aim:vre} and~\ref{aim:sharing}.
\TODO{POD -- write detailed description}

\item \label{objective:demo}Demonstrate the effectiveness of Virtual Research Environment
  built on top of \TheProject components for a number of real-world
  use cases taken from different domains, or crossing previously
  little connected domains. This addresses part of Aim~\ref{aim:vre}.

%Long term sustainability
\item \label{objective:disseminate}Effective Dissemination
\TODO{VP -- write detailed description}

\end{enumerate}

\subsection*{Detailed Descriptions of Objectives} % delete if a
                                % heading isn't needed here

\paragraph{Objective~\ref{objective:framework}: Virtual Research
  Environment Kit}\ 

Computational techniques have become a core asset for research in pure
  mathematics and its applications in the last decades. Mathematics
  communities have come together to develop powerful computational
  tools, such as GAP, \PariGP, SAGE or Singular, and valuable on-line
  services such as the Encyclopedia of Integer Sequences and the ATLAS
  of Group Representations. \TODO{cite} In building these systems, 
  mathematicians have gained strong
  experience in collaborative software development, with pioneering
  work and continuing leadership of Europe.

A number of approaches to linking these resources have been developed,
such as the SCSCP protocol from the Framework 6 SCIEnce project, and
the incorporation of a variety of free software tools in the SAGE
system, but the overall model is still that of a single mathematician
running programmes or interacting with a ``notebook''
page. The software provides little or no support for other aspects of
mathematical research: collaboration,
archival, reproducibility or linkage between programmes, data and
publication. Databases are updated mainly by mathematicians directly,
retaining no record of the source of new entries, and providing no way
of refering to the actual version of the data used in a particular
computation.

The first objective of this project is to design an architecture which
will allow existing mathematical software systems (suitably updated), 
off-the shelf non-mathematical tools and a small number of new
components to be flexibly combined to produce a VRE that will
effectively support collaborative mathematical research throughout
it's entire life-cycle. This will include software APIs and standards,
frameworks for assuring the semantic consistency of similar
mathematical objects in different systems.  It will be informed by the
outputs of objective \ref{objective:social}, ensuring that the VREs
fit the ways that mathematicians actually work.

\paragraph{Objective~\ref{objectives:core}:\TODO{title, CORE}}\
\TODO{SL Write detailed description}

Most of the direct mathematical capabilities of our software will come
from existing open source mathematical systems. For instance we will
use the power of the GAP Library for computational group theory or
\PariGP for number theory. Generic services such as storage, version
control, authentication and resource accounting will come from
off-the-shelf components building on standard infrastructures. 

Nevertheless some new tools will be needed \TODO{what? or at least examples}


\TODO{Keyword: flexible/versatile virtual environment}

Our research will cover a wide variety of aspects, ranging from
software development models, user interfaces \TODO{virtual
  environments?}  deployment frameworks and novel collaborative tools,
component architecture, design, and standardization of software
components and databases, to links to publication, data archival and
reproducibility of experiments, development models and tools, and
social aspects. It will build on the remarkable success of the open
source ecosystem and consolidate Europe's leading position in
computational mathematics.

Following the call specifications, all software, data, and
publications resulting from this proposal will be open.


\paragraph{Objective~\ref{objective:community}: \TODO{TITLE: COMMUNITY}}\ 
\TODO{NT detailed description}

\paragraph{Objective~\ref{objective:updates}: \TODO{TITLE: UPDATES}}\ 
\TODO{SL detailed description}


\paragraph{Objective~\ref{objective:sustainable}: A sustainable
  ecosystem of software components}\ 

The success of large specialized software like \PariGP, \Singular or
\GAP in the last decades has shown the viability of the academic open
source development model for such. For a long time, it was bitterly
debated whether this model would have any chance to scale to general
purpose systems for pure mathematics. The rapid takeoff of Sage in the
last decade has proven the viability of the ``developed by users for
users'' model: despite its large international community of about 150
active developers, it's running on a tiny specific budget, with most
activities being funded indirectly by research grants that require
specific development.

This was made possible by reusing existing components whenever
possible (e.g. hundreds of specialized open source math libraries, or
the \Python programming language with its developers tools and huge
library), by spinning off software development (e.g. the \Cython
compiler) to larger communities whenever possible, and by carefully
designing the development workflow.

Yet, long term critical non mathematical features like portability,
modularization, packaging, user interfaces, large data, parallelism,
or outreach toward related software, have been lagging behind. Indeed
they can hardly be implemented as a side product of research projects,
and \textbf{need to be assigned to a few full time developers}. Regular
funding is also needed to better structure the computational
mathematics community in Europe and support its upcoming major
widening through training, development workshops, exchanges, ...

\textbf{One purpose of this grant is to initiate this process and
  invest on the long run to reduce the recurrent needs.}

The principle is that, with the growth of the user base, a tiny number
of institutions or companies will hire a full-time developer to
support critical needs of their in-house research or development.
Opportunities for such hiring are for example actively investigated at
the Laboratoire de Recherche en Informatique. It should be noted that,
at the scale of a large university, the cost of licenses for analogous
commercial software can reach the same order of magnitude as that of a
developer.

%\TODO{Examples: LRI? Full time devs supported by research grant, like
%  for Linbox? Others?}

To reduce the number of required full time developers \TheProject will
invest toward, factoring out joint needs, and outsourcing or spinning
off more components to larger communities.
%
For example, \TheProject will save much recurrent efforts to the
mathematics community by providing a temporary boost to outsource the
development of the user interface of each computational component to
\IPython and make \IPython stand to the stringent needs of the
community.  Later on, thanks to its large user base both in academia
and industry, \IPython will continue to thrive without specific
funding or major contributions from the mathematics community.

\TheProject will also foster the productivity within the ecosystem by
investigating better collaboration processes between components, and
identifying, sharing, and promoting software development best
practices.

% open source development models for
\paragraph{Objective~\ref{objective:social}: \TODO{TITLE: SOCIAL}}\ 
\TODO{UM detailed description}

\paragraph{Objective~\ref{objective:data}: Next generation mathematical databases}\ 
\TODO{POD detailed description}

\paragraph{Objective~\ref{objective:demo}: \TODO{TITLE: DEMO}}\ 
\TODO{detailed description}


\paragraph{Objective~\ref{objective:disseminate}: \TODO{Title: disseminate}}\ 
\TODO{VP -- write detailed description}

The success of any research software or service is strongly related to its ability to attract and convince a great amount of users. Our different communities (Sage, Gap, \PariGP, Singular...) have developed a solid experience and network. As an example, Sage has gathered thousands of users in less than 10 years. This was achieved thanks to a very strong community building philosophy, especially through the organization of Sage-Days all over the world. The first Sage-days happened in 2006, today we count 63 of them, including 10 for 2014 only, and also Sage Education days, Sage Bug days, Sage Doc days and more. Most of us, \TheProject{} project members, have been involved in these events either as organizers or participants and it appears as the most efficient way to promote our software. More precisely, our objective is to create constant dialogue with the different communities: frequent workshops, conferences, user groups, mailing lists. By building on existing tools, we intend to involve the communities in the development process itself in the spirit of open-source software. 

We also intend to reach a larger crowd of researchers by cutting down non-research technical issues to access existing tools: building better documentation and tutorials, developing easy-to-install distributions, easy web and cloud access, better user interfaces, better interactions between different software. Doing so, our objective is to help the communities to grow themselves and interact together using our work. 


\clearpage

\draftpage

\subsection{Relation to the Work Programme}

\eucommentary{1-2 pages; Eugenia will help there}

\eucommentary{
Indicate the work programme topic to which your proposal relates, and
explain how your proposal addresses the specific challenge and scope
of that topic, as set out in the work programme.}

\eucommentary{
  \verbatiminput{call_description}
}

\draftpage
% ---------------------------------------------------------------------------
%  Section 1.3: Concept and Approach
% ---------------------------------------------------------------------------
\input{concept.tex}

\draftpage
% ---------------------------------------------------------------------------
%  Section 1.4: Ambition
% ---------------------------------------------------------------------------
\input{ambition.tex}

\draftpage
% ---------------------------------------------------------------------------
%  Section 2: Impact
% ---------------------------------------------------------------------------
\input{impact.tex}

\clearpage

% ---------------------------------------------------------------------------
%  Section 3: Implementation
% ---------------------------------------------------------------------------

\section{Implementation}

\TODO{Typical granularity: 5-8 work packages with 3-5 tasks and one
  deliverable per task; 10 milestones}

\subsection{Work Plan --- Work packages, deliverables and milestones}
\label{sect:workplan}

\eucommentary{Please provide the following:\\
\begin{itemize}
\item
brief presentation of the overall structure of the work plan;
\item
timing of the different work packages and their components (Gantt chart or similar);
\item
detailed work description, i.e.:
\begin{itemize}
\item
a description of each work package (table 3.1a);
\item
a list of work packages (table 3.1b);
\item
a list of major deliverables (table 3.1c);
\end{itemize}
\item
graphical presentation of the components showing how they inter-relate (Pert chart or similar).
\end{itemize}
}

\subsubsection*{Overall Structure of the Work Plan}

The work plan is broken down into XX workpackages as shown
in Figure~\ref{}: WP2 deals with  ...
In addition, there is one management work package (WP1) and one
general dissemination work package (\ref{m}). The Gantt chart on
Page~\pageref{fig:gantt} illustrates the timeline for the
various tasks for these work packages, including inter-task
dependencies.

%\newpage
\subsubsection*{How the Work Packages will Achieve the Project Objectives}
\label{sssec:how_the_work_packages_will_achieve}

\TOWRITE{ALL}{This needs to explain that we're actually going to meet the
objectives.  Needs to be done after objectives and WPs.}

The project objectives (Section~\ref{sect:objectives},
page~\pageref{sect:objectives}) and the corresponding work
packages that contribute to achieving those objectives are:

\begin{center}
\begin{tabular}{|l|l|l|}\hline
\textbf{Objective} & \textbf{Purpose} & \textbf{WPs} \\\hline \hline
Objective 1 & XX & \textbf{WPX} \\\hline
\end{tabular}
\end{center}

\paragraph*{Work Programme for Objective 1: }

Objective 1 is covered by WPX, which will ...

\landscape

\subsubsection*{Work Plan Timing: GANTT Chart showing Task Dependencies and Information Flows}


\vspace{-0.7in} \centerline{\hbox to \columnwidth{\hss%\includeimage[scale=0.85,angle=270]{ParaPhrase-Gantt2.pdf}
\hss}}
\label{fig:gantt}
\vspace{-1in} % Fool LaTeX into avoiding unnecessary page break
\endlandscape

\newpage

%\input{deliverables-dates}
\input{deliverables}

\addtocounter{subsubsection}{1}
\addcontentsline{toc}{subsubsection}{\protect\numberline{\thesubsubsection}Work
Package List}
\fbox{\begin{minipage}{\textwidth}\begin{center}{\Large\bf
        Work package list} % (full duration of project)}
  \end{center}
  \end{minipage}}

\bigskip\bigskip

\begin{tabular}{|p{1.2cm}|p{9.15cm}|p{0.8cm}|p{1.2cm}|p{1cm}|p{0.9cm}|p{0.9cm}|}
\hline
{\bf Work \mbox{package} No} & {\bf Work package title} &
{\bf Lead \mbox{partic.} no.} &
{\bf Lead short name} &
{\bf Person months} & {\bf Start month} & {\bf End month} \\\hline

\newcounter{wp}

% 1 Management
\addtocounter{wp}{1}
\workpackageentry{\thewp}{PS}{}{1}{60}

% 2 Community building and engagement
\addtocounter{wp}{1}
\workpackageentry{\thewp}{PS}{}{}{}

% 3 Component architecture
\addtocounter{wp}{1}
\workpackageentry{\thewp}{UV}{}{}{}

% 4 User interfaces
\addtocounter{wp}{1}
\workpackageentry{\thewp}{SR}{}{}{}

% 5 HPC and massively parallel computation
\addtocounter{wp}{1}
\workpackageentry{\thewp}{UJF}{}{}{}

% 6 Next generation databases
\addtocounter{wp}{1}
\workpackageentry{\thewp}{JU}{}{}{}

% 7 Social aspects
\addtocounter{wp}{1}
\workpackageentry{\thewp}{UO}{}{}{}

% 8 Dissemination
\addtocounter{wp}{1}
\workpackageentry{\thewp}{SA}{}{}{}

{\textbf{Total}} & & & &
\textbf{\large XXX}&
&
\\\hline
\end{tabular}


% \textbf{Summary:}\\[1ex]

\newpage

\fbox{\begin{minipage}{\textwidth}\begin{center}\Large\bf List of Deliverables
  \end{center}
  \end{minipage}}

\label{sect:deliverables}

\bigskip\bigskip\bigskip

\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{|p{0.8cm}|p{8.75cm}|p{0.8cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|p{1.2cm}|}  \hline
\textbf{Del. no.}              & \textbf{Deliverable name}        & \textbf{WP no.} & \textbf{Lead}
& \textbf{Type}              & \textbf{Dissemi- nation level}   & \textbf{Delivery date}
\\ \hline

%% Year 1

%\ref{del:xx}  & Requirements Analysis
%& WP? & & R & CO &  ?? \\
\hline
\end{tabular}
\end{center}
\end{minipage}


\newpage

%% Set up the milestone numbers.
\input{milestones}

\fbox{\begin{minipage}{\textwidth}\begin{center}\Large\bf List of milestones
  \end{center}
  \end{minipage}}
\label{sect:milestones}

\bigskip\bigskip\bigskip

\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular*}{\textwidth}{|p{1.5cm}|p{6.7cm}|p{2.5cm}|p{1.5cm}|p{3.6cm}|}  \hline
\textbf{Milestone number} & \textbf{Milestone name} & \textbf{Related work
  package(s)} & \textbf{Estimated date} & \textbf{Means of
  verification} (deliverables shown here + success criteria below) \\
\hline
\ref{mil:initial} &
  Completed initial requirements analysis.  &
  WPX &
  1 &
\ref{del:requirements-analysis}.
\\
\ref{mil:final} &
&
WPX &
&
\\
\hline
\end{tabular*}
\end{center}
\end{minipage}

\vspace{10pt}
\begin{center}
\begin{tabular*}{\textwidth}{|p{1.5cm}|p{13.3cm}|p{1.9cm}|}\hline
\textbf{Milestone} & \textbf{Success Criteria} & \textbf{Contributes to
  Objective(s)} \\\hline
\ref{mil:initial} &
Completed requirements analysis (Deliverable~\ref{del:requirements-analysis}). &
 \textbf{1, 3.}
\\
\ref{mil:final} &
XX
& \textbf{XX}
\\\hline
\end{tabular*}
\end{center}

\eucommentary{
KEY
Estimated date
Measured in months from the project start date (month 1)
Means of verification
Show how you will confirm that the milestone has been attained. Refer to indicators if appropriate.
For example: a laboratory prototype that is ‘up and running’; software released and validated by a
user group; field survey complete and data quality validated.
}

% ---------------------------------------------------------------------------
% Include Workpackage descriptions
% ---------------------------------------------------------------------------

\input{WorkPackages/WorkPackages}

\TODO{Milestones need to be discussed and then described here.}

\newpage

\TODO{Check this for any necessary changes.}


\subsection{Management Structure and Procedures}
\label{sect:mgt}

\eucommentary{Will get help from Orsay's grant services}

\eucommentary{Describe the organisational structure and the decision-making
(including a list of milestones (table 3.2a)).\\
Explain why the organisational structure and decision-making mechanisms are
 appropriate to the complexity and scale of the project.\\
Describe, where relevant, how effective innovation management will be
addressed in the management structure and work plan.\\
Describe any critical risks, relating to project implementation, that
the stated project's objectives may not be achieved. Detail any risk
mitigation measures. Please provide a table with critical risks
identified and mitigating actions (table 3.2b).}

\draftpage
\subsection{Consortium as a Whole}
\eucommentary{\begin{itemize}
\item
Describe the consortium. How will it match the project's objectives?
How do the members complement one another (and cover the value chain,
where appropriate)? In what way does each of them contribute to the
project? How will they be able to work effectively together?
\item
If applicable, describe the industrial/commercial involvement in the
project to ensure exploitation of the results and explain why this is
consistent with and will help to achieve the specific measures which
are proposed for exploitation of the results of the project (see section 2.3).
\item
Other countries: If one or more of the participants requesting EU funding
is based in a country that is not automatically eligible for such funding
(entities from Member States of the EU, from Associated Countries and
from one of the countries in the exhaustive list included in General
Annex A of the work programme are automatically eligible for EU funding),
 explain why the participation of the entity in question is essential to carrying out the project
\end{itemize}
}

\TODO{The participants are core developers of the involved components}

\TODO{Experience in community building and engagement}



Writing interfaces between computer algebra systems from different areas and collaboritive 
software development are important themes within the DFG Priority Project SPP1489.
As in the {\sc{Sage}} community, networking measures include the regular exchange 
of developers and the regular organization of software workshops (coding sprints) which 
bring whole teams together for solution finding and intense code writing. Particular tight 
collaborations exist between the {\sc{GAP}} and the {\sc{Singular}} communities, with 
major {\sc{GAP-Singular}} developers meetings taking alternately place at St. Andrews,
Kaiserslautern, and Aachen. See \url{http://www.computeralgebra.de/}.

\TODO{User interfaces: recruitement of IPython developers}

\TODO{Pythran and HPC: a key asset will be the recruitment of two of
  the lead developers of the \Pythran \Python-to-C compiler.}

\TOWRITE{JGD/CP}{Linbox: recruitement of previous ANR developper}

\TODO{Explanation of why we want to include Seattle (sage-math cloud,
  is a key component; access to IP).}

\draftpage

\subsection{Resources to be Committed}

\eucommentary{Will get help from Orsay's grant services}

\eucommentary{Please provide the following:
\begin{itemize}
\item
a table showing number of person/months required (table 3.4a)
\item
a table showing 'other direct costs' (table 3.4b) for participants where
those costs exceed 15\% of the personnel costs (according to the budget
table in section 3 of the administrative proposal forms)
\end{itemize}}

\input{resources}

\subsubsection*{Management Level Description of Resources and Budget}

\TODO{This needs to be updated in line with the rest of the
project.}

The project will employ XX person-months of effort over YY years,
comprising ...


% ---------------------------------------------------------------------------
%  Section 4: Members of the Consortium
% ---------------------------------------------------------------------------

\newpage

\eucommentary{This section is not covered by the page limit.\\
The information provided here will be used to judge the operational capacity.}

\section{Members of the Consortium}

\subsection{Participants}

\eucommentary{Please provide, for each participant, the following (if available):\\
\begin{itemize}
\item
a description of the legal entity and its main tasks,
with an explanation of how its profile matches the tasks in the proposal;
\item
a curriculum vitae or description of the profile of the persons,
including their gender, who will be primarily responsible for carrying
out the proposed research and/or innovation activities;
%
this includes a description of the profile of the to-be-recruited personnel
\item
a list of up to 5 relevant publications, and/or products, services
(including widely-used datasets or software), or other achievements
relevant to the call content;
\item
a list of up to 5 relevant previous projects or activities, connected
to the subject of this proposal;
\item
a description of any significant infrastructure and/or any major items
of technical equipment, relevant to the proposed work;
\item
any other supporting documents specified in the work programme for this call.
\end{itemize}}

\clearpage
\input{Participants/StAndrews.tex}
\clearpage
\input{Participants/ParisSud.tex}
\clearpage
\input{Participants/Kaiserslautern.tex}
\clearpage
\input{Participants/Oxford.tex}
\clearpage
\input{Participants/Bordeaux.tex}
\clearpage
\input{Participants/Grenoble.tex}
\clearpage
\input{Participants/Logilab.tex}
\clearpage
\input{Participants/Warwick.tex}
\clearpage
\input{Participants/Bremen.tex}
\clearpage
\input{Participants/Zurich.tex}
\clearpage
\input{Participants/Simula.tex}
\clearpage
\input{Participants/Southampton.tex}
\clearpage
\input{Participants/Versailles.tex}
\clearpage

\TOWRITE{MK}{Description of University of Silesia}
\TOWRITE{MK}{PIC for University of Silesia}
\TOWRITE{WS}{Description of University of Washington, CV}

\TODO{Reorder partners accordingly to the list on the front page?}

\cite{science-project}

\subsection{Third Parties Involved in the Project (including use of third party resources)}

\eucommentary{Please complete, for each participant, the table
(see page 27 of "VRETemplate.PDF"),
or simply state "No third parties involved", if applicable.}

No third parties involved.

\TODO{Or Seattle?}

% ---------------------------------------------------------------------------
%  Section 5: Ethics and Security
% ---------------------------------------------------------------------------

\newpage

\section{Ethics and Security}

\eucommentary{This section is not covered by the page limit.}

\subsection{Ethics}

\eucommentary{
If you have entered any ethics issues in the ethical issue table in the administrative proposal forms, you must:\\
$\bullet$ submit an ethics self-assessment, which: \\
-- describes how the proposal meets the national legal and ethical requirements of the
country or countries where the tasks raising ethical issues are to be carried out;\\
-- explains in detail how you intend to address the issues in the ethical issues table, in
particular as regards:
research objectives (e.g. study of vulnerable populations, dual use, etc.),
research methodology (e.g. clinical trials, involvement of children and related
consent procedures, protection of any data collected, etc.),
the potential impact of the research (e.g. dual use issues, environmental damage,
stigmatisation of particular social groups, political or financial retaliation,
benefit-sharing, malevolent use , etc.)\\
$\bullet$ provide the documents that you need under national law(if you already have them), e.g.:\\
-- an ethics committee opinion;\\
-- the document notifying activities raising ethical issues or authorising such activities\\
If these documents are not in English, you must also submit an English summary of them
(containing, if available, the conclusions of the committee or authority concerned).\\
If you plan to request these documents specifically for the project
you are proposing, your request must contain an explicit reference to the project title}

\subsection{Security}

Please indicate if your proposal will involve:

\begin{itemize}
\item
activities or results raising security issues: NO
\item
'EU-classified information' as background or results: NO
\end{itemize}

\newpage

\label{bibliography}
\addcontentsline{toc}{section}{References}

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
