\TOWRITE{NT}{Rework this section}

\subsubsection{Travel, dissemination and outreach}
\label{sect:budget-details-travel}

This community building nature of this grant proposal requires a large
number of staff exchanges, work shops with project partners and
workshops engaging the wider community in addition to the usual
management and project review meetings. For dissemination, we need to
target the computer science and focussed community and their conferences
as well as the domains benefitting from \TheProject, such as 
Mathematics and Science.

\subsubsection{Travel, dissemination}

We use the following guidelines for expected travel expenses:
\euro{2200} for attendance of a typical one week international
conference outside Europe (including travel, subsistence,
accommodation and registration), \euro{1200} for a corresponding
conference in Europe, \euro{750} for a one-week visit to project
partners, for example for coding sprints and one-to-one visits for
particular research. We expect a similar cost per week when hosting
visitors. For the 6-monthly project meetings, we expect on average a
cost of \euro{400} for travel, accommodation and subsistence.

For a partner site with one investigator and one full time researcher,
we expect that that both will attend all of the 9 project meetings that take 
place every 6 months (cost of 9 * 2 * 400 =
\euro{7200}), and that the site spends \euro{2000} per year to host
external visitors contributing to the project (total \euro{8000}). We
expect the investigator and research together to do 4 one-week visits
to other sites (each at \euro{750}, totals \euro{12000}). 


For dissemination, we expect the researcher to attend on average 1
international conference and 2 European meetings per year (totals
\euro{18400}) and the investigator to attend one international and one
european gathering (totals \euro{13600}).

Where there are multiple investigators per site, they will share the
travel and associated costs outlined above. Where there are multiple
researchers, or researchers not employed for the full 48 months, the
travel budget is reduced accordingly.

\subsubsection{Outreach costs}

We request funds for outreach activities such as workshops that
facilitate community building, disseminate best practice and
encourages sustained contributions of the community to the project and
beyond the lifetime of the funding. Details are given in the tables
below and in the work packages. 

We also request \euro{1000} per year per partner to pay for open
access publication charges.


\TODO{Give details in work packages on what workshops are planned} 

\TODO{List workshop expenses and travel expenses in tables below}.


\eucommentary{Will get help from Orsay's grant services}

\eucommentary{Please provide the following:
\begin{itemize}
\item
a table showing number of person/months required (table 3.4a)
\item
a table showing 'other direct costs' (table 3.4b) for participants where
those costs exceed 15\% of the personnel costs (according to the budget
table in section 3 of the administrative proposal forms)
\end{itemize}}

%\newpage

%\landscape

\fbox{\begin{minipage}{\textwidth}

\begin{center}\Large\bf
Summary of staff effort
\end{center}
\end{minipage}}

\eucommentary{Please indicate the number of person/months over the whole 
duration of the planned work, for each work package, for each participant. 
Identify the work-package leader for each WP by showing the relevant 
person-month figure in bold.}

\bigskip

%--------------

\subsection{Resources Southampton}

Southampton requests 32 person months for a researcher (expected to start in month 4 of the project), 6 person months for the lead PI (Hans Fangohr) and 2 person months for the co investigator (Ian Hawke). The lead PI will take on all management responsibilities. The researcher will not be employed for the whole project duration, and the PIs will carry out all tasks for the project in the remaining period. 

\bigskip
\begin{table}[h!]
\begin{tabular}{|r|r|p{9cm}|}
\hline
\textbf{} & \textbf{Cost (\euro)} & \textbf{Justification} \\\hline
\textbf{Travel} & 51,500& Travel as outlined in \ref{sect:budget-details-travel}\\\hline
\textbf{Equipment} & 4,000 & 2 High Performance Laptops\\\hline
\textbf{Equipment} & 6,000 & HPC Workstation to host \OOMMFNB webserver, \taskref{UI}{oommf-nb-ve}\\\hline

\textbf{Other goods and services} & 19,600 & 4 Dissemination workshops (travel \& room hire), \taskref{dissem}{dissemination-of-oommf-nb-workshops}\\\hline
\textbf{Audit} & 5,200 & Audit cost (management)\\\hline
\textbf{Total} & 86,300\\\cline{1-2}
\end{tabular}
\caption{Overview: Non-staff resources to be committed at Southampton (all in \texteuro)}\label{tab:resources-non-staff-southampton}\vspace*{-1em}
\end{table}


%--------------


\subsection{Resources partner X}
\subsection{Resources partner Y}
\subsection{\ldots}










\subsection{Resources partner Summary}

\begin{table}[ht]\centering
  \TODO{Table 3.4.a: insert here table from Figure 3, and transpose; see
    Table 3.4.a in the word template}
  \TODO{The work package leader will usually have the largest investment}
\caption{Overview: Resources to be committed (all in \texteuro)}\label{tab:resources}\vspace*{-1em}
\end{table}
\fbox{\begin{minipage}{\textwidth}

\eucommentary{Please complete the table below for each participant if the sum of the costs for’ travel’, ‘equipment’,
and ‘goods and services’ exceeds 15% of the personnel costs for that participant (according to the
budget table in section 3 of the proposal administrative forms).}

\begin{center}\Large\bf
Other direct cost items
\end{center}
\end{minipage}}

\bigskip

\begin{tabular}{|r|l|p{9cm}|}
\hline
\textbf{} & \textbf{Cost (\euro)} & \textbf{Justification} \\\hline
\textbf{Travel} & & \\\hline
\textbf{Equipment} & & \\\hline
\textbf{Other goods and services} & & \\\hline
\textbf{Total} & \\\cline{1-2}
\end{tabular}




%%% Local Variables:
%%% mode: latex
%%% TeX-master: "proposal"
%%% End:

%  LocalWords:  newpage fbox minipage textwidth eucommentary bigskip rr rr rr hline
%  LocalWords:  vspace texteuro textbf textbf textbf
