\documentclass{deliverablereport}

\usepackage{xparse}
\usepackage{etoolbox}
\usepackage{caption}
 \ExplSyntaxOn

\newcounter{eventcounter}

\newenvironment{event}[6]{
\vspace{0.5cm}
\refstepcounter{eventcounter}
\label{event-#2}


\noindent\textbf{Event~\theeventcounter -~ #1}\newline % title

\noindent #3 \newline % location and date

\noindent ODK~partners~involved:~ \clist_map_inline:nn{#4}{\site{##1}~}\newline %partners

\ifx&#5&%
      % no participant #
\else
\noindent #5~participants\newline
\fi

\ifx&#6&%
      % no website
\else
\noindent \url{#6}\newline
\fi



}{\begin{center}\noindent\rule{4cm}{0.4pt}\end{center}}

 \ExplSyntaxOff


\deliverable{dissem}{workshops-1}
\duedate{31/08/2016 (M12)}
\deliverydate{31/08/2016}
\author{Viviane Pons}

\begin{document}
\maketitle
\githubissuedescription
\newpage
\tableofcontents
\newpage

% Other tasks and deliv impacted: T3.6, T4.1, T6.1, T6.3, T6.10, D6.2 WP5 D4.4 D4.7 T3.1 T4.4 T4.6

\section{Project meetings and development workshops}

We call a development workshop an event with a restricted number of participants
who meet to work on a specific task. These workshops are an inherent part
of \ODK development process as described in \taskref{dissem}{devel-workshops}:
 they bring together
developers from within and outside of \ODK and allow effective work
and discussions on many technical aspects. They also participate in building
and maintaining a community of developers inside \ODK and within the
open-source communities we belong to.

We list here all workshops which have been organized or co-organized by \ODK
as well as external workshops which have been attended by \ODK participants
with a significant impact on the project. We also include project meetings as they
participate to the same goal of bringing together participants of the project and
always include some development time.

Throughout the first year of the project, we have had
\begin{itemize}
\item 3 project meetings,
\item 2 workshops fully funded and organized by \ODK,
\item 2 workshops partially funded and co-organized by \ODK,
\item 1 external workshop.
\end{itemize}
This includes in particular one development edition of Sage Days and one Atelier
Pari that we had announced on \longtaskref{dissem}{devel-workshops}.
 These events had a major impact on other \ODK workpackages, tasks, and deliverables: WP3:
\taskref{component-architecture}{portability},
\taskref{component-architecture}{extract-smc}
WP4:
\taskref{UI}{ipython-kernels},
\taskref{UI}{sage-sphinx},
\taskref{UI}{structdocs},
\delivref{UI}{ipython-kernels-basic},
\delivref{UI}{ipython-kernels},
WP6:
\taskref{dksbases}{data-design},
\delivref{dksbases}{design}


\input{events/2015-09-02-Kickoff}
\newpage
\input{events/2015-11-SD70}

\input{events/2016-01-Atelier-PARI-GP}

\input{events/2016-01-GAP-Sage-days}

%\input{events/2016-01-25-DKS}

\input{events/2016-04-04-SD77}

\input{events/2016-05-wp6-workshop}

\input{events/2016-06-27-Bremen}

\clearpage

\section{Dissemination and outreaching activities}

We describe here all activities related to \taskref{dissem}{dissemination}:
these are all events oriented towards dissemination, training, and outreach. This
includes events organized or co-organized by \ODK and also
participating in external events and many communication activities.

\subsection{Organization of Sage Days in established mathematical communities}

One goal of \ODK is to support local communities of researchers
and developers who contribute to the open-source softwares related to
the project. For Sage, this means supporting the organization of Sage-Days
workshops that arise from within all the different mathematical communities. The main 
goal of these workshops is mostly to improve the Sage coverage of some mathematical
area. They also play a major role in training and communication. The
impact for \ODK can be summarized this way:

\begin{itemize}
\item \textbf{Making ODK known to the end users}: by supporting Sage Days,
\ODK makes itself known to the Sage community and can
thus share the many developments of the project.

\item \textbf{Improving the overall quality of Sage}: by fostering researchers
in specific areas, Sage Days help bring interesting mathematics into
the software, which is beneficial for Sage and so \ODK.

\item \textbf{Training, bringing more user}: Sage Days are the perfect place
for new comers, especially students, to get their first experience with the software.

\item \textbf{Fostering a community}: Sage Days are helping making Sage a vibrant
community, which is vital for the success of \ODK.
\end{itemize}

Throughout the first year of the project, \ODK has been organizing or co-organizing
four Sage Days: one about geometry and dynamics of surfaces, one in differential geometry and topology,
 one in combinatorics and one in coding theory.

\input{events/2016-05-SD73}

\input{events/2016-05-SD74}

\input{events/2016-06-SD78}

\input{events/2016-08-SD75}
\newpage
\subsection{Training activities for Sage in developing countries}

As open-source software developers, we wish our products
to be accessible to as many people as possible. Even though we offer
 a free access, there is still a technical gap in many 
developing countries that 
often prevents schools and researchers to benefit from our softwares.
This is why we believe the role of \ODK is to foster 
a wider community that does not leave a part of the world behind. In 
this section, we describe training activities that have been conducted 
through \ODK in this regard.



The events \ref{event-2015DIMACOS}, \ref{event-2016HADAT}, and
\ref{event-2016CATN} described in this section are part of a long term
plan from Adrien Boussicault (Bordeaux University) to create an active
and autonomous Sage community in the Mediterranean area by repeated
events and the creation of a local team. We also include event
\ref{event-2016ECCO}, where Viviane Pons (Université Paris-Sud)
represented \ODK (ODK), which took place in Colombia and was
\href{http://blogs.ams.org/matheducation/2016/08/22/an-inclusive-maths-conference-ecco-2016/}
{a positive example of the inclusive and open spirit} we want to
achieve.

\input{events/2015-11-DIMACOS}

\input{events/2016-03-HADAT}

\input{events/2016-03-CATN}

\input{events/2016-06-ECCO}
\newpage

\subsection{Communication and participation to external events}

Dissemination activities also include the participation of \ODK
members to many different conferences of various size and topics
in computer science, mathematics, physics, and more. The goal is
to reach potential end-users, build bridges between communities and stay aware 
of current development in the scientific community.

We list here major events and communication. We have also put in place
a blog on our website: \url{http://opendreamkit.org/activities/} to track
these activities.

\input{events/2015-10-PyConFr}

\input{events/2015-11-thirty-years-atlas}

\input{events/2016-01-MMM}

\input{events/2016-05-se4science}

\input{events/2016-05-PyCon}

\input{events/2016-07-talk-lisbon}

\input{events/2016-08-EuroSciPy}


\section{Upcoming events and plans for the future}

This past year, \ODK has been organizing or co-organizing
14 different events, thus helping and supporting the open-source
communities on which the project is built. Some of these events had
been announced in the proposal: the Sage development workshop (event \ref{event-sd77}) , the Atelier
Pari (event \ref{event-AtelierPARI2016}), the FPSAC Sage Days (event \ref{event-sd78}), as
well as the events in developing countries. We have also decided
to support extra events which were directly helping \ODK goals
or beneficial for the entire community. 
We believe that this persistent presence of the \ODK project within
the open-source community activities is a key part of our success 
and we plan on keeping it on. The next few years look just as much eventful
as this past year. 

\begin{itemize}
\item Some \textbf{development workshops} are already planned such as \textbf{Atelier Pari} in Lyon,
a \textbf{high performance computing} in Grenoble and probably at least one \textbf{Sage Days} in Orsay.

\item We are working on three \textbf{major dissemination events} that should tentatively take place
at CIRM, Dagstuhl, and ICMS. We have formed organizing committees and are working on proposals.

\item The first ever \textbf{Women in Sage} in Europe has been announced and will take place
in France in January.
\end{itemize}

\bibliographystyle{plain}
\bibliography{report}

\footnotesize{All referenced ODK talks are available as annexes}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
