\hypertarget{deliverable-description-as-taken-from-github-issue-41-on-2019-12-10}{%
\section*{\texorpdfstring{Deliverable description, as taken from Github
issue
\href{https://github.com/OpenDreamKit/OpenDreamKit/issues/41}{\#41} on
2019-12-10}{Deliverable description, as taken from Github issue \#41 on 2019-12-10}}\label{deliverable-description-as-taken-from-github-issue-41-on-2019-12-10}}

\begin{itemize}
\tightlist
\item
  \textbf{WP2:}
  \href{https://github.com/OpenDreamKit/OpenDreamKit/tree/master/WP2}{Community
  Building, Training, Dissemination, Exploitation, and Outreach}
\item
  \textbf{Lead Institution:} Université Paris-Sud
\item
  \textbf{Due:} 2019-08-31 (month 48)
\item
  \textbf{Nature:} Websites, Media, etc.
\item
  \textbf{Task:} T2.1
  (\href{https://github.com/OpenDreamKit/OpenDreamKit/issues/24}{\#24})
\item
  \textbf{Proposal:}
  \href{https://github.com/OpenDreamKit/OpenDreamKit/raw/master/Proposal/proposal-www.pdf}{p.38}
\item
  \textbf{\href{https://github.com/OpenDreamKit/OpenDreamKit/raw/master/WP2/D2.16/report-final.pdf}{Report}}
  (\href{https://github.com/OpenDreamKit/OpenDreamKit/raw/master/WP2/D2.16/}{sources})
\end{itemize}

This deliverable briefly reviews the efforts of the consortium within
Work Package 2 (Community Building, Training, Dissemination,
Exploitation, and Outreach) to communicate the outcomes of the project
to the broader audience. This includes the production of Press Releases,
an important part of our communication strategy. The aim is to call
attention of multiple audiences about our research in a way
understandable by non-specialists. It also is to address the EU
perspective of research and innovation funding, by considering some
aspects such as the European transnational cooperation in the consortium
and the scientific excellence with a better use of results to the
scientific community.

Under the leadership of the coordinating institution (UPSud), the
beneficiaries prepared collectively a joint text. This text served as
template that each beneficiary could adapt to suit its own circumstances
and translate into the local language as required. The produced press
releases were then circulated to the respective press or communication
offices in early December 2019. This also includes other forms of
end-of-project communication through publications in EC and non EC
channels, OpenDreamKit's web site, community mailing lists, and social
media.

TODO (not to be included in the final deliverable report):

Press releasing process:

\begin{itemize}
\item
  \textbf{CONTACT} the Press Office or Communication department of your
  institution (to identify if the topic is suitable for a press release
  or not, or if there is another way to get media attention such as
  Press article published in the University website or through the
  newsletter). Best deadline: Next week.
\item
  \textbf{TOPIC APPROVAL} by the Communication Department/ Press Office,
  \textbf{deadline 15 days before the publication}.
\item
  \textbf{WRITING}: The press officer is usually in charge of writing
  the press release and distribute it to the general mainstream press.
  The content has to be easy to communicate to journalists. The UPSud
  communication department is preparing a general text that you will
  receive and could adapt at your convenience.
\item
  \textbf{RELEASING}: The approved release will then be sent to relevant
  journalists contacts of your institution ( general press in the high
  education, research area and local press) and uploaded to newswire
  services. Deadline for publication (to negociate with your respective
  communication departments): \textbf{before the Review meeting}.
\end{itemize}

TODO ( Contact Press offices to plan a Press release)

\begin{itemize}
\tightlist
\item
  \(\checkmark\) UPSud, @IzabelaFaguet , @nthiery
\item[$\square$]
  CNRS, @videlec
\item[$\square$]
  UNIKL, @wdecker
\item[$\square$]
  UOXF, @dimpase
\item[$\square$]
  USlaski, @marcinofulus
\item[$\square$]
  USTAN, @stevelinton, @alex-konovalov
\item[$\square$]
  UVSQ, @defeo
\item[$\square$]
  UWarwick, @JohnCremona
\item[$\square$]
  Logilab, @ocayrol
\item[$\square$]
  Simula, @minrk
\item[$\square$]
  Ugent, @jdemeyer
\item[$\square$]
  XFEL, @fangohr
\item
  \(\checkmark\) Fau, @florian-rabe @kohlhase
\end{itemize}
