\documentclass{deliverablereport}

\deliverable{dissem}{press-release-2}
\deliverydate{XX/YY/201Z}
\duedate{31/08/2019 (M48)}
\author{Izabela Faguet, Viviane Pons, Nicolas M. Thiéry}

\begin{document}
\maketitle
% This will be the abstract, fetched from the github description
\githubissuedescription

% write the report here

\section{Executive summary}

\TODO{Merge with the github issue description}

The present document constitutes Deliverable D2.16 (Press Release) in the framework of WP2
(Community Building, Training, Dissemination, Exploitation, and Outreach), regarding Tasks 2.1 
(Dissemination and Communication activities). Press releases are an important 
part of our communication stategy. With the press releases we plan to call
attention of multiple audiences about our research in a way understandable
by non-specialists and address the EU perspective of research and innovation 
funding, by considering some aspects such as the European transnational 
cooperation in the consortium and the scientific excellence with a better use
of results to the scientific community.

\section{End of project press releases}

\TODO{Merge the following sections}

During the first months of the project, we have covered 6 press releases and we
planned to do the same at the end to promote the actions and the results by providing
targeted information to broad audience (including the media and the public).
First we will submit a general communication about the project on our website presenting 
the results of the assessment made by the referees and the Project Officers during our 
final review meeting.

\section{Collaborative process}

The press releases are the result of a joint effort of the partners
who are actively involved in the promotion of the press releases;
their collective effort will result in at least 12 releases in general
press in the high education, research area and local press. We
followed our usual open bottom up approach. % insérer schéma

\section{Press releasing process}

 1. The coordinator have already contacted the Press Office of UPSud to inform about the topic
 of the project that was already approved by the Communication officer. 
 2.The UPsud communication department was in charge of preparing a general text that will be sent
 to the partners and that could be adapted at their convenience. We received the article proposal 
 and some changes were made. it has been translated in English and sent to the partners 
 for distribution to their respective Press offices. 
% insérer une annexe avec l'article et insérer un lien de renvoi
 3.The approved release will then be sent to relevant journalists contacts of each institution 
 and uploaded to the newswire services. The press releases will be published by end of november. 
 Each partner still needs to negociate with their respective communication departments the 
 definitive deadline for publication.
 4. We also plan to launch email campaigns in order to report to the scientific community
 about the significant developments on ODK project activities and results. 

\section{End of project communication through OpenDreamKit's channels}

OpenDreamKit's website (\url{https://opendreamkit.org/}) served as a
major dissemination tool in terms of project’s concept, objectives and
outcomes. It provides information on the project content, objectives,
partners, planned activities, events organized within the project and
deliverables. It's complemented by a
\href{https://twitter.com/OpenDreamKit}{@OpenDreamKit} Twitter
account, and a video channel. Both serve as direct communication
instrument to reach the general public and interact with other
communities such as GAP, SageMath, Cocalc, or Jupyter.

An overview of the project outcomes was produced by UPSud with the
technical contribution of the ODK partners (see appendix XXX), and
will be published on OpenDreamKit's web site; it takes the form of an
extended version of the Summary for Publication produced at the
occasion the last reporting period. It complements the press release
by being more detailed on the concrete actions and the impact on end
users, with link to related material (reports, blog posts, ...). Soon
after its publication, a blog about the future of OpenDreamKit
community will posted, and both will be advertised through social
media (twitter, ...).

% In order to reach a broad target, the presence of the ODK project in social media was
% also encouraged. 

% The website will have direct access to these social networks by clicking over the icon 
% situated on a visible part of the website. In this way, it will be easy for every user 
% to participate in this when the website is visited.

During the project, seven video interviews have been created to
communicate the project, its main goals and the consortium to a wider
public; and 3 explainer videos that explaines our project in a way
understandable for a non scientific public. The videos will be used
across online platforms, and promoted through the project’s online
communications channels such as YouTube. A Youtube account for
OpenDreamKit project was set up to host the video series. The number
of views, and comments for each video were monitored as KPIs in the
technical report.

\subsection{End of project communication through the European Commission's channels}

The Coordinator has contacted the communication officer  who will support the dissemination
of the results of the project ad will promote them through the EC's communication channels. 
The communication officer will publish an article and a presentation of the project in the 
newsletter ” Digital Excellence \& Science Infrastructure “ of the European Commission, 
on the European Commission website \url{https://ec.europa.eu/digital-single-market/} to reach 
a wider audience, potentiating it outreach. 
HORIZON, the EU Research \& Innovation e-magazine, will also be used to inform about 
the benefits and progress that ODK will generate in Europe, informing about 
the open debates created and the results.

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
