\begin{event}{Sage Days 79}{SD79}{Jerusalem, Nov. 21 -- Nov. 24, 2016}{PS, UB}{46}{4}{https://wiki.sagemath.org/days79}

\textbf{Main goals.} This workshop was dedicated to the thematics ``geometric combinatorics and symbolic dynamics'' in \Sage. It was also a training program
for the local community in Israel with participants from the main universities (Herbrew University, Bar-Ilan University, Beer-Sheva, Ben-Gurion University, Tel-Aviv University, University of Haifa, Weizmann Institute)

\textbf{ODK implication.} The workshop was mostly organized locally by Jean-Philippe Labbé and funded by the ERC Grant entitled ``Avenues in Probabilistic and Geometric Combinatorics''. ODK was supporting the event by...

\textbf{Event summary.} The event started by introduction talks and tutorials as well as installation sessions about \Sage so that new comers could be initiated and guided. The rest of the week featured many specific tutorials about some mathematics features related to the thematic of the workshop. Lots of time was let open for coding and projects.

\textbf{Demographic.} 29 participants out of the 46 came from local universities in Israel. 

\textbf{Results and impact.} As the first Sage Days in Israel, the main objective of this meeting was to introduce the software to local participants. Around half of the participants were beginners, we made around 15 installations of \Sage on Windows, Mac and Linux. We allocated a lot of time for beginners to learn from tutorials but also directly from the more experienced users by personalized help. Having a flexible schedule allowed to have such help sessions.

Here is a summary of the accomplishments of the week:

\begin{itemize}
    \item    Worked was carried out on 47 \Sage tickets: \url{https://trac.sagemath.org/query?keywords=~days79}
    \item    Around 15 installations on Windows, Mac and Linux.
    \item    Creation of a sample \Sage package to document the process.
    \item    Around 10 participants got to know Sage better doing tutorials
    \item    Experienced Sage users answered tons of questions surrounding usage of Sage
    \item    We got 5-6 more experienced users to contribute to Sage by reviewing, reporting, writing tickets
    \item    Many discussions and healthy debates about view, plot, show and polytope in Sage
    \item    Some projects and discussions will be continued through future collaborations and meetings in 2017 
\end{itemize}


\end{event}