\begin{event}{Atelier PARI/GP 2017b}{AtelierPARI2017b}{Clermont-Ferrand (FR),
2017-06-19 to 2017-06-23}{PS,UB,UV,UW}{36}{http://pari.math.u-bordeaux.fr/Events/PARI2017b/}

\textbf{Main goals.}

The PARI/GP Ateliers were established in 2012 as a yearly meeting
between developers and users of the PARI/GP system.

The main goals are advertising new features and improvements,
discussing further developments, sharing best practices, and collaborative
code writing (hacking sessions, doc reviews, bug-squashing parties).

You can find the list of previous PARI Ateliers at
\url{http://pari.math.u-bordeaux.fr/ateliers.html}

\textbf{ODK implication.} 
%Describe how ODK was involved and give a rough estimation of cost for ODK

\ODK participants: B. Allombert, K. Belabas, J.-P. Flori

\ODK provided the main funding source for the workshop (accommodation,
subsistence and travel expenses), for about XXXk\euro. The following
institutions co-funded the event through institute of mathematics (Laboratoire
  Blaise Pascal): GDR Structuration de la Théorie des Nombres (CNRS), Région
  Auvergne Rhône-Alpes, Conseil départemental Puy-de-Dôme, Université
  Auvergne-Rhônes-Alpes and the city of Clermont-Ferrand.

\textbf{Event summary.} 
%Give a summary of your event

The 8th Atelier PARI/GP took place in Clermont-Ferrand (France) from june
19th to 23rd with a special focus on Elliptic curves, Modular Forms and
$L$-Functions.

There were 36 registered participants from 16 different institutions
(no registration fees).

A typical day of the workshop had introductory talks and tutorials
in the morning; afternoons allowed ample time for hacking sessions,
discussions and training.

The Atelier featured 13 morning talks.

\textbf{Results and impact.} 
% What did you achieve with this event? (If ever it impacted 
% other ODK tasks and deliverables, mention it here)
This was a dissemination event with a special focus on recently developped
packages (Elliptic curves over number fields, Modular forms), before they were
merged into a formal release. We got lots of feedback on the interface and
documentation and a number of bugs and glitches were fixed as a result.
\end{event}
