\begin{event}{Sage/GAP Days 86: Algebra and Words in Combinatorics }{sd86}{Montréal (Canada) 2017-04-17 to 21}{PS}{31}{2}{https://wiki.sagemath.org/days86}

  \textbf{Main goals.} This training workshop was organized at the
  occasion of a four month CRM thematic semester
  [Algebra and words in combinatorics(http://www.crm.umontreal.ca/Combinatorics2017/index.php/)
  at the [Laboratoire de combinatoire et d'informatique mathématique (LaCIM)](http://lacim.uqam.ca/en/) of
  [Université du Québec à Montréal](http://www.uqam.ca/).

  Taking advantage of the participation of Nicolas M. Thiéry to the
  thematic semester (as member of the scientific committee), this
  workshop was followed by a
  \href{https://wiki.sagemath.org/Montreal}{weekly half-day meeting}
  around computational tools for combinatorics until July 2017, and
  \href{https://more-sagemath-tutorials.readthedocs.io/en/latest/2017-05-29-CRM/}{computational
    sessions} at the occasion of certain spring schools of the
  semester.

  \textbf{ODK implication.} The Sage Days was organized by Nicolas M.
  Thiéry (\ODK coordinator, Paris Sud) and Sébastien Labbé (since then
  \ODK participant at CNRS). \ODK funded some of the invited speakers
  (\approx 4k€). Later events were organized by Nicolas. Nicolas's
  participation to the semester (\approx 6k€) was funded by a
  combination of \ODK's direct and indirect costs.

  \textbf{Event summary.} An intensive week with many brainstorms and
  coding sprints + follow up meetings.

  \textbf{Demographic.} 31 participants (including 2 from \ODK), about
  half of which being female.

  \textbf{Results and impact.} Beside training many new \Sage users
  and developers, the workshop and followup events boosted an active
  local user group at LaCIM, including a new generation of graduate
  students and postdocs, but also François Bergeron, a prominent
  figure in experimental combinatorics who used the occasion to switch
  to \Sage. Discussions prompted the organization by this group of
  \Sage Days 95, a one week Women in \Sage workshop in July 2018,
  where \ODK member (Sheffield) and Research Software Engineer Tania
  Allard was invited as role model.

  Altogether
  \href{https://trac.sagemath.org/query?status=closed&status=needs_info&status=needs_review&status=needs_work&status=new&status=positive_review&keywords=~days86&col=id&col=summary&col=status&col=time&col=changetime&col=author&col=reviewer&order=status}{18
    \Sage tickets} were actively worked on during the week.
\end{event}
