\begin{event}{Atelier PARI/GP 2017c}{AtelierPARI2017c}{Oujda (MO),
2017-11-22 to 2017-11-23}{UB}{36}{6}{http://pari.math.u-bordeaux.fr/Events/PARI2017c/}

\textbf{Main goals.}

This was a teaching and dissemination meeting, by invitation from the math
institute in Oujda as a satellite event for their ``Journ\'ees Alg\`ebre,
  Th\'eorie des Nombres et Application''.

\textbf{\ODK implication.}
%Describe how \ODK was involved and give a rough estimation of cost for \ODK

\ODK participants: B. Allombert and A. Page from Bordeaux.

\ODK funded travel costs for the two instructors for about 1k\euro. The
  organizing Oujda institute of mathematics co-funded the event, paying for
  all local expenses.

\textbf{Event summary.}
%Give a summary of your event

This Atelier PARI/GP took place in Oujda (Morocco) from november 22nd to
23rd, it was followed by a 2-day Moroccan national research conference on
  Algebra, Number Theory and their Applications. There were 70 participants
  from all over Morocco (no registration fees).

The 2-day Atelier featured a general introduction to PARI/GP and two
  specialized courses in the mornings (graduate level):
\begin{itemize}
\item Bill Allombert ``Cryptographie et courbes elliptiques'',
\item Aurel Page ``Th\'eorie alg\'ebrique des nombres''.
\end{itemize}
Afternoons were devoted to practice sessions. Bill Allombert also gave
  a research talk on ``Aspects combinatoires et algorithmiques des fonctions
  $L$ d'Artin'' in the ensuing conference.

Slides for all talks are available at
\url{http://pari.math.u-bordeaux.fr/Events/PARI2017c/}

\textbf{Results and impact.}
% What did you achieve with this event? (If ever it impacted
% other \ODK tasks and deliverables, mention it here)

This was the first purely instructional PARI/GP event, it allowed us to test
the format and teaching material, thereby paving the way for future such events.
\end{event}
