\begin{event}{ICMS 2018 -- Session 11: Backtrack search techniques in groups and combinatorics}{ICMS18-PBT}{South Bend, USA, 25th July 2018}{UStan}{15}{1}{http://icms-conference.org/2018/sessions/session11/}

\textbf{Main goals.}
We invited experts from AI, combinatorics, computational group theory and
related areas with the aim of sharing and exchanging ideas, problems,
results and implementations.

\textbf{ODK implication.} Organised by Markus Pfeiffer (UStan), and his
colleague Christopher Jefferson. Some of Markus Pfeiffers work improves the
software tools used in backtrack search in GAP.

ODK funded Markus Pfeiffer's travel to ICMS and the conference fee, and the
accommodation for the session participants Paula Hähndel and Wilf Wilson (both
from Halle, Germany).

\textbf{Event summary.} We saw a great selection of talks from experts around
the world on the current state of the art in backtrack search: Chris Jefferson
on partition backtrack, Robin Candy (PhD student of Brendan McKay) on canonical
labelling subgroups under conjugacy, Paula Hähndel on using Orbital Graphs
(which are used in partition backtrack and canonical labelling), Anton Betten
and Matthias Koch on ``Snakes and Ladders'', Abdullah Al-Azemi (Kuwait
University) on HPC-Techniques in backtrack search, and Markus Pfeiffer on
backtrack search in the free group.

In particular, two popular techniques ``Partition Backtrack'' and ``Snakes and
Ladders'' were discussed, and for the first time in recent decades experts in
this are were able to sit together and learn from each other how to combine the
two approaches to an even more powerful one.

\textbf{Results and impact.} New collaborations were setup, Robin Candy is going
to visit Halle (Germany) and St Andrews later this year to work on improving
backtrack search for canonical labels under subgroup conjugacy. Wilf Wilson from
Halle who is now working on a joint project with Chris Jefferson, Rebecca
Waldecker and Markus Pfeiffer was able to meet experts and get an introduction
to the field.

\end{event}
