\begin{event}{Reproducible Python}{SHF_Pycon2018}{Cleveland, USA, May 2018}{USH}{60}{1}{https://us.pycon.org/2018/}

\textbf{Main goals.} The PyCon 2018 conference, which took place in Cleveland, is the largest annual gathering for the community using and developing the open-source Python programming language. It is produced and underwritten by the Python Software Foundation, the 501(c)(3) nonprofit organization dedicated to advancing and promoting Python. Through PyCon, the PSF advances its mission of growing the international community of Python programmers.

\textbf{ODK implication.} ODK Sheffield member, Tania Allard, developed and delivered a workshop using ODK supported technology at PyCon2018.

\textbf{Event summary.} Workshop and open materials on Reproducible analysis in Python: these materials cover the essentials on how to develop workflows with a reproducibility approach.

The workshop was first delivered in PyCon 2018 to over 60 attendees from all over the world.

The materials were further extended to add more content on Jupyter notebook validation using ODK-developed nbval and also on property based testing

All the materials are licensed under CC-BY and can be found at \url{https://github.com/trallard/ReproduciblePython} and are also shared using binder

\textbf{Results and impact.} As a follow up for this workshop Tania Allard was been invited to give a talk about reproducibility in data pipelines at the RAPIDS conference in London (\url{https://www.dotmesh.com/blog/rapids-2018/})

\end{event}
