\section{Project Requirements}\label{sec:requirements}

\subsection{Employment Status Information}\label{sec:req:employment-status}

\begin{todo}{from the proposal template}
  For each applicant, state the last name, first name, and employment status (including
  duration of contract and funding body, if on a fixed-term contract).
\end{todo}

\subsection{First-time Proposal Data}\label{sec:req:first}

\begin{todo}{from the proposal template}
  Only if applicable: Last name, first name of first-time applicant.

  If this is your first proposal, reviewers will consider this fact when assessing your
  pro- posal. Previous proposals for research fellowships, publication funding, travel
  allow- ances, or funding for scientific networks are not considered first proposals. If
  you are submitting a “first-time proposal” and it is part of a joint proposal, please
  note that your independent project must be distinct from the other projects.

  If you have already submitted a proposal as an applicant for a research grant and have
  received a letter informing you of the funding decision, or if you have led an independ-
  ent junior research group or project in a Collaborative Research Centre or Research
  Unit, you are no longer eligible to submit a “first proposal”. If you have submitted a
  “first-time proposal” and it was rejected, you may resubmit the application, in revised
  form, as a first-time proposal for the same project.
\end{todo}

\subsection{Composition of the Project Group}\label{sec:req:group}

\begin{todo}{from the proposal template}
  List only those individuals who will work on the project but will not be paid out of the
  project funds. State each person’s name, academic title, employment status, and type of
  funding.

  Please list separately the individuals paid by your institution and those paid using
  other third-party funding (including fellowships).
\end{todo}

\begin{sitedescription}{jacu}
  The KWARC (Knowledge Adaptation and Reasoning for Content) research group headed by
  Michael Kohlhase for has the following members
  \begin{compactdesc}
  \item[Dr. N.N.] is the \ldots She has a background in\ldots.
  \end{compactdesc}
  Additionally, the group has attracted about 10 undergraduate and master's students that
  actively take part in the project work and various aspects of research.
\end{sitedescription}

\begin{sitedescription}{pcg}
  Power Consulting GmbH is the leading provider of semantic document solutions. Dr. Senior
  Researcher leads an applied research group consisting of
  \begin{compactdesc}
  \item[Dr. N.N.] is the \ldots She has a background in\ldots.
  \end{compactdesc}
  The group has access to seven programming slaves specializing in web development and
  document transformation techniques
\end{sitedescription}


\subsection{Cooperation with other Researchers}\label{sec:coop}

\subsubsection{Planned Cooperations}\label{sec:coop:planned}
\begin{todo}{from the proposal template}
  Researchers with whom you have agreed to cooperate on this project
\end{todo}
\begin{compactdesc}
\item[Prof. Dr. Super Akquisiteur (Uni Paderborn)] knows exactly what to do to get funding
  with DFG, we will interview him closely and integrate all his intuitions into the {\pn}
  templates.
\item[Prof. Dr. Habe Nichts (Uni Hinterpfuiteufel)] has never gotten a grant proposal
  through with DFG, we will try to avoid his mistakes.
\item[Dr. Sach Bearbeiter (DFG)] will consult with the DFG requirements to be met in the
  proposals.
\item[Dr. Donald Knuth (Stanford University)] is so surprised that we want to do grant
  proposals in {\TeX/\LaTeX} that he will help us with any problems we have in coding in
  this wonderful programming language.
\end{compactdesc}

\subsubsection{Scientific Collaborations in the past Three Years}\label{sec:past-coop}

\begin{todo}{from the proposal template}
  Researchers with whom you have collaborated scientifically within the past three years

  This information will assist the DFG’s Head Office in avoiding potential conflicts of
  in- terest during the review process.
\end{todo}

\subsection{Scientific Equipment}\label{sec:req:equipment}

Jacobs University provides laptops or desktop workstations for all academic
employees. Great Consulting GmbH. is rolling in money anyways and has all of the latest
gadgets.


\subsection{Project-Relevant Interests in Commercial Enterprises\qquad n/a}

%%% Local Variables: 
%%% mode: LaTeX
%%% TeX-master: "proposal"
%%% End: 

% LocalWords:  Durchf uhrung subsubsection ipower Hinterpfuiteufel Sach Aktivit
% LocalWords:  Erkl arungen
