\section{Introduction}\label{sec:intro}

There is a large and vibrant ecosystem of open-source software systems for mathematics. 
These range from calculators, which perform simple computations, via mathematical databases, which curate collections of a mathematical objects, to powerful modeling tools and computer algebra systems (CAS).

Most of these systems are very specific -- they focus on one or very few aspects of mathematics.
For example, among databases, the ``Online Encyclopedia of Integer Sequences'' (OEIS) focuses on sequences over $\mathbb{Z}$ and their properties, and the ``L-Functions and Modular Forms
Database'' (\LMFDB)~\cite{Cremona:LMFDB16,lmfdb:on} on objects in number theory pertaining to Langland's program.
 Among CAS, \GAP~\cite{GAP:on} excels at discrete algebra with a focus on group theory, \Singular~\cite{singular:on} focuses on polynomial computations with special emphasis on commutative and non-commutative algebra, algebraic geometry, and singularity theory, and \Sage~\cite{SageMath:on} aims to be a general purpose software for computational pure mathematics by loosely integrating many systems including the aforementioned ones.

For a mathematician, however, (a user, which we call Jane) the systems themselves are not relevant.
Instead, she only cares about being able to solve problems. 
Because it is typically not possible to solve a mathematical problem using only a single program, Jane has to work with multiple systems and combine the results to reach a solution. 
Currently there is very little tool support for this practice, so Jane has to isolate sub-problems that the respective systems are amenable to, formulate them in the respective input language, collect intermediate results and reformulate them for the next system -- a tedious and error-prone process at best, a significant impediment to scientific progress at worst.
Solutions for some situations certainly exist, which can help get Jane unstuck, but these are ad-hoc and only for specific often-used system combinations. 
Moreover, each of these ad hoc solutions requires a lot of maintenance and scales badly to multi-system integration. 

One goal of the OpenDreamKit project is tackling these problems systematically by building virtual research environments (VRE) on top of the existing systems.
To build a VRE from individual systems, we need a joint user interface -- the OpenDreamKit project adopts Jupyter~\cite{jupyter-project:on} and active documents~\cite{KohDavGin:psewads11} -- and an interoperability layer that allows passing problems and results between the disparate systems. 
For the latter, it proposes the Math-in-the-Middle (MitM~\cite{DehKohKon:iop16}) paradigm, an interoperability framework based on a central, system-independent ontology of mathematical knowledge.
In this paper we instantiate the MitM paradigm in a concrete case study using a distributed computation involving \GAP, \Sage, and \Singular.

We will use the following running example from computational group theory: Jane
wants to experiment with invariant theory of finite groups. She works in the
polynomial ring $R=\mathbb{Z}[X_1,\ldots,X_n]$, and wants to construct an ideal $I$
in this ring that is fixed by a group $G \leq S_n$ acting on the variables,
linking properties of the group to properties of $I$ and the quotient of
$R$ by $I$.

To construct an ideal that is invariant
under the group action, it is natural to pick some polynomial $p$ from $R$
and consider the ideal $I$ of $R$ that is generated by all elements of the orbit
$O=Orbit(G,R,p)\subseteq R$.
For effective further computation with $I$, she needs a G\"obner base of $I$.

Jane is a \Sage user and wants to receive the result in \Sage, but she wants to
use \GAP's orbit algorithm and \Singular's Gr\"obner base algorithm, which she
knows to be very efficient. For the sake of example, we will work with $n=4$,
$G=D_4$ (the dihedral group\footnote{Incidentally, this group is called $D_4$ in
\Sage but $D_8$ in \GAP due to differing conventions in different mathematical
communities -- a small example of the obstacles to system interoperability that
MitM tackles.}), and $p=3\cdot X_1+2\cdot X_2$, but our results apply to
arbitrary values.

In Section~\ref{sec:mitm}, we recap the MitM paradigm.
MitM solutions consist of three parts: a central ontology, specifications of the abstract languages of the involved systems (which we call \emph{system dialects}), and the distributed computation infrastructure that connects the systems via the ontology as an intermediate representation.
The rest of the paper develops these three parts for our case study:
In Section~\ref{sec:cgt}, we contribute a fragment to the MitM ontology that formalizes computational group theory.
In Section~\ref{sec:apit}, we specify the abstract languages of \GAP, \Sage, and \Singular and their relation to the ontology.
Finally in Section~\ref{sec:case}, we present the resulting virtual research environment built on these systems in action.
Section~\ref{sec:concl} concludes the paper and compares MitM-based interoperability with other approaches.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

%  LocalWords:  sec:intro Sloane:oeis12,oeis mathbb Cremona:LMFDB16,lmfdb:on GAP:on \Sage:on singular:on DehKohKon:iop16 emph sec:mitm sec:cgt sec:apit sec:mitm_poc sec:concl MitM-based sec:case Jupyter jupyter-project:on KohDavGin:psewads11 ldots,X_n leq obne cdot cdot
%  LocalWords:  SageMath:on ldots,X_4 subseteq obner Groebner formalizations
