\documentclass{deliverablereport}

\usepackage[style=alphabetic,backend=bibtex]{biblatex}
\addbibresource{report.bib}
\addbibresource{../../lib/publications.bib}

\usepackage{xparse}
\usepackage{etoolbox}
\usepackage{caption}
\graphicspath{{events/}}
 \ExplSyntaxOn

\newcounter{eventcounter}

\newenvironment{event}[7]{
\vspace{0.5cm}
\refstepcounter{eventcounter}
\label{event-#2}
\addcontentsline{toc}{subsection}{\hspace{3ex}-\ Event~\theeventcounter -~ #1}


\noindent\textbf{Event~\theeventcounter -~ #1}\newline % title

\noindent #3 \newline % location and date

\noindent ODK~partners~involved:~ \clist_map_inline:nn{#4}{\site{##1}~}\newline %partners

\ifx&#5&%
      % no participant #
\else
\noindent #5~participants~
\ifx&#6 &%
    % no odk participant
\else 
~(including~#6~from~within~ODK)\newline
\fi
\fi

\ifx&#7&%
      % no website
\else
\noindent \url{#7}\newline
\fi



}{\begin{center}\noindent\rule{4cm}{0.4pt}\end{center}}

 \ExplSyntaxOff


\deliverable{dissem}{workshops-4}
\duedate{31/08/2019 (M48)}
\deliverydate{09/10/2019}
\author{Izabela Faguet, Viviane Pons, Nicolas M. Thiéry et al.}

\begin{document}
\maketitle
\githubissuedescription
\newpage
\tableofcontents
\newpage

\section{Development workshops}

Throughout year 4 of the project, we organized seven development
workshops with a limited number of participants who met to work on a
specific task. These workshops are an inherent part of \ODK
development process as described in \taskref{dissem}{devel-workshops}:
they bring together developers from within and outside of \ODK and
allow effective work and discussions on many technical aspects. They
also participate in building and maintaining a community of developers
inside \ODK and within the open-source communities we belong to. Some
of the workshops included a training and/or dissemination approach.

\input{events/2018-12-04-LinBoxWinterDays}

\input{events/2019-01-Atelier-PARI-GP}

\input{events/2019-06-11-XFEL-mumax.tex}

\input{events/2019-06-15-SageGapDays101}

\input{events/2019-08-15-GAP-Singular-school-meeting}

\input{events/2019-08-MathDataWorkshop}

\input{events/2019-08-24-ReportWritingSprint.tex}

\clearpage
\section{Dissemination and outreach activities}

We describe here all activities related to \taskref{dissem}{dissemination}:
these are all events oriented towards dissemination, training, and outreach. This
includes events organized or co-organized by \ODK and also
participating in external events and many communication activities.

\subsection{Training and community building workshops in established mathematical communities}

We have organized or coorganized a series of workshops to support
communities of mathematicians in their use of open source
computational software and VRE. Such workshops typically consist of a
blend of tutorials at all levels and coding sprints with the following
aims:

\begin{itemize}
\item \textbf{Training and attracting more users}: this is the perfect
  place for newcomers, especially students, to get acquainted with the
  systems; meanwhile, more experienced users get training on more
  advanced topics and on how to collaborate and share their
  production.

\item \textbf{Disseminating and evaluating \ODK outcomes}: running
  many training sessions give ample opportunities to witness how
  end-users adopt or struggle with the \ODK outcomes, and how they fit
  their needs.

\item \textbf{Improving the overall quality}: by fostering researchers
  in specific areas, Sage Days help bring interesting mathematics into
  the software, which is beneficial for Sage and so \ODK.

\item \textbf{Community building}: the workshops foster a
  collaborative spirit where everyone is encouraged to share their
  expertise, help each other, make teams to tackle shared needs, and
  contribute back.
\end{itemize}


\input{events/2018-09-24-STM-Koper}

\input{events/2018-11-19-Aachen_school}

\input{events/2019-02-11-CIRM-FreeComputationalMathematics}

\input{events/2019-03-04-heraklion}

\input{events/2019-04-Atelier-PARI-GP}

\input{events/2019-07-08-SageDays105.tex}

\input{events/2019-07-22-SageDays100.tex}

\input{events/2019-08-04-ESSLLI-Zoo.tex}


\subsection{Fostering diversity}

\ODK is deeply concerned with the diversity gap that exists in science
in general and more specifically in mathematics and software
development. Here are the events we organized during this period to
support specifically women and developing countries in sciences.

\input{events/2019-04-08-WomenInSage}

\input{events/2019-07-15-SageDays102}

\subsection{Research workshops}

\input{events/2019-02-04-JNCF}

\subsection{Communication and participation in external events}

\input{events/2018-09-17-GAPDaysFall2018}

\input{events/2018-10-06-PyconFR}

\input{events/2018-11-01-jupyterlab}

\input{events/2019-03-CIRM}

\input{events/2019-03-18-GAPDaysSpring2019}

\input{events/2019-05-27-BigProof.tex}

\input{events/2019-05-29-nbgrader}

\input{events/2019-07-01-FPSAC}

\input{events/2019-07-08-CICM.tex}

\input{events/2019-07-10-CICM-LML.tex}

\input{events/2018-08-27-AMRS}

%\newpage\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
