\begin{event}{AMRS2018 --- Conference on Applied Mathematics and Random Structures}{AMRS2018}{Birzeit, Palestine, 27th-30th of August 2019}
{Paris 13 University}{41}{1}{https://lipn.univ-paris13.fr/~nicodeme/birzeit18RAND/}

\textbf{Main goals.}AMRS is a conference intended to foster collaborations between Palestinian and French universities or research centres; 
to sustain the critical efforts of the mathematics community in Palestine to set up a doctoral school. AMRS provided an opportunity for dissemination of important and recent results in the field of Applied Mathematics and Statistics with Applications to Economy, Industry, and Science.

\textbf{\ODK implication.}
Thierry Monteil, an ODK member gave a tutorial for the freeware computer algebra system SAGE during the conference. 
There were no costs to \ODK other than travel costs for him.

\textbf{Event summary.}
AMRS conference gathered 41 attendees around various fields of Applied mathematics: Statistics, Game The-ory, Analytic Combinatorics, 
Probability, Differ-ential or Difference Equations, Matrices, Stochas-tic Optimal Control Problems, Partial DifferentialEquations, 
Binary Search Trees, Ideals on SkewMatrices. 
There were 13 talks given by 7 Palestinian professors, and 6 French); and 10 talks by students:

\textbf{Results and impact.}
Thierry Monteil has introduced  the  basic  concepts  of  SAGE  and  present  many examples based on graphs algorithms to 24 students

19 from Birzeit–
3 from Al Quds–
1 from Polytechnic Palestine UniversityHebron (PPU)–
1 from Malaya
There was a large majority of female students in the audience,19 out of 24 students

\end{event}
