\begin{event}{JNCF2019: Journées Nationales de Calcul Formel}{JNCF2019}{Luminy (FR),
  2019-02-04 to
  2019-02-08}{UV,UJF}{58}{3}{http://www.jncf2019.uvsq.fr/}
  
\textbf{Main goals.} This is the yearly meeting of the French community in
Computer Algebra, open to an international audience, with lectures and
contributed talks given mostly in English.

\textbf{\ODK implication.} Luca De Feo co-organized the JNCF 2019. The
organization of the workshop was coordinated with the organization of
the ``Free Computational Mathematics'' conference, taking place the
following week in the same venue, to encourage cross-participation.

Luca De Feo and Clément Pernet gave presentations on Monday on the
achievements of \ODK relevant to the Computer Algebra
community.

\ODK participants: Luca De Feo from University of Versailles,
Jean-Guillaume Dumas and Clément Pernet from University Grenobles
Alpes.

\ODK sponsored the event (3000 euros %\TODO{Double-check the figure}).

\textbf{Event summary.} This conference took place in Luminy (France) from
February 4th to 8th. About 58 mathematicians and computer scientists
participated to the event. Three of the participants also participated
in the ``Free Computational Mathematics'' conference the following
week, while four other noted members of the French community in
Computer Algebra, Joris van der Hoeven, Marie Françoise Roy, Fredrik
Johansson and Bill Allombert, elected to only participate in the latter
due to time constraints.

Slides for Luca De Feo and Clément Pernet's presentations are
available online from the conference web page:
\url{http://www.jncf2019.uvsq.fr/edt.html}.

\textbf{Results and impact.} Computer Algebra is the birth place of
computer-aided Mathematics, and all of \ODK software owe something to
the field. Over the years, \ODK has constantly used the JNCF as an
occasion to disseminate its achievements relevant to the community,
through contributed talks.

The last edition of JNCF to happen concomitantly with the project was
the occasion to double the dissemination efforts, and organize the
JNCF in tandem with the main \ODK conference ``Free Computational
Mathematics'': two invited talks at the ``Free Computational
Mathematics'' were delivered by noted members of the French Computer
Algebra community, while one of the invited lectures at JNCF was
delivered by Mohamed Barakat, a GAP developer and founder of the OSCAR
consortium, close to the \ODK community. A full session was devoted to
mathematical software on Monday afternoon, with two presentations
given by \ODK members, and one by a MapleSoft executive, which
triggered visible interest in the audience.

This was a successful workshop, and a great occasion to deliver 
successful developments of \ODK to its user base.
\end{event}
