\begin{sitedescription}{UB}
% PIC: 
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there

The Centre National de la Recherche Scientifique (National Center for
Scientific Research, CNRS) is a public organization under the responsibility of
the French Ministry of Education and Research. Its missions are
\begin{compactitem}
\item To evaluate and carry out research and
bringing social, cultural, and economic benefits for society.
\item To contribute to the application and promotion of research results.
\item To develop scientific information.
\item To support research training.
\item To participate in the analysis of the national and international
scientific climate in order to develop a
national policy.
\end{compactitem}
It consists of ten institutes in Biology, Humanities, Information Sciences and Technologies, etc. CNRS laboratories (or research units) are located throughout France, and employ a large body of tenured researchers, engineers, and support staff. 

Three research units are concerned by this proposal. The Laboratoire
Math\'ematiques d'Orsay (LMO) which is a joint unit with Universit\'e Paris-Sud
which is a partner of this proposal and has already been described
(see~\ref{desc:ParisSud}). Two laboratories in Bordeaux (FR), the Laboratoire
d'Informatique Bordelais (CNRS-LaBRI) and the Institut Math\'ematiques de
Bordeaux (CNRS-IMB). Those two units are joint with the Universit\'e de
Bordeaux which is a third party of this proposal
(see~\ref{section:ThirdParties}).

Bordeaux is an important center of studies and research in France
with approximately 50,000 students, 2,000 PhD students and 5,000
researchers. The University of Bordeaux was founded in the 15th
century and nowadays the city regroups two universities, dozens of
schools and 100 research laboratories with partner institution such
as CNRS, Inserm, INRA and INRIA.

The Institut Math\'emathiques de Bordeaux (CNRS-IMB) is a leading
institution in Number Theory. It is the home of the software
\PariGP and the Journal de Th\'eorie des Nombres de Bordeaux.

The city of Bordeaux also hosts two important young laboratories
for computer science: Laboratoire Bordelais d'Informatique (CNRS-LaBRI)
and INRIA-Bordeaux.

\medskip
In the context of this proposal, Bordeaux has a long standing experience in
Algorithmic Number Theory and two significant hardware infrastructures
(Plafrim and Avakas). The main task of Bordeaux in this project is the
extension of \PariGP in relation with the other software components and
the .

The Bordeaux site will lead the development of \PariGP within
workpackages \WPref{UI},\WPref{hpc}, \WPref{dissem}. It will also play
an important role in \WPref{hpc} with respect to combinatorics
and the organisation of workshop in favour of developing countries in \WPref{dissem}.

\subsubsection*{Curriculum vitae}
%Note that for purely administrative reasons, Loic Gouarin will be attached to
%CNRS but is naturally based at Paris sud - commented out, this is
%duplicated at Loic's CV (AK)

% Curriculum of the personnel at this institution

\input{CVs/Vincent.Delecroix.tex}
\input{CVs/Karim.Belabas.tex}
\input{CVs/Bill.Allombert.tex}
\input{CVs/Adrien.Boussicault.tex}
\input{CVs/Bordeaux-not-known.tex}
\begin{participant}[type=R,PM=6,gender=male,salary=5600]{Lo\"ic Gouarin}
Lo\"ic Gouarin is a CNRS employee but we refer to the Paris-Sud Universit\'e
description ({sec:ParisSud}) for his CV.
\end{participant}

\subsubsection*{Publications, products, achievements}

Some recent publications :
\begin{compactenum}
\item 
Karim Belabas, Eduardo Friedman,
\textit{Computing the residue of the Dedekind zeta function}.
Math. Comp. 84 (2015), no. 291, 357--369. 

\item
The PARI Group; \PariGP version 2.7.0, Bordeaux, 2014,
\url{http://pari.math.u-bordeaux.fr/}.

\item
Karim Belabas et al.
\textit{Explicit methods in number theory. Rational points and Diophantine equations},
179 pages, Panoramas et Synthèses 36, 179p., 2012.

\item
Bill Allombert, Yuri Bilu and Amalia Pizarro-Madariaga,
\textit{CM-Points on Straight Lines}, to appear in ``Analytic Number Theory'' (dedicated do H. Maier),
Springer.

\item
Vincent Delecroix,
\textit{Cardinality of Rauzy classes}
Ann. Inst. Fourier, 63 no 5 (2013), p. 1651-1715.

\item
Jean-Christophe Aval, Adrien Boussicault, Mathilde Bouvel, Matteo Silimbani
\textit{Combinatorics of non-ambiguous trees},
Advances in Applied Mathematics 56 (2014), p. 78-108.
\end{compactenum}

\subsubsection*{Previous projects or activities}

Current grants:
\begin{compactenum}
\item
 ANR PEACE (2012-2015).
    Goal: The discrete logarithm problem on algebraic curves is one of the rare
    contact points between deep theoretical questions in arithmetic geometry and
    every day applications. On the one side it involves a better understanding,
    from an effective point of view, of moduli space of curves, of abelian
    varieties, the maps that link these spaces and the objects they classify.
    On the other side, new and efficient algorithms to compute the discrete
    logarithm problem would have dramatic consequences on the security and
    efficiency of already deployed cryptographic devices. 

\item
ERC starting grant ANTICS (2011-2016).
    Goal: "Rebuild algorithmic number theory on the firm grounds of theoretical
    computer science".
    Challenges: complexity (how fast can an algorithm be?), reliability
    (how correct should an algorithm be?), parallelisation.
\end{compactenum}

\subsubsection*{Significant infrastructure}
\begin{compactenum}
\item The Plafrim is a regional federation hosted at INRIA Bordeaux (in partnership with the CNRS-LaBRI and CNRS-IMB). It has an important cluster devoted to experimental code (1188 cores).
\item The M\'esocenter de Calcul Intensif Aquitain (MCIA) is localised
in Bordeaux. It hosts the Avakas cluster (3328 cores,  38 TFlops) and the
M3PEC cluster (432 cores).
\end{compactenum}

\end{sitedescription}

%KEY-MORE-TODOS


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  sitedescription th eorie des nombres Plafrim mesocentre Avakas developped
%  LocalWords:  subsubsection Belabas Synthèses Allombert Bilu Pizarro-Madariaga Maier
%  LocalWords:  parallelisation TOWRITE
