\begin{sitedescription}{UJF}

% PIC: UJF == 999907429
% see: http://ec.europa.eu/research/participants/portal/desktop/en/organisation

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there

The Université Joseph Fourier (UJF)  is a research intensive university in an
international and high tech environment with $16\,600$ full time students
including $1\,200$ in a doctoral program, $1\,500$ lecturers and researchers and
$1\,500$ administrative and technical staff. Ranked 150th at the latest Shanga\"i
ranking, 5th french university, and top 75 natural sciences and mathematics.
UJF laboratories have successfully contributed to the FP7 european with 
105 projects, 26 of them being coordinated by UJF. Under H2020, UJF continues to
actively participate with already 4 projects.

The UJF partner gathers two teams from the Laboratoire Jean Kuntzmann
(CASYS team, with Jean-Guillaume Dumas) and the Laboratoire d’Informatique de
Grenoble (MOAIS, Clément Pernet). The 
CASYS team is specialised in Algebraic computations, cryptology, codes and hybrid symbolic-numeric dynamical systems. The MOAIS team is specialised in programming and scheduling
design on distributed resources for applications based on interactive simulation. The software
developed by this partner is significant. It includes the exact arithmetic
library \Givaro, the exact linear algebra libraries \fflas and \Linbox, and the
parallel programming runtime system \software{XKAAPI}.
Relying on this expertise, the UJF site will be coordinating the tasks related
to High Performance Mathematical Computing, forming Work Package~5:
\taskref{hpc}{hpc-pari}, 
\taskref{hpc}{hpc-gap}, 
\taskref{hpc}{hpc-linbox}, 
\taskref{hpc}{hpc-singular},
\taskref{hpc}{hpc-mpir}, 
\taskref{hpc}{hpc-combi},
\taskref{hpc}{pythran},
\taskref{hpc}{hpc-jupyter}.
It will also be in charge of the task~\taskref{hpc}{hpc-linbox} on the
development of high performance exact linear algebra with the \Linbox library.
\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

\input{CVs/Clement.Pernet.tex}
\input{CVs/Jean-Guillaume.Dumas.tex}
\input{CVs/Pierrick.Brunet.tex}
%\input{CVs/First.Last.tex}

\begin{participant}[type=res,PM=24,salary=4200,gender=]{Unknown}
Full time developer on \Linbox, and its integration into \Sage in the framework
of Task~\taskref{hpc}{hpc-linbox}. The recruited
person should master \software{C++} programming and library development in general. A good
knowledge on scientific computing and in particular linear algebra is also required. 
\end{participant}

\subsubsection*{Publications, products, achievements}

% \begin{compactenum}
% \item Coauthoring of the open source book ``Calcul Mathématique avec
%   Sage'', the first of its kind comprehensive introduction to
%   computational mathematics in Sage for education.
% \end{compactenum}
\begin{description}
\item[Software projects]\
  \begin{description}
 \item[\software{fflas-ffpack}:] An open-source \software{C++} library offering dense
    linear algebra kernels over a finite field. In the same  spirit as the
    numerical \software{BLAS} (Basic Linear Algebra Subroutines), and
    \software{LAPACK} libraries, it delivers high performance for the most
    commonly used routines of scientific computing: matrix multiplication,
    solving linear systems, computing echelon forms, determinants,
    characteristic polynomials, etc. This library has set the standard
    approach for high performance exact dense linear algebra. It is currently
    used in \Sage, and has inspired the design of similar routines in
    most commercial computer algebra systems: \Maple, \Magma, etc.
  \item[\Linbox:] An open-source \software{C++} middleware library for
    exact linear algebra. It uses \software{fflas-ffpack} for its dense finite
    field linear algebra part and extends its functionalities to other
    computation domains (integers, rationals, polynomial rings) and type
    matrices (sparse and structures matrices, black-box
    matrices). \Linbox is integrated in \Sage. 
  \item[\Pythran:] \label{pythran-descr}An open-source \Python-to-\texttt{C++} optimizing compiler
    offering an high performance runtime for Scientific Python kernels. Dynamicity of
    the \Python language is not compliant with static compilation. That's why only
    a subset of the \Python language is supported by \Pythran. Thanks to these
    restrictions, \Pythran generate code up to 3000 faster than original module.
    \Pythran takes advantage of multi-cores and SIMD instruction units and,
    thanks to type inference, it requires little annotations. Its runtime
    supports a growing subset of the \Numpy package.
  \end{description}
\item[Selected Publications]\ 
\medskip\noindent
\begin{compactenum}[1.]
\item Coauthoring of the open source book ``Calcul Mathématique avec
  Sage'', the first of its kind comprehensive introduction to
  computational mathematics in Sage for education.

\item Parallel computation of echelon forms (with J-G. Dumas, T. Gautier and Z. Sultan). 
\emph{In Proc. Euro-Par'14}  (2014),  LNCS 499--510. DOI: 10.1007/978-3-319-09873-9\_42.

\item Pythran: Enabling static optimization of scientific python programs
  (Serge Guelton, Pierrick Brunet, Alan Raynaud, Adrien Merlini, and Mehdi Amini.)
\emph{Proceedings of the \Python for Scientific Computing Conference (SciPy)} June 2013.

\item Fast Computation of Hermite Normal forms of random integer matrices (with
W. Stein).
\emph{J. of Number Theory} {\bf{130.7}} (2010), 1675--16833. DOI: 10.1016/j.jnt.2010.01.017

\item Dense Linear Algebra over Word-size Prime Fields (with J.-G. Dumas and P. Giorgi). 
\emph{Trans. on Math. Software} {\bf{35.3}} (2008), 1--42. DOI: 10.1145/1391989.1391992.
\end{compactenum}

\end{description}

\subsubsection*{Previous projects or activities}

\begin{compactenum}
\item Direction of the ANR program on High-Performance Algebraic
  Computations 2012-2015.
\item Participation to the NSF-Inria associate teams QOLAPS (with NCSU, USA)
\item Coordination of a CNRS PEPS grant (parallel computer algebra)
\item Organisation of the ISSAC'12 conference, the main
  international conference in computer algebra, and of PASCO'15, 
  the ISSAC'15 satellite conference on parallel computer algebra.
\end{compactenum}


\end{sitedescription}



\begin{draft}
\vspace{1cm}

\begin{verbatim}
- [X] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [X] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [X] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [X] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
  
- [X] Do all our tasks include us in the list of sites involved?
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS



%%% Local Variables: 
%%% TeX-master: "../proposal.tex"
%%% End: 

%  LocalWords:  sitedescription Laboratoire Kuntzmann Fousse d’Infor matique Pernet Roch
%  LocalWords:  Galet Taktuk ployment subsubsection Calcul Mathématique avec texttt emph
%  LocalWords:  fflas-ffpack texttt LinBox Pythran Dynamicity medskip noindent Guelton
%  LocalWords:  Pierrick Amini SciPy Hermite Storjohann NSF-Inria TOWRITE
