\begin{sitedescription}{UO}

% PIC: 999984350 THE CHANCELLOR, MASTERS AND SCHOLARS OF THE UNIVERSITY OF OXFORD OXFORD UK GB12550673

% See proposal.tex, Members of the Consortium for a complete description of what should go there

The University of Oxford is in the top ten universities worldwide in the Shanghai 2013 and 2014
rankings (the Universities of Cambridge and Oxford are the only non-US university there).
It employs over 10,000 staff and has a student population of over 21,000.
Most staff are directly appointed and managed by one of the University’s 130 departments or
other units within a highly devolved operational structure - this includes 5,900 ‘academic related’
staff (postgraduate research, computing, senior library, and administrative staff) and
2,820 ‘support’ staff (including clerical, library, technical, and manual staff). There are also
over 1,600 academic staff (professors, readers, lecturers), whose appointments are in the
main overseen by a combination of broader divisional and local faculty board/departmental
structures. 

The Department of Computer Science is one of the longest established
Computer Science departments in the country. Formerly known as the
University Computing Laboratory, it is home to a community of world-class research and
teaching. Research activities encompass core Computer Science, as well as Security,
Algorithms, computational biology, quantum computing, computational linguistics,
information systems, software verification and software engineering. 

The department is home to undergraduates, full-time and part-time Master's
students, and has a strong doctoral programme.  Students are highly skilled and
motivated, and as practice shows, easily start contributing to open-source
projects such as Sagemath.

\subsubsection*{Curriculum vitae}

\input{CVs/Ursula.Martin.tex}
\input{CVs/Edith.Elkind.tex}
\input{CVs/Dima.Pasechnik.tex}

\subsubsection*{Publications, products, achievements}
\begin{enumerate}
\item U. Martin. Computational logic and the social. J.Logic \& Computation (2014) [doi:10.1093/logcom/exu036]
\item U. Martin, A. Pease. Mathematical Practice, Crowdsourcing, and Social Machines, in Springer LNCS vol. 7961 (2013) [doi:10.1007/978-3-642-39320-4\_7]
\item G. Chalkiadakis, E. Elkind, M. Wooldridge. Computational Aspects of Cooperative Game Theory, Morgan \& Claypool, 2011 [doi:10.2200/S00355ED1V01Y201107AIM016]
\item E. Elkind, D. Pasechnik, Y. Zick. Dynamic weighted voting games, in Proc. AAMAS 2013 [http://dl.acm.org/citation.cfm?id=2484920.2485003]
\item S.H. Chan, H.D.L. Hollmann, D. Pasechnik. Sandpile groups of generalized de Bruijn and Kautz graphs and circulant matrices over finite fields,
J. Algebra 421(2015) [doi:10.1016/j.jalgebra.2014.08.029]
\end{enumerate}

\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item % {Projects and activities in Oxford} 
U. Martin holds an EPSRC Senior Fellowship, 2014-2017, to study the social machine of mathematics.
\item E. Elkind will hold an ERC Starter Grant, 2015-2020, to study and develop algorithms for collective decision
making on structured domains.
\item D. Pasechnik supervised  students contributing,  and significantly contributes himself, to Sagemath, to 
OEIS, and to GAP.
\end{enumerate}

\subsubsection*{Significant infrastructure}
Oxford has world-class computational facilities, including numerous HPC clusters and a dedicated
centre, Advanced Research Computing, to support HPC users.
Another dedicated facility, Oxford's e-Research Centre, facilitates 
digital research and drives innovation in technology. Last but not the least,
the library of Oxford University is one of the most complete libraries in the UK.
\end{sitedescription}




%KEY-MORE-TODOS


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  sitedescription programme Sagemath subsubsection Pease Chalkiadakis Zick
%  LocalWords:  Elkind Wooldridge Claypool Pasechnik Hollmann Bruijn Kautz TOWRITE
