\begin{sitedescription}{PS}

The Université Paris-Sud is among the 40 top universities worldwide in the
2013 Shanghai ranking, and is one of the two best French research
universities. With about 27000 students, 1800 permanent teaching staff
and 1300 permanent research scientists from national research
organisations (CNRS, Inserm, INRA, Inria), it is the largest campus in
France. Since 2006, scientists from the University were awarded two
Fields medals, one Nobel Prize and a number of other international
(European Inventor Award 2013, Wolf Prize 2010, Holweck Prize 2009,
Japan prize 2007) and national prizes.  The Université Paris-Sud has a
complete array of competences, ranging from the purest of exact
sciences to clinical practices in medicine, covering life and health
sciences, legal sciences and economics. Research at the Université
Paris-Sud, an essential part of academic understanding, is
complemented by research activities with a high valorisation
potential. Research contracts and partnership with companies make the
Université Paris-Sud a key actor and a major player in French
research.  The Université is located close to the Plateau de Saclay,
the largest cluster of public and private R\&D institutions in France
(with ca. 16000 research staff), and is one of the core members of the
University Paris Saclay – a world class university and a
world-renowned research and innovation hub.

In the context of this project, the Université Paris-Sud is the
home of one of the largest group of \Sage developers worldwide.
It's a member of the Open Source Thematic Group of the Systematic
Paris Region Systems and ICT Cluster. The University also hosts a
major research group (VALS) working on proof assistants (Coq) and
another on Human Centered Computing, which will facilitate reaching
toward those communities.

The Université Paris-Sud will lead the project (\WPref{management}),
host the majority of the man power for \WPref{component-architecture}
Component Architecture, and lead tasks on or around the \Sage
computational system (\taskref{component-architecture}{extract-smc},
\taskref{component-architecture}{interface-systems},
\taskref{UI}{ipython-kernels}, \taskref{UI}{sage-sphinx},
\taskref{UI}{dynamic-inspect}, \taskref{hpc}{hpc-combi},
\taskref{data-memo}). Finally, it will lead \WPref{dissem} and in
particular host or coorganize many of the community building,
training, and dissemination actions.

% The main participants have accumulated 15 years of experience of
% collaborative open source software development for mathematics
% leadership, and community animation.

\subsubsection*{Curriculum vitae of the investigators}

\input{CVs/Viviane.Pons.tex}
%\input{CVs/Nathann.Cohen}
\input{CVs/Florent.Hivert.tex}
\input{CVs/Nicolas.Thiery.tex}
\input{CVs/Samuel.Lelievre.tex}

\begin{participant}[type=res,PM=48,salary=5500]{NN}
\end{participant}
\begin{participant}[type=res,PM=36,salary=5500]{NN}
  We will hire two full time experienced software developers to work
  under the leadership of Nicolas Thiéry on the technical tasks
  pursued by Paris Sud, in particular in
  \WPref{component-architecture} Component Architecture and \WPref{UI}
  User Interface. When relevant, the mentoring will be complemented by
  Luca De Feo (\site{UV}), or experts of the tasks at hand from the
  larger community.
  %
  The fellow will have a strong software engineering experience,
  ideally in the Python ecosystem. We further require good
  communication and team working skills, in particular to work in
  tight collaboration with international open-source developer
  communities.
\end{participant}

%\begin{participant}[type=res,PM=36,salary=3100]{NN}
%  PhD.
%\end{participant}

\begin{participant}[type=res,PM=24,salary=3932]{NN}
  We will hire an experienced part time project manager to help with
  the overall management during the whole duration of \TheProject.
\end{participant}

% Participation:
% NT: Project Management: 4PM, Component Architecture: 4PM,
%     Training/Dissemination: 2PM, User Interfaces: 2PM
% Dev 1: Component Architecture: 48PM
% Dev 2: User Interface: 24PM, CA: 10PM, HPC: 2PM
% Florent: HPC: 4PM, User Interface: 2PM (Sphinx & co)
% Viviane: Training/Dissemination: 6PM
% Loic: Training/Dissemination: 3PM, User Interfaces: 2PM
% Sam: Training/Dissemination: 6PM
% Project manager: Project management: 24PM
% PHD: ???
% Total: 

\subsubsection*{Publications, achievements}

\begin{enumerate}
\item Lead of the \SageCombinat software project.
\item Coauthoring of the open source book ``Calcul Mathématique avec
  Sage'', the first of its kind comprehensive introduction to
  computational mathematics in \Sage for education.
\item 500 tickets contributed to \Sage.
\end{enumerate}


\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item Home of six Sage Days (week-long training and development workshops).
\item Coorganizer of six other Sage Days.
\item Founder and regular organizer of a bimonthly \Sage User Group
  meeting in the greater Paris area.
%\item Expertise exchanges with Logilab
%\item \TODO{XXX}
\end{enumerate}

\subsubsection*{Significant infrastructure}

The Université Paris Sud hosts the lead developers of the open source
cloud infrastructure \texttt{Stratuslab} and its reference
infrastructure (400 cores). The participants are regular users
of this infrastructure, and in close contact with the developers.

Paris Sud also hosts the WILDER platform, an experimental wall-sized
high-resolution interactive touch-screen for conducting research on
collaborative human-computer interaction and the visualization of
large datasets.

\end{sitedescription}



\begin{draft}
\vspace{1cm}\TOWRITE{VP}{Complete check list below -- delete completed items if you wish}

\begin{verbatim}
- [ ] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [X] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [ ] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [X] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
  
- [ ] Do all our tasks include us in the list of sites involved?
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: 

%  LocalWords:  sitedescription Paris-Sud organisations Inserm Inria Holweck valorisation
%  LocalWords:  Saclay subsubsection faut formel des projets antérieurs Acronyme titre
%  LocalWords:  agence financement durée Pareil les publi année SageCombinat Calcul avec
%  LocalWords:  Mathématique Logilab Sud texttt Stratuslab Chapuis
