\begin{sitedescription}{USH}

% Our person months should be 54 
% Dissemination: 24, HPC: 12, Social: 18

% PIC: 999976881
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there
The University of Sheffield is a leading Reasearch University in the
United Kingdom that was ranked 69th in the World in the most recent
2014 QS World University Rankings and was ranked in the top ten for
the most recent UK wide research asssment exercise.  Professor
Lawrence is based in The Sheffield Institute for Translational
Neuroscience (SITraN) and the Department of Computer Science at the
University of Sheffield have a unique partnership based on two shared
professorial appointments, Professor Winston Hide and Professor Neil
Lawrence.  SITraN is a world leading research centre for
neurodegenerative disease, located in a purpose built building on the
University of Sheffield campus adjacent to the Medical School at
Sheffield Teaching Hospitals NHS Trust Royal Hallamshire Hospital. 

The Department of Computer Science was ranked 5th across UK
departments in ``Research Quality'' by the recent UK-wide Research
Evaluation Framework. It has a particular history of working with data
with internationally leading groups in Machine Learning, Speech and
Language Processing.

The University of Sheffield will focus on tasks on HPC (\taskref{hpc}{hpc-jupyter}),
dissemination (\taskref{dissem}{project-intro}) and social aspects of the project
(\taskref{social-aspects}{social-output}). It will host one of the main project meetings
and run regular workshops on data science and Gaussian processes which incorporate
\TheProject outputs.

\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

\input{CVs/Neil.Lawrence.tex}
\input{CVs/Mike.Croucher.tex}
\input{CVs/Sheffield-not-known1.tex}
\input{CVs/Sheffield-not-known2.tex}
\begin{participant}[type=res,PM=1]{NN}
  We will support one month of an administrators time for assistance
  with workshop organisation and project administration across the
  whole duration of \TheProject.
\end{participant}


%\input{CVs/First.Last.tex}

\subsubsection*{Publications, products, achievements}

\begin{enumerate}
\item N. Fusi, C. Lippert, N. D. Lawrence and O. Stegle. (2014) "Warped linear mixed models for the genetic analysis of transformed phenotypes" in Nature Communications 5 (4890)
\item J. Hensman, M. Rattray and N. D. Lawrence. (2014) "Fast nonparametric clustering of structured time-series" in IEEE Transactions on Pattern Analysis and Machine Intelligence
\item M. A. \'Alvarez, D. Luengo and N. D. Lawrence. (2013) "Linear latent force models using Gaussian processes" in IEEE Transactions on Pattern Analysis and Machine Intelligence 35 (11), pp 2693--2705
\item N. Fusi, O. Stegle and N. D. Lawrence. (2012) "Joint modelling of confounding factors and prominent genetic regulators provides increased accuracy in genetical genomics studies" in PLoS Computat Biol 8, pp e1002330
\end{enumerate}

\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item Organisers of the Gaussian Process Summer Schools (three 3-day workshops on Gaussian process models in python and the IPython notebook).
\item Organisers of five Gaussian Process and Data Science Road Shows (educating on data science and Gaussian process models in Uganda, Colombia, Italy, Australia and Kenya) Each workshop is 3-4 days long. 
\end{enumerate}

\subsubsection*{Significant infrastructure}
\begin{enumerate}
\item The Sheffield Institute for Translational Neuroscience is a 18 million pound world leading institute for research into neurodegenerative disorders. It houses clinicians, biologists and computationalists under a single roof and contains the Sheffield Microarray and Next Generation sequencing Core Facility. The institute provides an exemplar of how mathematical ideas can be rapidly translated to analysis through provision of appropriate software frameworks.
\item The Sheffield group are regular contributors to open source software including a python framework for Gaussian process modelling (\url{https://github.com/SheffieldML/GPy}), contributions to the Bioconductor framework for computational biology (puma, tigre, gprege), and more recently frameworks for open data science (\url{https://github.com/sods/ods}).  
\item The Sheffield group has taken a leading role in education in data science and machine learning with a series of workshops and summer schools which use \Jupyter as the main interface for practical implementation of ideas (\url{http://ml.dcs.shef.ac.uk/gpss/}).

\end{enumerate}
\end{sitedescription}


\begin{draft}
\vspace{1cm}
\begin{verbatim}
- [X] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [X] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [X] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [X] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
  
- [X] Do all our tasks include us in the list of sites involved?
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS
