\begin{sitedescription}{USO}

% PIC: 
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there


  The University of Southampton (USO) is one of the leading
  universities in the United Kingdom, was founded in 1952 and is a
  member of prestigious Russell Group of UK Universities. The
  University of Southampton has more than 19,000 undergraduate
  students and 4,000 postgraduates and is an excellent venue for
  conducting cutting-edge research and for providing high quality
  education. The university is truly international, drawing students
  from over 130 different countries and benefiting from a wide and
  varied culture. It is ranked in the top 1\% of universities
  worldwide (QS world university rankings 2014-15) and in the top 15
  of research led universities in the UK, and is participating in a
  high number of collaborative research projects and related
  initiatives. To ensure the impact of its research projects,
  University of Southampton’s Research \& Innovation Services (R\&IS)
  is responsible for professional protection of IP and supporting
  commercial development with industry. R\&IS has had considerable
  success, licensing annual revenue in excess of \EUR{1}million and
  launching twelve successful spin-out companies since 2000.  The
  university has a strong track record of working in European
  projects, especially within the Framework Programme. The EC 6th FP7
  Monitoring Report ranked UoS 17th out of all higher and secondary
  education organisations for number of FP7 participations during
  2007-2012. Throughout the FP7 UoS has received \EUR{132M} in
  research grants and has been involved in 319 projects, including 63
  ICT and 8 INFRASTRUCTURES Collaborative Projects. 

  The Faculty of Engineering and the Environment (FEE) consists of 370
  research postgraduate students and 340 academic and research
  staff, and is one of the lead engineering faculties in Europe, educating
  a range of professionals and generating research of the highest
  quality. Southampton's world-leading engineering ranking is
  confirmed by being ranked first in the UK for the volume and quality
  of the research in Electronic Engineering, Electrical Engineering
  and General Engineering in the latest Research Excellence Framework
  (REF) 2014. The faculty also hosts the University Technology Centres
  and Research Framework Agreements with key partners including:
  Airbus, Rolls-Royce, Lloyd’s Register, Microsoft and Network
  Rail. FEE has a strong background in working on international
  research projects, including 84 EU FP7 projects worth over \euro
  28M.


\medskip In the context of this proposal, Southampton has long
standing experience in high performance computer simulation to advance
science and engineering, with significant hardware infrastructure, a
critical mass of several hundred researchers working in the field, and
a dedicated doctoral training center in computational modelling.
Southampton's main tasks in this project are the extension and application of the
\Jupyter{} Notebook technology in computational materials research to
provide a virtual research environment demonstrator for a large research community.

\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

\input{CVs/Hans.Fangohr.tex}
\input{CVs/Ian.Hawke.tex}
\input{CVs/Southampton-not-known.tex}
%\input{CVs/First.Last.tex}

\subsubsection*{Publications, products, achievements}

\begin{enumerate}
\item Open Source micromagnetic simulation framework Nmag,
  \href{http://nmag.soton.ac.uk}{http://nmag.soton.ac.uk}, Thomas
  Fischbacher, Matteo Franchin, Giuliano Bordignon, Hans Fangohr: \emph{
A Systematic Approach to Multiphysics Extensions of Finite-Element-Based Micromagnetic Simulations: Nmag 
IEEE Transactions on Magnetics \textbf{43}, 6, 2896-2898 (2007)}
\item Other open source contributions to the micromagnetic simulation
  community: OVF2VTK, higher order anisotropy extensions to OOMMF,
  OVF2MFM, summarised at
  \href{http://www.southampton.ac.uk/~fangohr/software/index.html}{http://www.southampton.ac.uk/~fangohr/software/index.html} 
\item H. Fangohr.
\emph{A Comparison of C, Matlab and Python as Teaching Languages in Engineering}
Lecture Notes on Computational Science \textbf{3039}, 1210-1217 (2004)
\end{enumerate}

\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item EPSRC Doctoral Training Centre in Complex Systems Simulations
  (\href{http://icss.soton.ac.uk}{http://icss.soton.ac.uk}), jointly
funded by EPSRC and the University of Southampton, \EUR{14m}, (2009--2018)
\end{enumerate}

\subsubsection*{Significant infrastructure}
\begin{enumerate}
\item The University of Southampton hosts the largest university owned
  Supercomputer "Iridis 4" in the UK (12,300 cores, 250 TFlops),
  the hardware (\EUR{3.75}m) is refreshed every 3 years.
\item A community of 200 academics and over 500 researchers and
  doctoral students are users of this facility and provide a wide
  network pushing forward excellent computational science in the
  context of solving real world problems.
\item EPSRC Centre for Doctoral Training in Computational Modelling in
  the United Kingdom,
  \EUR{12}m. (\href{http://ngcm.soton.ac.uk}{http://ngcm.soton.ac.uk}),
  (2013--2022)
\end{enumerate}
\end{sitedescription}



\begin{draft}
\vspace{1cm}\TOWRITE{HF}{Complete check list below -- delete completed items if you wish}

\begin{verbatim}
- [X] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [X] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [X] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [X] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  sitedescription Programme organisations programmes Centres subsubsection
%  LocalWords:  micromagnetic Nmag Fischbacher Franchin Bordignon Fangohr emph textbf
%  LocalWords:  Multiphysics summarised Iridis TFlops Modelling
