\begin{sitedescription}{SA}

% PIC: 
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga
% 

% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there

The Centre for Interdisciplinary Research in Computational Algebra (CIRCA)
fosters research at the interface of Mathematics and Computer Science including
abstract and algorithmic algebra and combinatorics, formal languages and
automata, mathematical software and constraint programming. Our success is 
founded on the close integration of theoretical and algorithmic research and 
the development and use of state-of-the-art software.

In 1997, CIRCA became the centre of the development of \GAP after the 
retirement of under~Prof.~J. Neub\"user who initiated the system in 
mid-1980s in Aachen. This move was supported by EPSRC, EU and Leverhulme 
grants. Now GAP and its packages comprise over 1 million lines of C and 
1.25 million lines of GAP code. The system is distributed freely under the GPL2,
and our records show that it has been installed in at least 3000 sites and 
cited in several thousands publications 
(see \url{http://bit.ly/gap_citations} and \url{http://www.gap-system.org/Doc/Bib/bib.html} ). 
GAP was used in landmark computations such as the ``Millennium Project'' 
to classify all groups of order up to 2000 
and the classification of the over $10^{19}$ semigroups of order 10. It is 
designed to be natural to use for mathematicians; to be powerful and flexible 
for experts and to be freely extensible so that it can encompass new mathematics. 
These objectives have been met and GAP was awarded the ACM/SIGSAM Richard D. Jenks
Memorial Prize for Excellence in Software Engineering applied to Computer Algebra in 2008.

Nowadays, as one of the centres of \GAP\ development, CIRCA has
excellent contacts with developers and users worldwide. Particularly 
relevant to this proposal are the Singular and homalg groups at 
Kaiserslautern, Soicher at QMUL, Praeger and others at UWA, Cooperman 
and his students at NEU and the other GAP centres in Aachen, Fort Collins 
and Braunschweig.

CIRCA's output includes first class results in Pure Mathematics and in
Computer Science, recognised by our highly-cited publications in top
international venues in both disciplines and our widely used research
software. Beyond the individual international connections of the 
investigators and research staff, CIRCA as a centre has national 
importance and international standing. CIRCA has been selected to host 
major conferences such as CP 2010, BCC 2009, PP 2007, and the 
``Groups St Andrews'' series in 2005 and 2013.


\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

\input{CVs/Steve.Linton.tex}
\input{CVs/Alexander.Konovalov.tex}
\input{CVs/Markus.Pfeiffer.tex}
%\input{CVs/First.Last.tex}

% TODO(AK): select and add more items below

\subsubsection*{Publications, products, achievements}

\begin{enumerate}
\item 
S.~Linton, K.~Hammond, A.~Konovalov, C.~Brown, P.W.~Trinder, H.-W.~Loidl, 
P.~Horn and D.~Roozemond, Easy Composition of Symbolic Computation Software using 
SCSCP: A New Lingua Franca for Symbolic Computation.
J. Symbolic Computation, 49 (2013), 95--119.
% http://dx.doi.org/10.1016/j.jsc.2011.12.019
\item
V.~Janjic, C.M.~Brown, M.~Neunh{\"o}ffer, K.~Hammond, S.~Linton, H-W.~Loidl. 
Space exploration using parallel orbits: a study in parallel symbolic computing.
in Parallel Computing: Accelerating Computational Science and Engineering (CSE). 
vol. 25, Advances in Parallel Computing, IOS Press, 2013, p. 225--232.
\item 
A.~Konovalov and S.~Linton. 
SCSCP --- Symbolic Computation Software Composability Protocol, 
Version 2.1.4; 2013. GAP package (\url{http://alexk.host.cs.st-andrews.ac.uk/scscp/}).
\item
V.~Slavici, D.~Kunkle, G.~Cooperman, S.~Linton. 
An efficient programming model for memory-intensive recursive 
algorithms using parallel disks. In ISSAC 2012: Proceedings of 
the 37th International Symposium on Symbolic and Algebraic Computation. 
New York, ACM Press, 2012, p. 327--334.
\item 
R.~Behrends, A.~Konovalov, S.~Linton, F.~L{\"u}beck, M.~Neunh{\"o}ffer.
Towards high-performance computational algebra with GAP.
Proceedings of the Third International Congress on Mathematical Software: 
Kobe, Japan, September 13-17, 2010. LNCS 6327, Springer, 2010, p. 58--61.
% http://dx.doi.org/10.1007/978-3-642-15582-6_12 

\end{enumerate}

\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item
\textbf{Multidisciplinary Critical Mass in Computational
Algebra and Applications} EP/C523229 2005--2010, \pounds 1.1m.
Through a range of subprojects this grant developed CIRCA as 
a broad centre of excellence in computational algebra. We 
extended \GAP, developed new algorithms, and opened up new 
applications in AI, combinatorics and physics. The project 
produced over 200 refereed publications in a huge range of 
venues, and delivered, as intended, a sustainable step change 
in the scale and breadth of CIRCA's research.
\item 
\textbf{SCIEnce: Symbolic Computation Infrastructure for Europe}
(FP6 eRII3-CT-026133) 2006--2011, \euro 3.2m, coordinator.
This project tackled the fragmentation of the European community 
of researchers in, and users of, symbolic computation. Among the 
nine partners were four major system developers (of \GAP, \Maple, 
\MuPAD and \KANT), an international research institute (RISC-Linz) 
and other groups with specialist expertise. Project activities 
included symbolic web services, system composability, symbolic 
grid and cloud computing and a program of visits, workshops and 
summer schools. One important outcome was a new protocol \SCSCP, 
now used well beyond the original project.
\item
\textbf{HPCGAP: High Performance Computational Algebra and Discrete Mathematics} 
EP/G055181 2009--2014, \pounds 1.5m, four sites. This project aimed
at reengineering GAP to support simple, safe and efficient parallel 
programming on a range of systems from multicore laptops and desktops, 
through departmental and university clusters to HPC systems. By the 
end of the project, we had adapted a complex system including a 
language runtime and over 400 000 lines of interpreted code to enable 
safe and efficient parallel programs. The proposed project is very timely as
the multi-threaded version of GAP is becoming mainstream, and users and 
package developers need training and support to parallelise their code.
\end{enumerate}

\subsubsection*{Significant infrastructure}

CIRCA provides hosting for the GAP website and ftp-servers, runs the 
automated system to check, fetch and test updates of GAP packages, uses 
the Jenkins continuous integration system to run regression tests for the 
development version and release candidates on its computer infrastructure, 
and manages GAP releases preparation.
\end{sitedescription}


\begin{draft}
  \vspace{1cm}\TOWRITE{SL}{Complete check list below -- delete completed items if you wish}

\begin{verbatim}
- [X] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [X] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [ ] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

     SL: all done excepot possibly for Social Aspects


- [X] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [ ] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.
  
- [X] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
  
- [ ] Do all our tasks include us in the list of sites involved?
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  sitedescription Neub centres recognised subsubsection Konovalov Roozemond
%  LocalWords:  textbf
