\begin{sitedescription}{UW}
% PIC: 999976784
% see: http://ec.europa.eu/research/participants/portal/desktop/en/orga

The Mathematics Institute at the University of Warwick was ranked 23rd
worldwide in the 2013 QS world university subject rankings, and third in
the UK in the 2014 Research Excellence assessment.  Five
members of the Department are Fellows of the Royal Society, and one,
Regius Professor Martin Hairer, was awarded a Fields Medal in 2014.
Mathematics and Statistics at Warwick currently hold £35.8M in
research grants from EPSRC (the next highest in the UK being Cambridge
at £22.8M and Oxford at £24.2M).  Nine members of the department
currently hold ERC grants.


% See ../proposal.tex, section Members of the Consortium for a
% complete description of what should go there

\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution

\input{CVs/John.Cremona}

%\input{CVs/First.Last.tex}
%\input{CVs/First.Last.tex}
%\input{CVs/First.Last.tex}

\subsubsection*{Publications, products, achievements}
\begin{enumerate}
\item
The Number Theory research group at Warwick was started only in 2006,
but has rapidly risen to international status and one of the largest
and most vibrant groups in Europe, comprising 25 members (professors,
lecturers, postdoctoral researchers and early stage researchers).  Of
the group's members, two (Loeffler and Dokchitser) hold Royal Society
Research Fellowships and one (Bartel) a Royal Commission 1851
Fellowship.  Loeffler won a Leverhulme Foundation Prize jointly with
Zerbes.
\item
Several members of the Number Theory group at Warwick are \Sage
developers, including John Cremona, who has contributed thousands of
lines of code to \Sage\ since 2006 both through his {\tt eclib} C++
library and through original \Python code which forms part of the
\Sage\ library; David Loeffler, who has contributed substantially to
the modular forms module in Sage; and postdoc Marc Masdeu, who has
worked on the \Sage-Flint interface.
\end{enumerate}

\subsubsection*{Previous projects or activities}

\begin{enumerate}
\item
In 2013 Professors John Cremona and Samir Siksek, together with
co-investigators at Bristol, were awarded a six-year major grant of
£2.2M from the UK Engineering and Physical Sciences Research Council
(EPSRC) to support the L-functions and Modular Forms Database (LMFDB)
project.  This grant funds three postdoctoral researchers at Warwick,
computer equipment to host its database and website, and regular LMFDB
workshops.
\item
Each year Warwick hosts a year-long Warwick EPSRC Symposium focussing
on one area of mathematical research.  The 2012-13 Number Theory
Symposium included six research workshops and a summer school ``Number
Theory for Cryptography'' and raised the international profile of the
number theory group substantially.
\end{enumerate}

\subsubsection*{Significant infrastructure}

Computing infrastructure available to the group is excellent, with
seven dedicated machines (over 300 cores) as well as access through
Warwick's Centre for Scientific Computing which hosts a 6000-core
linux cluster and a 3500-core cluster of workstations.
\end{sitedescription}



\begin{draft}
\vspace{1cm}\TOWRITE{JEC}{Complete check list below -- delete completed items if you wish}

\begin{verbatim}
- [ ] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
  Take into account person months associated with work package 1, time
  of all staff to be hired and work on the project (including
  investigators). Figure 5 helps with a quick check of the sums over
  different work packages.

- [ ] completed site specific resource summary in resources.tex,
  including table of non-staff costs. This is compulsory (EU
  regulations) if the non-staff cost exceed 15% of the total cost, and
  is likely to be the case for most of the partners. We ask everybody
  to do it, to be consistent and show transparently how we have
  planned our total budget.

- [ ] Have all our tasks a designated lead institution? Check in the
  Work Packages that all the tasks you are involved in have a
  dedicated lead party. If the lead party is "USO", then use:
  \begin{task}[lead=USO]

- [ ] Have all our deliverables a designated lead institution [using
  the 'lead=' key]?

- [ ] In the "Members of the consortium section", have we addressed "a
  description of the legal entity and its main tasks, with an
  explanation of how its profile matches the tasks in the
  proposal"? See Entry for Paris Sud and Southampton as examples.

- [ ] In the Members of the consortium section, have we given
  descriptions of all the people we intend to hire (even if we don't
  know who that is yet). 
\end{verbatim}
\end{draft}

%KEY-MORE-TODOS



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:


%  LocalWords:  sitedescription Regius Hairer subsubsection Loeffler Dokchitser Zerbes tt
%  LocalWords:  eclib Masdeu Siksek
