
\section{KPIs}

\begin{introduction}
 The following Key Performance Indicators (KPI) show how OpenDreamKit addresses the specific impacts listed in the work programme. KPIs were 
 thought through by the members of OpenDreamKit so that they are meaningful, reusable, realistic and easily measurable. the following 
 qualitative and quantitative indicators are divided into the four aims of OpenDreamKit. If quantitative indicators are more useful for 
 reporting and internal evaluation, qualitative indicators will give content for futher dissemination and communication purposes, for 
 example through the project website.
\end{introduction}


\subsubsection{KPIs for Aim 1}

\begin{Aim 1}
  Improve the productivity of researchers in pure mathematics and applications by promoting collaborations based on mathematical
  software, data, and knowledge.\end{recommendation}


Qualitative metrics: success stories, reported as blog posts.

             success stories reported as blogposts

           %Statistiques sur le site web Open dream kit b de blogposts:
          



Story of cocalc adoption for teaching in UK after ODK's Kickoff
            
Support was given to a number of lecturers in Sheffield to migrate to Jupyter and CoCalc (formerly SageMathCloud) but also to those that had 
already been using CoCalc and Jupyter notebooks for their courses. These included lecturers from Computer Science, Physics, Biomedical 
Science, Bioinformatics, and Materials Science. (D2.17, T2.6) A previously generated CoCalc tutorial was extended by adding tutorial 
sections for students having courses in CoCalc as well as with a hands-on tutorial for lecturers to get started. The material can be found 
as a website at https://tutorial.cocalc.com/ 

%ttps://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/7cbee70c3bff6d0778733be931c927981df87e65/meetings/2018-10-28-Luxembourg/SteeringCommitteeMeeting/ProgressReports/Sheffield-Leeds.md


Reproducible GAP experiments on Binder (AlexK)
AlexK %lien avec ODK Qui est-ce?
used GAP Jupyter interface in teaching at PGTC 2018 and "Software tools for mathematics" workshop.
Here https://github.com/alex-konovalov/gap-teaching, you can find a collection of GAP Jupyter notebooks. It uses the Docker container with 
the latest public release of GAP, which is maintained in a separate repository at https://github.com/gap-system/gap-docker.

             Stories of Mike (Michael Croucher)
Michael is a Research Software Engineer at the unievrsity pf SHeffield, passionate about improving the quality of research
software. He enables researchers to ask larger and more complex research questions by improving the software they develop. Along with the 
Software Sustainability Institute, the UK Research Software Engineering Association and the EU-funded OpenDreamKit project, Michael Croucher 
actively campaign to improve the career prospects of the talented people who underpin a huge amount of computational research…. He explained 
in a blogpost on october 2018 " How OpenDreamKit supported the RSE revolution" saying that "By 2015, there were a small number of central 
‘Research Software Engineering Groups’ within UK Universities with his group at Sheffield  being among the first. OpenDreamKit was one of 
the first projects they won that demonstrated that funders would support RSEs on major grants – this improved credibility of the new role a 
great deal and helped secure its future at Sheffield."% dernière phrase incomrpéhensible pour moi
 https://opendreamkit.org/2018/10/29/ODK-RSE/
                              

\subsubsection{KPIs for Aim 2 / adoption of \ODK's technologies}

\begin{Aim 2}
  Make it easy for teams of researchers of any size to set up custom, collaborative Virtual Research Environments tailored to their
  specific needs, resources and workflows. The VRE should support the entire life-cycle of computational work in mathematical research, from 
  initial exploration to publication teaching and outreach;
\end{recommendation}

Qualitative metrics:
\begin{itemize}
\item Success stories about \ODK based VRE deployments (university-wide, Mathrice, EGI, Micromagnetics VRE, ...), 
and generally speaking adoption of \ODK's components.
\item added metrics:
 
 
       June 2018, Jupyter was awarded prestigious 2017 ACM Software System Award
       Previous winners include: UNIX, TCP/IP, the Web, TeX, Java, GCC, LLVM
       %ttps://blog.jupyter.org/jupyter-receives-the-acm-software-system-award-d433b0dfe3a2
              
       Partnership with EGI (main stakeholder of the EOSC): there is an ongoing collaboration between EGI and OpenDreamKit to deploy
       JupyterHub and BinderHub-based EGI services. Proofs of concepts for both have been deployed by Enol Fernandez from EGI. Both parties
       are very satisfied with the collaboration and want to strengthen it. A Technology Provider Agreement was signed between EGI and
       Simula, on behalfF of OpenDreamKit/Jupyter developers. Joint applications to upcoming EOSC calls, and in particular INFRAEOSC-02-2019
       https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/7cbee70c3bff6d0778733be931c927981df87e65/meetings/2018-10-28-Luxembourg
       /SteeringCommitteeMeeting/ProgressReports/ParisSud.md
       %ttps://github.com/OpenDreamKit/OpenDreamKit/blob/88ee2544e31e661c4eac7f39fa7333f24b62ab2a/WP4/D4.8/report.tex
              
       %2-3 blog sur comment deployer VRE (luca defeo, loic gouarin) deploying jupyter hub:
       Deploying a containerized JupyterHub server with Docker
       %https://opendreamkit.org/2018/10/17/jupyterhub-docker/
       Toward versatile JupyterHub deployments, with the Binder and JupyterHub convergence
       %https://opendreamkit.org/2018/03/15/jupyterhub-binder-convergence/
       Loic gouarin: Deploying JupyterHub with Kubernetes on OpenStack
       %https://blog.jupyter.org/how-to-deploy-jupyterhub-with-kubernetes-on-openstack-f8f6120d4b1
     
       Use of Jupyter components by Logilab
       Simulagora
       Logilab VRE deployment for application development and deployment.
       Can use JupyterLab for application development, which can then be deployed with a simplified parameters form input. % ils utilisent des composants de Jupyter de plus en plus

         
  Quantitative metrics:
\begin{description}
\item List of known \ODK based VRE deployments, as tracked on \url{https://github.com/OpenDreamKit/OpenDreamKit/issues/174}
\item Number of installs of \ODK's components via platform-specific distribution channels: Debian popcon, Arch statistics, installer
  downloads, \dots
%\item[WP3] Decrease in set-up time for dissemination workshops based
%  on \ODK components. This is hard to measure objectively, better to
%  gather user stories.
\end{description}

A prototype VRE shall be extended with improved ease of deployment, new functionality such as interactive 3D visualization and real-time 
collaboration, enabling researchers to collaborate productively in a shared computational environment. Integrating notebooks and semantic 
knowledge into a publication / knowledge system enable a continuous process of leveraging OpenDreamKit components from research to 
publication.
A group of mathematical researchers with access to common computational resources, such as a shared lab computer or cloud servers,has 
already been able to deploy a prototype VRE with JUPYTERHUB, integrating OpenDreamKit components. The Jupyter kernels for mathematical 
software developed as part of OpenDreamKit make computational mathematical components accessible in a JUPYTER environment, enabling a 
Jupyter-based deployment of the relevant tools for the researchers. The process of working on notebooks is greatly improved by review tools 
developed as part of WP4, enabling researchers to collaborate to some degree in a shared computational environment.

% à formuler https://opendreamkit.org/meetings/2018-10-28-Luxembourg/ProjectReview/wp4.pdf
%https://github.com/OpenDreamKit/OpenDreamKit/issues/174
Millions of notebooks are already online (over 3M on GitHub alone)
OpenDreamKit kernels
The Jupyter kernels for mathematical software  developed as part of OpenDreamKit make computational mathematical components accessible in a 
Jupyter environment, enabling a Jupyter-based deployment of the relevant tools for the researchers.

Now 117 Jupyter kernels (49 when ODK started), 6 contributed to by ODK.
Further improved kernels from first reporting period (GAP, PARI, Singular). 
Delivered as D4.7, and MMT as part of WP6
Also contributed to kernels: cling (C++), SageMath

KPI: ODK Kernels on GitHub
Notebooks found on GitHub using each kernel (code):
SageMath: 6199
Xeus-cling: 684
GAP: 63
Singular: 8
PARI/GP: 3
MMT: 1

We have made collaborations with various institutions and projects to deploy instances of JupyterHub and CoCalc (formerly 
SageMathCloud),Collecting estimates of the difficulties involved in such deployments in various use cases will help us to plan future 
deployments and seek what could be done to ease deployment.
    
9 Jupyter deployments:

  Running:
  [x] Local CoCalc instance at Universität Zürich.
      Deployed September 2015 - February 2016.
      People involved: @pdehaye, @williamstein
      
  [x] [Instance of JupyterHub](https://jupyter.math.cnrs.fr/hub/) deployed by the [Mathrice group](http://mathrice.fr/)
      Host Infrastructure: France Grille's LAL cloud
      Users: members of math labs in France
      Main use case: casual use 
      People involved: the Mathrice group, @nthiery

  [x] Local [JupyterHub instance at Université Paris Sud / Paris Saclay](http://jupytercloud.lal.in2p3.fr/)
      Host Infrastructure: France Grille's LAL cloud
      Users: personnel and students of UPSud / Paris Saclay
      Main use case: use in classroom (Python, Sage, C++), casual use
      People involved: @VivianePons, @nthiery, @gouarin 
      
  [x] JupyterHub instance deployed on USheffield's HPC system
      http://docs.iceberg.shef.ac.uk/en/latest/using-iceberg/accessing/jupyterhub.html
      People involved: @mikecroucher
  
  [x] JupyterHub instance(s) deployed at UVSQ
      https://opendreamkit.org/2018/10/17/jupyterhub-docker/
      Main use case: use in classroom (Sage, Python, C, Apache Spark), casual use
     People involved: @defeo 
   
  [x] [Gallery of JupyterHub instances](https://jupyterhub.readthedocs.io/en/latest/gallery-jhub-deployments.html)
  [x] JupyterHub and Binder instances deployed on EGI infrastructure; see #205.
  [x] Easy deployment of live GAP/SageMath/... notebooks with [mybinder](mybinder.org), 
  thanks to the Docker containers (#58);potential alternatives: [Debian packaging](https://wiki.debian.org/DebianScience/Sage) 
  and [Conda packaging](https://wiki.sagemath.org/Conda).
  People involved: @nthiery, ...
  [x] Local instance of CoCalc (using the Docker container) at the University of Gent
    Main use case: teaching for mathematics students
    Deployed at jupyter.mathhub.info With MMT kernel People involved: @tkw1536

  
  3 interests in Jupyter deployments:
  [ ] JupyterHub deployment at the [Einstein Institute of Mathematics](http://math.huji.ac.il), part of the Hebrew University of Jerusalem
      Related: [Sage Days 79](https://wiki.sagemath.org/days79)
  [ ] Local Cocalc instance at UPSud
      Host infrastructure: UPSud's cloud.
      Users: personnel and students of UPSud
      Main use case: use in classroom (Python, Sage, C++), casual use
      People involved: @VivianePons, @nthiery, @gouarin 
      For now, priority has been given to the above JupyterHub instance
  [ ] Integration of Sage in the tmpnb.org's temporary notebook server
      People involved: @rgbkrk, @nthiery
      Status: some experiments run during a sprint at Pycon'15. Now that Sage's
      Jupyter kernel is well integrated in the stable version of Sage, it's just a question
      of installing Sage in tmpnb's docker container.
      This is superseded by [binder](http://mybinder.org); see above.
\end{itemize}

Use of ODK's computational components in collaborative workspaces
“ODK’s computational components available on major platforms” (month 42).
       User story: users shall be able to easily install ODK’s computational components on the three major platforms (Windows, Mac, Linux)
       via their standard distribution channels.
       
       There are Packages available for all components on Debian, Ubuntu, Fedora, Arch, Gentoo, ...Experimental Conda Forge packages  % à voir
      Arch Based upon voluntary reports from ~30k users %graphique à rajouter
      Debian Based on ~200K voluntary submissions %image à rajouter
      Ubuntu, based on ~2.8M voluntary submissions %image à rajouter
      
      Docker hub 
          10k pulls for sagemath-jupyter
          5.7k pulls for sagemath
          3.8K pulls for gap-docker

      
\subsubsection{KPIs for Aim 3}

\begin{recommendation}{Aim 3}
  Identify and promote best practices in computational mathematical research including: making results easily reproducible; producing
  reusable and easily accessible software; sharing data in a semantically sound way; exploiting and supporting the growing
  ecosystem of computational tools.
\end{recommendation}

Qualitative metrics: Success stories
\begin{itemize}
\item Mike Croucher's talk ``Is your research software correct''
Best practice and tools for correct and reproducible research]
        
        The excellent talk \href{https://mikecroucher.github.io/MLPM_talk/}{``Is your research software correct''} by Mike 
         Croucher, highlights crucial best practice whenever software is used in research, including open code and data sharing, 
         automation, use of high level languages, software training, version control, pair programming, literate computing, or testing. A 
         lot of the work in ODK relates to disseminating this set of best practice (\longWPref{dissem}), and enabling it through appropriate 
         technology (\longWPref{UI}).  Just to cite a few examples, \longdelivref{UI}{jupyter-collab}, and \longdelivref{UI}{jupyter-test} 
         enable respectively version control and testing in the \Jupyter literate computing technology, while Mike's talk is and will be 
         delivered in several of ODK's many training events (see \longdelivref{dissem}{workshops-1} for the list of training events in year 
         1).
         
         OpenDreamKit member, Tania Allard, ran a hands-on workshop on Jupyter notebooks for reproducible research. This workshop focused on 
         the use of Jupyter notebooks as a means to disseminate reproducible analysis workflows and how this can be leveraged using tools 
         such as nbdime and nbval. Both nbdime and nbval were developed by members of the OpenDreamKit project as a response to the growing 
         popularity of the Jupyter notebooks and the lack of native integration between these technologies and existing version control and 
         validation/testing tools.
         An exceptional win was that this workshop was, in fact, one of the most popular events of the conference and we were asked to run 
         it twice as it was massively oversubscribed. This reflects, on one hand, the popularity of Jupyter notebooks due to the boom of 
         literate programming and its focus on human-readable code. Allowing researchers to share their findings and the code they used 
         along the way in a compelling narrative. On the other hand, it demonstrates the importance of reproducible science and the need for 
         tools that help RSE and researchers to achieve this goal, which aligns perfectly with the goals of OpenDreamKit.
         https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/6faf6eb2f1532f342f86c8da633078067ca40c85/_posts/2018-03-07-
         opendreamkit-at-the-rse-conference.md

 
\item Some research paper that showcases a range of best practices supported by ODK work (paper written collaboratively on e.g. github,
  software distributed as e.g. SageMath package, live demo and logbooks on binder, nbdime for collaboration, ...).\item ...\end{itemize}

%https://opendreamkit.org/meetings/2018-10-28-Luxembourg/ProjectReview/WP6.pdf
%envoyer mail à Michael Kohlhase
Multi-Site involvement of Researchers (Mobility of Researchers)
PD. Dr. Florian Rabe (Joint appointment UPSud/FAU)
Felix Schmoll Summer Internship (From JacU to St.Andrews)
Prof. Nathan Carter (Bentley Univ.) in St. Andrews (Sabbatical)
Heavy interest by the theorem proving community about MitM Ontology
Logipedia (http://logipedia.science) adopts the MitM principle of integrating (logical) systems by aligning concepts.
First ODK-external MitM “user” for the next months: Andrea Thevis, Saarbrücken 


The Math-in-the-Middle (MitM) ontology and the system API theories in the MitM paradigm are big theory graphs with thousands of 
nodes and edges. Understanding and interacting with such large and complex objects is very difficult.The FAU group has conducted
research into whether virtual reality technologies are helpful for this task. We have presented a first working prototype at the
Conference on Intelligent Computer Mathematics CICM 2018 and the author: Richard Marcus - a master's student at FAU has received a
prize for best presentation.https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/1f175efc40146570fafa28123bed6b1198c41b00
/_posts/2018-08-20-tgview3d.md


Publishing reproducible logbooks %Jane voir site ODK
    Some instances
    Computing isomorphisms and embeddings of finite fields, Ludovic Brieulle, Luca De Feo, Javad Doliskani, Jean-Pierre Flori and Éric 
    Schost, arXiv:1705.01221 [cs.SC] 
    The 0-Rook monoid and its representation theory, Hivert and Gay
    %https://opendreamkit.org/2017/11/02/use-case-publishing-reproducible-notebooks/
    %https://opendreamkit.org/tag/use-case
    %https://opendreamkit.org/2018/03/15/jupyterhub-binder-convergence/
     

Qantitative metrics:
    \begin{itemize}
    \item Number of PyPI hosted packages for \Sage, and similarly for other components.
     https://github.com/OpenDreamKit/OpenDreamKit/blob/88ee2544e31e661c4eac7f39fa7333f24b62ab2a/WP3/D3.3/report.tex
  
  %demander précisions à Luca de feo
  %https://mybinder.org/v2/gh/opendreamkit/opendreamkit.github.io/rp2-wp3-binder-presentation?filepath=meetings%2F2018-10-28-Luxembourg%2FProjectReview%2FWP3.ipynb
  Conda packages
      From Conda Forge:
      Jupyter 380K downloads: https://anaconda.org/conda-forge/jupyter
      Sage 6K downloads: https://anaconda.org/conda-forge/sage

      Sage on PyPI
      80 packages in PyPI, not all Sage-related.
      Make them easily discoverable, document workflows.


    GAP packages code coverage: 69% (4.9) → 75% (4.10).
    Freshness of GAP packages: 50% released in the last year.
    SageMath on Windows: 44% (happy) Windows users.
    Packaging (not counting alt. methods, such as Conda):
    Arch: 50% of Jupyter users are also ODK users;
    Debian: 10% of Jupyter users are also ODK users.
    Medium sized VRE deployments: 20h of work, 244 LOCs.
    Docker Hub: 4K-10K pulls of ODK images.
 %https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/af29108dd216972485a17a3444caf6eee569033a/meetings/2018-10-28-Luxembourg/ProjectReview/WP3.ipynb

The OpenDreamKit project aims to build a virtual research environment for mathematics from an ecosystem of open source mathematical software 
systems. A crucial subgoal in this is to enable the systems to communicate mathematical objects, e.g. concrete elliptic curves or groups.
Several approaches exist (direct c-level library calls, semantic handles, …),, cattering for different needs, and OpenDreamKit pursues them 
all. Of particular interest is the use of OpenMath, a format which is designed for a content-oriented representation of mathematical 
concepts, objects, and models. To create common meaning space, the OpenDreamKit project is pursuing the Math-in-the-Middle (MitM) paradigm 
to establish an ontology of mathematics that can act as a joint context for the communication.

In the OpenMath terminology, the MitM ontology acts as a set of content dictionaries (CDs) that anchor the OpenMath objects semantically. 
For communication between the systems we only need to equip them with OpenMath phrasebooks (I/O libraries for OpenMath objects). The MitM 
ontology will be semi-automatically curated and connected to “system interface ontologies” (CDs for the system objects).

%Michael Kohlhase
%https://opendreamkit.org/meetings/2018-10-28-Luxembourg/ProjectReview/WP6.pdf
MitM-connected Systems: four (GAP, Sage, LMFDB, Singular)(See D6.5)
Formal MitM Ontology: 55 files, 2600 LoF, 360 commits (See D6.8)
Informal MitM Ontology: 815 theories, 1700 concepts in English, German,(Romanian, Chinese)
MitM System API Theories (GAP, Sage, LMFDB, Singular): 1.000+ Theories, 22.000 Concepts.

First Math-In-The-Middle-based interoperability prototype
Thanks to a fully functional prototype integrating of at least the systems GAP, SAGE, S INGULAR, and LMFDB via the SCSCP Protocol, end users 
shall be able to run calculations involving any combination of those systems from any of them. This prototype will be the basis for 
integration work for additional systems and the user interface from WP4.(#264)

The current state of play is that we have initial exports of system interface ontologies for three systems (the exporters are under 
development still, so your mileage may vary).

    SageMath (265 CDs); see them on MathHub.info
    GAP (210 CDs with 2996 symbols) see them on MathHub.info, and
    (partially) and the LMFDB.

%Computational Foundation for Python/Sage”. In the course of the deliberations in the WP6 workshops we saw a shift from the development of computational foundations and verification towards API/Interface function specifications to enable semantic system interoperability via the Math-in-the-Middle Ontology. Consequently, emphasis has changed to the generation of API Content Dictionaries (API CDs) for GAP, LMFDB and SAGE. We have a prototypical set of GAP and SAGE Content Dictionaries in OMDoc/MMT form (GAP: 218 CDs, 2996 entries; SAGE: 512 CDs, 2800 entries overall). The computational foundations exist but are rather more simple than originally anticipated. Much of the functionality has been offloaded to the SCSCP standard – remote procedure call with OpenMath representations of the mathematical objects – developed in the SCIENCE Project. As a direct consequence of the work in OpenDreamKit the OpenMath Society has promoted the SCSCP protocol into as an OpenMath Standard.Conversely, the GAP and SAGE CDs are rather more elaborated than anticipated in the proposal, and thus form a viable basis for alignment with the MitM Ontology.
% https://github.com/OpenDreamKit/OpenDreamKit/blob/88ee2544e31e661c4eac7f39fa7333f24b62ab2a/terminations/report/wp6.tex
%https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/333b21524f753dd49e441b4b1157ded68999eeff/meetings/2018-02-01-SteeringCommitteeMeeting/ProgressReports/Zurich.md

\item Number of additional systems made interoperable with the Math-in-the-Middle architecture, on top of the three for the Month
  36 prototype
  % Michael: I am expecting about 2 (OEIS and Findstat) any more than that would be a success.conv: {dissem: PU, due: '24', issue: '138', label: D6.4, lead: ZH, nature: DEC,



Usage/impact statistics (since last reporting period)
Further development on the nbdime project, delivered in the first period (D4.6). Has been met with enthusiasm, adoption in the community.
Jupyter Notebook and Jupyter Lab extensions, git integration.

Nbdime: 855 stars on github, 64 contributors (36 in 12 months prior), 611 
comments, 239 new issues (241 closed).

Workshop on live documents hosted in Oslo. Resulted in new package: thebelab, for embedding execution on any page, integrating tools from 
JupyterLab and MyBinder.org, demonstrating value of coordination
Thebelab: 44 stars, 15 contributors, 151 new issues (118 closed), 323 
comments.

3D visualization in Jupyter notebooks D4.12: Jupyter extension for 3D vis, demonstrated with fluid-dynamics
K3D-Jupyter is a package created by OpenDreamKit which provides a fast and simple 3d plotting tool in the Jupyter notebook. The primary aim 
of K3D is to be easy for use as stand alone package but also to allow interoperation with existing libraries as VTK. The power of ipywidgets 
makes it also a fast and performant visualisation tool for HPC computing e.g. fluid dynamics.

K3D-Jupyter: 48 stars, 14 contributors, 140 new issues (129 closed), 303 
comments.
%https://opendreamkit.org/meetings/2018-10-28-Luxembourg/ProjectReview/wp4.pdf

Packages:
k3d-jupyter
ipydatawidgets
ipyscales○unray
Improved distribution

\subsubsection{KPIs for Aim 4 / Dissemination / Impact}

\begin{recommendation}{Aim 4}
  Maximise sustainability and impact in mathematics, neighbouring fields, and scientific computing.\end{recommendation}


An important part of the success of the ODK project is linked to its ability to foster a community in the spirit of the open source projects 
it is built on. Part of this relies on the organization and participation to scientific and development events of many different scales and 
objectives. We gave a first overview of the actions taken during the first year of the project. This includes:

    Organization of development workshops
    Organization of dissemination workshops in different thematic linked to ODK
    Training intervention in external events
    Communication interventions and participation to external events
    Actions taken to foster a community in developing countries


Qualitative metrics: success stories. For example:
\begin{itemize}
\item Women in \Sage workshops;
\item Use of \Jupyter, \cocalc, ... for teaching, at Sheffield, at the
  African Institute for Mathematical Sciences, ...;
\item Stories about the impact of the Micromagnetics VRE;
\item Impact of nbdime, 3D widgets, Thebelab, ...
% D4.13 (Sphinx) might have impact beyond ODK. For example, Simon King
%is interested in things done for D4.13 for his
%\software{p\_group\_cohomology} package. It might also lead to a PEP,
%which by itself counts as impact.
\end{itemize}

   
 The under-representation of women in the scientific world is even more visible if we intersect science with software
  development. As we know, we have many talented women in our community, and we have organised some events targeted at women in the
  spirit of the "Women in Sage" days that happened many times in the US already. %e are planning to have one more on 2019......

Women in Sage Events:
    Organised by OpenDreamKit in Paris January 2017
    Another event (independent of ODK) in Montreal in June 2018
    Next ODK one: Spring/Summer 2019 in Crete

Women in computing
    Developed training materials and provided training for over 130 women in the last 12 months at Sheffield and Manchester in partnership 
    with CodeFirstGirls.
    Tania Allard participated in the Diversity and Inclusion in Scientific Computing unconference by direct invite of NumFOCUS
    Tania Allard was diversity chair for the 2017 International Research Software Engineering conference.

Women in sage workshops
    Last January, Viviane Pons, Jessica Striker and Jennifer Balakrishnan organized the first WomenInSage event in Europe with OpenDreamKit.      
    20 women spent a week together coding and learning in a rented house in the Paris area.
    We took advantage of the diverse knowledge background of our group to work together and learn from each other. It was an occasion for 
    many "first times" among participants who had very little experience with Sage:

    5 participants installed a source version of Sage for the first time (so that they could edit the source).
    3 used git for the first time.
    5 used git within Sage for the first time.
    11 got their first Trac account .
    5 got their first contribution to a Sage ticket.
    8 are in the process of getting their first code integrated to Sage.

We worked on 14 tickets during the week, 6 of those which have been merged since the conference. All participants said they had learned new 
things and it would impact their careers. %https://github.com/OpenDreamKit/OpenDreamKit.github.io/blob/6faf6eb2f1532f342f86c8da633078067ca40c85/_posts/2017-04-06-WomenInSage.md

\subsection{Teaching}\label{teaching}
We showcase here a few of the many uses in teaching of the Jupyter kernels we contributed to:

     \item Since 2017, Jupyter is used at Université Paris Sud for teaching C++ to over 300 students.This was initiated in particular by
     \ODK participants Loïc Gouarin, Viviane Pons, and Nicolas M.\ Thiéry. The mix of narrative documents and interactive programming 
     fostered active participation from the students while our web-based deployment made it easier for them to work from home.The course 
     material is available from \url{http://Nicolas.Thiery.name/Enseignement/Info111}.
     
     \item In early spring 2017, Prof.~Dr.~W.~Decker and Prof.~Dr.~G.~Pfister gave a three-week course on computational algebraic geometry 
     at the African Institute for Mathematical sciences (Cape Town, South Africa) with lectures and computer lab sessions.
     The course was attended by about 50 students from all over Africa. In the lab sessions, the students learned how to experiment with the
     computer algebra system Singular. It proved extremely valuable that the students could run Singular in the Jupyter notebook.
     
     \item The \GAP Jupyter kernel was used by Pedro Garcia-Sanchez to teach a master course in mathematical software at the University of
     Granada. See  \url{https://github.com/pedritomelenas/Software-Matematicas-GAP}. Pedro has taken on the technology, and is now involved
     in the   development of interactive visualization widgets for discrete maths (package Francy; see also~\longdelivref{UI}{ipython
     advanced-interacts}) in particular for use in other courses.
     
     \item The \GAP Jupyter kernel was also used by Alexander Konovalov for running a \href{https://github.com/alex-konovalov/gap-teaching
     /}{GAP tutorial} at the Postgraduate Group Theory Conference \href{http://www-groups.mcs.st-and.ac.uk/~pgtc2018/}{PGTC 2018}.
     \end{itemize}



Quantitative metrics:
\begin{itemize}
\item Statistics on workshops organized and conference presentations delivered as part of our dissemination activities, including estimates of number of attendees and what (if anything) happened as a follow-up.
\item Number of courses and departments we worked with directly and an estimate of how many students this subsequently affected.

  For example:
  \begin{itemize}
  \item Department of physics, University of Sheffield. Estimated number of new users: 500 undergraduate students
  \item Department of Biomedical Sciences, University of Sheffield.BMS353. 2015: N students. 2016: Y students
  \end{itemize}
\end{itemize}

\begin{itemize}
  \item University of Ghent
  Department of Mathematics (Bachelor, Master). Courses: CoCalc, Jupyter, SageMath. Estimated number of new users: 31 students
  Department of Chemistry (Master). Courses: SageMath. Estimated number of new users: 37 students
  
  \item Paris Sud University 
  Department of Algèbre et Calcul Formel. Courses: Jupyter, SageMath. Estimated number of new users: 25 students yearly.
  Department of Introduction à l'Informatique. Courses: Jupyter, C++. Estimated number of new users: 400 students yearly.
  Department of Algorithmique, ingénieurs. Courses: Jupyter. Estimated number of new users: 20 students yearly.
  Department of Algorithmique M1. Courses: Jupyter. Estimated number of new users: 20 students yearly.
  Department of Combinatorics M2. Courses: Jupyter. Estimated number of new users: 20 students yearly.
  Department of Projet Math-info L1. Courses: SageMath, Jupyter. Estimated number of new users: 40 students yearly.
  
  \item FAU Erlangen-Nürnberg
  Department of Logic-based Natural Language Semantics. Courses: MMT is the logic, domain modeling and reasoning system. Estimated number of new users: 5-10 yearly. 
  Department of Knowledge Representation for Mathematics and Technology. Courses: MMT is the logic, domain modeling and reasoning system. Estimated number of new users: 5-10 yearly. 
 
 \item  UVSQ
 Department of Algèbre Commutative et Effective (M). Courses: Jupyterhub, SageMath, Planetaryum. Estimated number of new users: 22 students.
 Department of Algorithmique et Programmation C (M). Courses: JupyterHub, JupyterLab, C. Estimated number of new users: 22 students.
 Department of Introduction au calcul formel (M). Courses: Jupyterhub, SageMath. Estimated number of new users: 18 students.
 Department of Algorithmique (M). Courses: Jupyterhub, JupyterLab, Python. Estimated number of users: 18 students.
 Some database stuff (M). Courses: JupyterHub, Apach Spark. Estimated number of users: 60 students. 

\item  Sheffield
 Department of physics. Courses: .................. . Estimated number of new users: 500 students.
 Department of Biomedical Sciences. Courses: .............................. . Estimated number of new users: .... students.
%#277) https://github.com/OpenDreamKit/OpenDreamKit/issues/277


Over the last four years OpenDreamKit has been involved in 66 events, 31 organized/co-organized by UPSud.

                        - 21 development workshops and project meetings, among these: 14 organized or co-organized by UPSud
                        - 26 training workshops or sessions,  7 org./co-org. by UPSud adding 1100 trainees
                        - 13 community building workshops, 8 org./co-org. by UPSud
                        - 6 research workshops, 2 org./co-org. by UPSud

And 19 external events ( 9 organized/co-organized by UPSud).

We have worked on our online communication
        new website
        6475 visits in 2018
        452 followers on Twitter
        Video interviews and explainer material
        
  Use of ODK technologies for teaching: success of the Bioinformatics Awareness Days
An event that took place in The Sheffield Institute for Translational Neuroscience on November 2017 where were given training activities 
about bioinformatics workflows using Jupyter notebooks with computation provided by the free Micosost Azure Notebook service.The event 
demonstrated that \ODK supported technologies could be applied to the field of Bioinformatics and led to a new collaboration between Dr 
Cutillo and \ODK member Mike Croucher.

Following the success of this workshop, Dr Cutillo independently taught an introductory workshop on statistics using Jupyter notebooks 
on Azure at Parthenope University of Naples (Materials at \url{https://github.com/luisacutillo78/RbasicStats}) r Cutillo has since moved 
to University of Leeds where she will be teaching statistics to 200+ undergraduates. She plans to use \ODK developed technologies 
in collaboration with the Research Software Engineering group at Leeds.

The event required also the development of a website that was linked to the Jupyter notebooks (\url{https://bitsandchips.me/BAD_days/}). The 
website caught the attention of Eleni Vasilaki, Head of Machine Learning at University of Sheffield who wanted to do something similar for 
her course on Adaptive Intelligence. We supported her in this endeavour and the result is at (\url{http://bitsandchips.me
/COM3240_Adaptive_Intelligence/}).

In order to better support this, \ODK member Tania Allard, developed a Jekyll template for use by academics and researchers using Jupyter 
notebooks for course materials and dissemination. This led to the development of a Python package: nbjekyll (\url{https://github.com
/trallard/nbjekyll}) that complements the Jekyll template.As well as being used internally at Sheffield, The nbjekyll package received some 
attention on twitter \url{https://twitter.com/jdblischak/status/1009800776305332224} and \url{https://twitter.com/walkingrandomly/status
/1009414151716909057} receiving a total of 42 retweets and 80 'likes'% ( à remettre à jour).
  






















