\documentclass{deliverablereport}

\deliverable{management}{ipr2}
\duedate{31/08/2018 (M36)}
\deliverydate{Unknown}

\usepackage[style=alphabetic,backend=bibtex]{biblatex}
\addbibresource{../../lib/kbibs/kwarcpubs.bib}
\addbibresource{../../lib/kbibs/extpubs.bib}
\addbibresource{../../lib/kbibs/kwarccrossrefs.bib}
\addbibresource{../../lib/kbibs/extcrossrefs.bib}
\addbibresource{../../lib/deliverables.bib}
\addbibresource{../../lib/publications.bib}
% temporary fix due to http://tex.stackexchange.com/questions/311426/bibliography-error-use-of-blxbblverbaddi-doesnt-match-its-definition-ve
\makeatletter\def\blx@maxline{77}\makeatother

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Styling: adapt amsart's subsubsection macro to put a newline after the title
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand\subsubsection{\@startsection{subsubsection}{2}%
  \z@{.5\linespacing\@plus.7\linespacing}{.1\linespacing}%
  {\normalfont\bfseries}}

% Variant of taskref that links to the section on the task in this file
\newcommand\localtaskref[2]{\hyperref[#1@#2]{\csname task@#1@#2@label\endcsname}}
\newcommand\longmilestoneref[1]{\textbf{\csname mile@#1@label\endcsname}:
  ``\csname mile@#1@title\endcsname''
  (month \csname mile@#1@month\endcsname)}
\makeatother

%\usepackage{todonotes}
\author{Nicolas M. Thiéry, Benoît Pilorget, et al.}

\begin{document}
\enlargethispage{4ex}
\maketitle
\githubissuedescription
\tableofcontents\newpage

\section{Progress on the project}

In this section, we give a general overview of the progress of the
project in the years 2017 and 2018. We start by recalling some context of \ODK's
approach that is important to understand and evaluate the
progress. Then we describe the general state, and in more detail
the progress of the work packages.

\subsection{Some context: \ODK's approach}
\ODK's approach to delivering a Virtual Research Environment (VRE) for
mathematics is not to build a monolithic one-size-fits-all VRE, but
rather a toolkit from which it is easy to set up VRE's that are
customised to specific needs by combining the appropriate components
(collaborative workspaces, user interfaces, computational software,
databases, \dots) on top of available physical resources (from
personal laptops to cloud infrastructure). This approach --- chosen by
design --- allows to flexibly put together lean computational
environments and tools for particular research challenges. These tools
provide the required functionality but due to the component based
approach carry no unnecessary bloat that would reduce effectiveness in
terms of installation process, size, computation time, and
reproducibility.

Most of the components preexist as an ecosystem of open source
software, developed by well established communities of developers. For
example, for interactive computing and data analysis, OpenDreamKit
promotes Jupyter, a web-based general purpose flexible notebook
interface\footnote{a notebook is a document that contains live code,
  equations, visualizations and explanatory text} that targets all
areas of science. A number of Virtual Research Environment already
exist, e.g.\ powered by SageMathCloud or JupyterHub.

Hence most of the work in \ODK\ is to foster this ecosystem, improving
the components themselves and their composability. The technical is
distributed over the work packages:
\begin{itemize}
\item Component architecture (WP3):
  \begin{itemize}
  \item ease of deployment: modularity, packaging, portability,
    distribution, for individual components and combinations thereof.
  \item sustainability of the ecosystem: improving the development workflows.
  \end{itemize}
\item User Interfaces (WP4): enable Jupyter as uniform notebook
  interface, and further improve it; foster the collaboration between
  SageMathCloud and JupyterHub; generally speaking investigate
  collaborative, reproducible, and active documents.
\item Performance (WP5): make the most of available hardware
  (multi-core, HPC, cloud), for individual computational components and
  combinations thereof.
\item Data/Knowledge/Software (WP6): enable rich and robust
  interaction between computational components, data bases, knowledge
  bases, users through explicit common semantic spaces, a language to
  express them, and tools to leverage them.
\end{itemize}
These technical work packages are supported by the following two activities:
\begin{itemize}
\item Community building and dissemination (WP2): developer and
  training workshops, conferences, teaching material with focus on
  making the created value accessible to a wide, varied and growing user community.
\item Studies of Social Aspects (WP7): analysis of user needs and
  research into collaborative and open software development in
  mathematics and science.
\end{itemize}

As a result of \ODK's approach, the work programme for \ODK\ consists
of a large array of loosely coupled tasks, each being useful in its
own right, and none being absolutely critical.

The second and third years confirmed that this is a strong feature of \ODK's
approach. Indeed, as analysed in the proposal, this kind of project is
subject to the following risks:
\begin{itemize}
\item Recruitment of qualified personnel;
\item Different groups not forming effective team;
\item Implementing infrastructure that does not match the needs of end-users;
\item Lack of predictability for tasks that are pursued jointly with
  the community;
\item Reliance on external software components.
\end{itemize}
Together with ambitious software challenges, this makes the accurate
prediction of workload and precise timeline of work packages
difficult, especially over a period of four years in a field of
rapidly evolving technologies.

The loose coupling allows some flexibility,
permitting to modify the tasks schedule and human resources allocation,
with little influence on the general aims and objectives.

\subsection{General progress}

Intensive work has now started on almost all fronts of the
project\footnote{status reports delivered at the St Andrews project
  meeting (January 2016) and at the Bremen's project meeting (June
  2016) helped the Coordinator to track the progress}. A few tasks
(and the corresponding deliverables) have been postponed by a couple
months due to recruitment delays. This concerns mostly the
micro-magnetic VRE demonstrator
(\taskref{dissem}{dissemination-of-oommf-nb-virtual-environment},
\taskref{dissem}{dissemination-of-oommf-nb-workshops},
\taskref{dissem}{ibook},
\taskref{component-architecture}{oommf-python-interface},
\taskref{UI}{oommf-py-ipython-attributes},
\taskref{UI}{oommf-tutorial-and-documentation},
\taskref{UI}{oommf-nb-ve},
\taskref{social-aspects}{oommf-nb-evaluation}), where recruitment
at Southampton initially proceeded at expected speed but eventually
experienced delays of several months outside our control due to unusually high demand on the UK Home
Office which had to process work permit paperwork for
the successful candidate (attributed to a high number of immigration applications in the run up to the
Brexit referendum). Some deliverables got
delayed as well by a couple months due to unexpected technical
difficulties or misplanning (e.g. \delivref{hpc}{pythran-typing},
\delivref{hpc}{sage-paral-tree}, \delivref{UI}{pari-python-lib1}). All
these delays have been included in the amendment of the Grant
Agreement, which was necessary to include UGent in the
consortium. On the other hand, we are happy to report below on very
strong recruitment (see Section 3.1), as well as unexpectedly rapid
progress on portability and packaging aspects. Also WP6
(Data/Knowledge/Software) has witnessed a particularly strong and
early uptake, with active involvement of many of the participants and
promising outcomes.

All in all, \ODK\ is running according to its plan, and its first
outcomes are already benefiting the mathematical community and
beyond. September 2016 will see the start of Key Performance
Indicators.  These KPIs, which will be more precisely and
realistically defined then, will give results for the 1st Reporting
Period (RP1) at month 18.  This way we will hopefully be able to see the
evolution of the impact OpenDreamKit has had between the RP1 and RP2,
at month 36.

\include{objectives}

\subsection{Achievements and ongoing progress in workpackages}

\include{WP1}
\include{WP2}
\include{WP3}
\include{WP4}
\include{WP5}
\include{WP6}

\section{Risk management}
\subsection{Recruitment of highly qualified staff}
Recruitment of highly qualified staff was planned to be a high risk
when the Proposal was written. And unfortunately it turned out we were
right. In such a field as computer science and software development,
potential candidates who are likely to be fairly young considering
only temporary positions are offered, are very scarce. Furthermore
they need to make a choice between public and private bodies which are
very attractive, and the choice between pure development and research.
Because of this difficulty to recruit in the past year, there have
been slight changes in the workplan, which do
not put the project results at risk.

The following people were hired in the past year:\\


\begin{tabular}{|l|c|r|r|r|}
\hline
NAME&GENDER&PARTNER&POSITION&HIRING DATE\\
\hline
Benoît PILORGET&M&\site{PS}&Project manager&17-09-2015\\
Jeroen DEMEYER&M&\site{PS}&Research engineer&01-03-2016\\
Erik BRAY&M&\site{PS}&Research engineer&01-01-2016\\
Christian MAEDER&M&JacobsUni&Senior researcher&01-01-2016\\
Tom WIESING&M&JacobsUni&Junior researcher&01-09-2015\\
Xu HE&M&JacobsUni&Junior Researcher&01-09-2015\\
Alexander BEST&M&UNIKL&Research engineer&01-02-2016\\
Anders JENSEN&M&UNIKL&Postdoc&01-11-2015\\
Alexander KRUPPA&M&UNIKL&Postdoc&01-08-2016\\
Jan AKSAMIT&M&USlaski&Technical staff&01-10-2015\\
Marijan BEG&M&Southampton&Research fellow&01-05-2016\\
B. RAGAN-KELLEY&M&Simula&Postdoc&01-09-2015\\
V.T. FAUSKE&M&Simula&Postdoc fellow&02-05-2016\\
\hline
\end{tabular}\\
~\\
 \ODK partners had to face some Human Resources issues in the past year:
\begin{itemize}
\item{\site{PS}:}
  Thanks to an early start in the recruitment process, and despite
  some difficulties in attracting experienced candidates for a part
  time position, the project manager position (24PM) was filled by
  Benoît Pilorget shortly after the start of the project. Unfortunately at month 36 the project
finds itself without a Project manager since the departure of B. Pilorget.

  The recruitment of \site{PS}'s first Research Engineer (48PM) was
  delayed by four months because the top ranked candidate for this
  position, Erik Bray, was originating from the US and needed time to
  arrange for his moving; there were also some administrative delays
  (visa, ...).

  The second Research Engineer position (36PM) was more problematic
  for internal administrative reasons. The top ranked candidate,
  Jeroen Demeyer, had the perfect profile; however for family reasons,
  he wished to work most of the time from Ghent in Belgium. After eight
  months investigating an administrative solution to hire him at
  \site{PS}, and a temporary four month solution, it was decided with
  OpenDreamKit's Steering Committee and Project Officer to instead add
  Ghent's university as new partner, hire Jeroen Demeyer there, with an
  adequate budget transfer and amendment to the Grant Agreement.

  Those delays have induced late start on several tasks, and costed
  much management time. However the excellence of the recruitment, well
  confirmed by the results obtained so far, was worth it and will soon
  compensate for the late start.

  In addition to this, a three year PhD position was open to work on
  WP6, starting from Month 12. By lack of suitable candidate, this
  position will be converted into a two year PostDoc position,
  presumably starting at Month 24. Active advertising has started and
  there are some tentative candidates. The relevant deliverables being
  due late in
  the project, no delay is to be expected from this change.\\

\item{CNRS:} Because the research engineer offer (48PM) was still not
  filled in the Summer 2016, the CNRS decided to divide the position
  in two full positions of 24 PM each. As a result, a candidate was
  already selected for one of the two positions and should begin his
  work this Fall 2016.  Thanks to the PM division, there should be no
  delay in any task or
  deliverable. \\

\item{JacobsUni:} Michael Kohlase, lead PI for Jacobs University, has
  moved on 01/09/2016 to Friedrich-Alexander-Universität
  Erlangen-Nürnberg, and most of his team will follow him. Since he is
  a critical asset for OpenDreamKit, a Grant Agreement amendment will
  be submitted in Winter 2016-2017 to update the consortium accordingly.\\

\item{UJF:} The original tentative candidate for UJF's Research
  Engineer position (12PM, planned to start on Month 1), Pierrick
  Brunet, finally declined the position to accept an alternative
  permanent offer. The position will be filled by another candidate in
  Autumn 2016. This induced a delay of Deliverable \textbf{D5.2} from Month 12
  to Month 18, without impact on other tasks.\\

\item{UNIKL:} UNIKL had to split the 12 PM planned for a software developer into 2 shorter positions (Anders Jensen and Alexander Kruppa) in order to deliver the planned work on time. Indeed the few qualified persons for this job were not able to accept this 12 months position during the timelapse planned within the project.\\

\item{USFD:} The University of Sheffield has also been struggling in the the hiring process of a postdoc (36PM). The position should be filled this Autumn.\\

\item{Southampton:} Southampton faced administrative difficulties in the recruitment of Marijan Beg (38PM) as a post-doc, due to the Croatian nationality of Mr Beg. His recruitment was delayed of four months, and therefore some tasks and deliverables, planned to be borne around the end of the project, were postponed of four months. However no serious delay nor implication on the main tasks of OpenDreamKit followed these difficulties.\\

\item{UVSQ:} Nicolas Gama is currently on a long-term leave until September 2017. This will not affect the project in any way.\\

\item{UZH:} The University of Zürich partner is only composed of one person, Paul-Olivier Dehaye, who does not enjoy a permanent position there. There have been worries that Mr Dehaye's contract with his university might end earlier than planned within OpenDreamKit. But thanks to the action of the OpenDreamKit steering committee, Mr Dehaye has been technically rehired by UZH as a scientific consultant for as long as the project needs.\\

\item{Simula:} Everything is fine concerning temporary staff
  recruitment on the Simula side, however we have had to endure the
  hazards of human ressources with Hans-Peter Langtanger (the PI when
  the Grant was signed) being on a long-term sick leave, and with
  Martin Alnaes replacing him as PI currently on a paternity
  leave. However Benjamin Ragan-Kelley has stepped in to lead the
  Simula contribution in the meantime and all planned tasks are on time.\\
\end{itemize}


Altogether, this first year confirmed that the recruitment of highly
qualified staff is indeed a risky endeavour, which induced delays on
several deliverables. However the planned mitigation measures --
taking into account the pool of potential candidates in the design of
the positions, aggressive advertisement, weak coupling between tasks
-- worked adequately: with appropriate reshuffling of the work plan,
we don't expect an impact on the overall progress of the project.

\subsection{Different groups not forming effective team}

As expected, this risk was tamed by the existence of many preexisting
collaborations between the partners and of ``joint itches to scratch
together'' (to use a common open source software metaphor). The
organization of many joint workshops (for example the Sage-GAP
workshop, the Atelier Pari attended by SageMath developers, the WP6
workshops) helped bootstrap joint activities through brainstorms and
coding sprints. Upcoming workshops are planned on Year~2 to strengthen
collaborations with the social aspects team in Oxford and the Singular
team in Kaiserslautern.


\subsection{Implementing infrastructure that does not match the needs of end-users}

The consortium is keeping in their minds the end-user needs. Since
OpenDreamKit is improving already existent software which have their
own users, their needs are naturally met. However Key performance
Indicators will evaluate the effects of OpenDreamKit on these
software. KPIs, indicated in the Proposal, will be launched this
Autumn with the help of the end-user group which was merged with the
Advisory Board. Constant links between the accomplished work and the
end-user needs should be made in WP2 deliverables and also in WP7
deliverables when relevant.  Open tracking of KPIs evolution can be
found on
\href{https://github.com/OpenDreamKit/OpenDreamKit/labels/KPI}{GitHub}.

\subsection{Lack of predictability for tasks that are pursued jointly
  with the community}

As planned, we are regularly shifting manpower around to adapt for the
variability of the involvement of the community in the different
tasks. For example, the SageMath Jupyter kernel of
\longdelivref{UI}{ipython-kernels-basic} was mostly implemented by the
community which allowed to focus on other tasks such as the long term
task~\longdelivref{component-architecture}{portability-cygwin}.  On
the other hand many other deliverables were implemented with very
little help from the community.

\subsection{Reliance on external software components}

There is not much to report on this front yet: none of the external
software component we rely on have failed us. Quite on the contrary,
critical software like \Jupyter have continued to blossom. Besides the
high modularity of the design means few components are critical to the
overall success of the project.

\section{Quality assurance plan}

\subsection{Deliverables quality: Quality Review Board}

The Quality Review Board is the Consortium Body that fosters best
possible quality in the deliverables. All four members of the board
have a research interest in the quality of software in computational
science, and use and share their experience to benefit the quality of
the work.

The board is chaired by Hans Fangohr, from the University of
Southampton and European XFEL GmbH. He is supported in this task by
Mike Croucher from the University of Leeds, Alexander Konovalov from
the University of St Andrews, and by Konrad Hinsen from the Centre de
Biophysique Moléculaire with whom a Non-Disclosure Agreement was
signed.

The members engage with European initiatives working towards
improvement of the software quality in research, in particular in
computational and data science; both as voluntary activities and key
of their professional roles. Mike Croucher is the head of research
computing at Leeds, well know through his outreach blog and an EPSRC
Research Software Engineering Fellow, Alexander Konovalov is a driving
force in the Software Sustainability Institute, Konrad Hinsen has
founded and is editing the ReScience Journal for reproducible Science,
and Hans Fangohr is founder and director of the UK's only centre for
doctoral training in computational modelling, a fellow of the Software
Sustainability Institute, was chairing the EPSRC's national scientific
advisory committee on high performance computing, and is leading big
data analysis infrastructure development at the European XFEL research
facility.

The quality review board has reviewed deliverables at the end of
reporting period 1, identified good practice - both in terms of
software engineering content but also presentation of the work -,
produced a report, and shared the findings with all members in the
project to improve the quality of the remaining deliverables. The
board has stuck to its no-blame culture in its reporting.

A similar process is underway for the end of reporting period 2.

%\ODK consortium about the deliverables content and layout.

% While the primary focus of the board is on the OpenDreamKit project
% and the software it develops,
% some of the lessons may be more widely applicable and be made publicly
% available.





% % original

% The content form of deliverables due by then had to meet the
% expectations of the Project Officer and of Reviewers. Following this
% experience, we have concluded that deliverables should be written in
% Latex using a style file created for this purpose. For deliverables
% that are not reports by themselves, it's appropriate to have a
% relatively short report with a link to the github issue, and a copy of
% the description of this issue. In all cases, the report shall be
% self-contained. Deliverables are indeed evaluated based upon their
% versions submitted on the EU portal without retrieving other
% resources. Links have no legal value, since there is no guarantee that
% the referenced material will remain unchanged.  Partners who have a
% deliverable due at month 12 (August 2016) have been following these
% tips. The feedback of the Official review will help the Quality review
% Board in ensuring the quality of reports meets the needs. This will be
% up to the Quality Review Board to meet after the 1st reporting period
% and to decide if the quality of deliverables is acceptable.  They will
% aim at identifying good practice and weaknesses, and to share the
% lessons with the project to improve any future project work. The board
% will focus on selected deliverables and investigate those in detail
% rather than attempting a superficial inspection of all deliverables.


\subsection{Infrastructure quality: End-user group}


It was decided by the Steering Committee during the
\href{http://opendreamkit.org/meetings/2015-09-02-Kickoff/management_structure/}{kick-off
  meeting} to slightly modify the management structure by having only
one gender-friendly Advisory Board composed of 6 people (as agreed a
few months later at the
\href{http://opendreamkit.org/meetings/2016-06-27-Bremen/minutes/}{Bremen
  meeting}), some of which to be end-users.

Members of the board are: Jacques Carette from the McMaster University, Istvan Csabaifrom the Eötvös University Budapest,
Françoise Genova from the Observatoire de Strasbourg, Konrad Hinsen from the Centre de Biophysique Moléculaire,
William Stein who is CEO of SageMath, Inc. (SME), and Paul Zimmermann from INRIA.



\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  maketitle githubissuedescription newpage newcommand xspace Jupyter dissem
%  LocalWords:  tableofcontents visualizations composability itemize analyzed taskref hpc
%  LocalWords:  dissemination-of-oommf-nb-virtual-environment taskref dissem taskref pn
%  LocalWords:  dissemination-of-oommf-nb-workshops dissem ibook taskref taskref taskref
%  LocalWords:  oommf-python-interface oommf-py-ipython-attributes taskref oommf-nb-ve
%  LocalWords:  oommf-tutorial-and-documentation taskref oommf-nb-evaluation taskrefs
%  LocalWords:  delivref pythran-typing sage-paral-tree subsubsection organized Dagstuhl
%  LocalWords:  co-organized organization modularization ipython-kernels nbdime Pythran
%  LocalWords:  jupyter-collab ystok WPref dksbases compactitem emph WPtref DehKohKon
%  LocalWords:  iop16 textbf tasktref lfmverif triformal formalized biformal ossp09 Dima
%  LocalWords:  hline Marijan Pilorget Pierrick Kruppa Dehaye Dehaye's Dehaye's Alnaes
%  LocalWords:  Konovalov Hinsen github printbibliography
