\documentclass[book]{deliverablereport}

\usepackage[style=alphabetic,backend=bibtex]{biblatex}
\addbibresource{../../lib/kbibs/kwarcpubs.bib}
\addbibresource{../../lib/kbibs/extpubs.bib}
\addbibresource{../../lib/kbibs/kwarccrossrefs.bib}
\addbibresource{../../lib/kbibs/extcrossrefs.bib}
\addbibresource{../../lib/deliverables.bib}
\addbibresource{local.bib}
% temporary fix due to http://tex.stackexchange.com/questions/311426/bibliography-error-use-of-blxbblverbaddi-doesnt-match-its-definition-ve
\makeatletter\def\blx@maxline{77}\makeatother
\usepackage{local}
\usepackage{ed}

\title{Notebook Import into MathHub.info (interactive display)}
\def\shorttitle{Notebooks in MathHub}

\deliverable{UI}{jupyter-import}
\deliverydate{15/07/2018}
\duedate{1/09/2018}

\author{Kai Amann}
\author{Michael Kohlhase}
\author{Florian Rabe}


\begin{document}
\begin{abstract}\strut\\
  \ednote{MK: to be done at the end}
\end{abstract}
\maketitle
%\newpage\strut\githubissuedescription
\newpage\tableofcontents\newpage

\section{Jupyter server on jupyter.mathub.info}
\ednote{KA: ask Tom about the server}
\ednote{KA: paper or report?}
\ednote{KA: show actual code? I suppose not}
The goal of this report is to integrate Jupyter notebooks into MathHub
and make them compatible with MMT, in a way that we can conveniently use 
MMT syntax in these notebooks and also a little bit of extra functionality
like e.g. the Jupyter widgets. The first step is setting up a Jupyter server,
which currently runs on \url{http://juypter.mathhub.info}. \ednote{KA: maybe show picture of it?}
For this server, we have developed a custom kernel, that forwards the input 
entered into the Jupyter notebook to the MMT backend. This then processes 
said input and sends the response back to the Jupyter frontend via the kernel.
We will cover the implementation of the Jupyter kernel and the MMT-backend,
later in this report.

\section{Jupyter kernel for MMT}\ednote{KA: isn't it more likely a MMT kernel for jupyter?}
In this section we will cover the implementation of the Jupyter kernel that
communicates with the Jupyter frontend and the MMT backend. 

\subsection{Command syntax}
\ednote{KA:I suppose we still want to be able to create theories etc.}
\ednote{KA: if we want to have other "commands" (e.g. active computation)} 
\ednote{KA: we will have to create a custom parser for that}
The syntax we can use in the Jupyter frontend is basically the same one we are able to 
use e.g. in JEdit extended with some additional features like active computation
or theory visualization via ?MathML?. 
\subsection{REPL extension in MMT}
\ednote{KA: + the JupyterKernel Class on the Scala side? or does this go to the next section
 I do have to look over this a little more since I did not really 
 have anything to do with the code there.
 I know that it handles Session management and passes the code
 to the parser (that is afaik also the parser for JEdit).}
\ednote{KA: here we would have to think of a Syntax that we want to use 
 and which applications we want for it e.g. the 
 keyword active computation opens some textbox 
 widgets for user input, based on the current theory}

\subsubsection{Pythonbridge, Implementation of the Python kernel}
Now to the actual implementation of the Jupyter Kernel. Although communication between 
the Python kernel and the Scala MMT backend via HTTP would suffice for basic things,
like e.g. MMT surface syntax commands, it proves exceedingly difficult when we want to 
provide more complex services to the notebook user that require the usage of Jupyter widgets,
like e.g. displaying a whole user-created theory as MathML, or active computation. 
So instead we use a Python-Java-bridge that lets us overcome the issues we have with HTTP and
make code more modular and easier to understand, on both Python and Scala side.
\ednote{KA: do we really want to explain the bridge here? if yes how detailed?}
\ednote{KA:  explain what widgets are and how they work? if yes how detailed?}
The Python-Java-Bridge basically works as follows: on the Scala side we only have method stubs
that are linked with methods on the Python side that implement the stubs. So when we call 
a method on the Scala side, we actually call the code that is implemented on the Python side 
and get the result of this function execution back in Scala. This mechanism allows us to 
conveniently register callbacks to the Jupyter widgets. You might ask yourself why those callbacks 
are so important. The answer is simple: a button that doesn't do anything when clicked is not 
a very useful one. Those callbacks enable widgets to interact with each other and the user alike. 
Therefore they are essential to using the full potential of the Jupyter widgets.
\ednote{KA: To use the Python-Java Bridge we first have to enable the Py4JGateway extension on our MMT server}


\section{Jupyter Notebooks on MathHub}
\ednote{KA: write some demo notebooks stored in gl.mathhub.info}
\subsection{special treatment when displaying notebooks in mathhub.info: "Runme button"}
\ednote{KA: show them in jupyter.mathhub.info}

\section{MitM-based notebook applications}
\subsection{MitM math formalization in MMT on MathHub}
\subsection{Domain specific applications: MOSIS}

\section{Plans for Future developments?}

\begin{oldpart}{MK}
  \section{Plan for the Document}
MathHub based on MMT for knowledge management. So MMT/Notebook integration is the main
feature we report on. Kai makes a first pass and then Florian and Michael iterate; do not
forget to talk about In-document computation~\cite{ODK-D4.9} and
MoSIS~\cite{PolKohKoe:kacse18}
\end{oldpart}

\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  maketitle newpage tableofcontents newpage newcommand xspace ednote mathdb
%  LocalWords:  standardize dktheories concl printbibliography pn textit mmt mitm emph
%  LocalWords:  WPref dksbases prioritized taskref organized delivref dkstheories Jupyter
%  LocalWords:  githubissuedescription JEdit visualization subsubsection Py4JGateway
%  LocalWords:  Runme formalization oldpart PolKohKoe:kacse18
