\documentclass{deliverablereport}

\deliverable{UI}{sage-sphinx}
\deliverydate{31/08/2018}
\duedate{31/08/2017 (M24)}
\author{Jeroen Demeyer}

\begin{document}
\maketitle
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

The \Sage computational system, like \Python and many other Python
projects, uses the \Sphinx documentation system. Similar in principle
to \lstinline{javadoc} or \lstinline{doxygen}, it generates structured
documentation in various formats (HTML, PDF, ...).
The documentation is generated either from \texttt{.rst} files
or from so-called \emph{docstrings}, that is snippets of documentation
interspersed in the code it documents.

One specific challenge is scaling to the sheer size of the \Sage
documentation (13k pages just for the reference manual): building it
requires a lot of time and memory. In addition -- due to those particularly
stringent needs and a long legacy -- \Sage was (at the time of writing
the proposal) using an outdated version of \Sphinx, crippled by many
layers of customisation and adaptations that had accumulated over the
years. This was becoming impossible to maintain, and impeding
\Sage's long term sustainability.

Task \taskref{UI}{sage-sphinx} tackles this situation with, broadly
speaking, two main goals:
\begin{enumerate}
\item a critically needed deep refactorisation to
  foster sustainability by outsourcing back to
  \Sphinx all generic aspects (parallel compilation, index generation,
  etc);
\item optimizing resource usage for the documentation build.
\end{enumerate}

Although a lot has already been achieved with more than twenty two
contributions to \Sage or upstream (see
\url{https://github.com/OpenDreamKit/OpenDreamKit/issues/87} for an up
to date list), progress has been
significantly slower than originally planned for. In large part, this
is because many upstream projects are involved: \Sphinx, \Docutils,
\Pygments, \Cython, and \Python. We are thus dependent on their own
independent schedule whenever an upstream contribution is needed;
an highlight of such sensitive contribution is PEP 580, a Python
Enhancement Proposal for the core language implementation (see \S~\ref{cython}). Such
a situation was foreseen in our Risk Management Plan and its impact on
the overall project is negligible: on the one hand, no other task
depends on \taskref{UI}{sage-sphinx}; on the other hand, it was easy to
shuffle around some of our internal work plan, and reserve PMs
for this task during OpenDreamKit's final year, at Ghent University
and Université Paris-Sud. Also, knowing that this was a long term
endeavour and that the timeline was not critical, we decided on
several occasions to invest in long term solutions that would benefit
not only \Sage but the community as a whole (see \S\,\ref{cython}).

We report here on the current achievements, and detail the work plan for the
final year. Formally speaking, this deliverable report concerns solely the
first goal of \taskref{UI}{sage-sphinx}; however, as the its goals are closely interrelated, reporting on
both is required to explain the big picture.

\section{Removing \Sage customisations}

\Sage uses a highly customised version of \Sphinx.
The word ``customised'' can mean any of the following:
an actual patch to the source code,
a monkey-patch (a patch at runtime),
a fork of some plug-in, or just a highly specialised configuration.

One of the goals of this deliverable is to remove these
customisations, allowing \Sage to use a standard \Sphinx.
This can be achieved in two ways:
either contributing back the customization to upstream \Sphinx to make
it a standard feature,
or changing \Sage (or some other project like \Cython or \Python)
to remove the need for the customisation.

This has multiple advantages:
first of all, having less custom code makes \Sage more maintainable.
Second the customization contributed upstream benefit everybody
and not just \Sage. Last, but not least, it helps packaging \Sage
(see \taskref{component-architecture}{mod-packaging}),
as packagers need not any produce and maintain a customized \Sphinx
package in addition to the standard one.

\subsection{General clean up and minor fixes}

We did quite a lot of minor clean up.
This is hard to list in detail.
In particular, various fixes were made to the build system for the documentation,
including using a more standard \Sphinx configuration, as well as
removing several irrelevant customisations to \texttt{autodoc},
the extension for generating documentation from docstrings.
We also made building the \Sage documentation a lot more reliable:
in the past, it often failed to build incrementally after an upgrade,
which required a resource expensive full rebuild. This was a burden
for developers but also a barrier for efficient continuous integration
(see \longdelivref{component-architecture}{multiplatform-buildbot}).

\subsection{Support for Cython functions}\label{cython}

\Sage uses a combination of Python and Cython modules.
Cython is a programming language which mostly uses the Python syntax,
but which is compiled to C instead of interpreted like Python.
By default, functions in Cython become Python \emph{built-in} functions.
These are a special kind of optimised function, implemented in C.

Unfortunately, built-in functions do not work well with \Sphinx, as they
do not have all the needed introspection capabilities.
In particular, \Sphinx cannot determine the input arguments of such functions.
Since the arguments are an important part of the documentation
of a function, this is a big problem.
\Sage and \Cython have various hacks to make it work anyway.
These are quite ugly, so it would not make much sense to support those in \Sphinx.

Fixing this properly requires adding support in \Python
for fast user-defined function classes.
For user-defined function classes, \Cython can add whatever attributes
are needed to support \Sphinx and other use cases that require deep introspection
of functions.
The problem is that such user-defined function classes
are necessarily slower than built-in functions.
This is not acceptable for \Sage where speed is very important.
In June 2018, we submitted
\href{https://www.python.org/dev/peps/pep-0580/}{PEP 580} (see Appendix~\ref{pep580}), a Python Enhancement Proposal to fix this.
At the May 2018 OpenDreamKit developers workshop in Cernay
(see~\longdelivref{dissem}{workshops-3}),
we already discussed an earlier variant of the PEP
with the core \Cython developers and they were very positive about it.
Both before and after submitting the PEP,
we discussed it online with the Python community,
leading to substantial improvements.
Unfortunately, the process of getting a PEP accepted is quite slow:
at the time of writing this report, the PEP has not been accepted (nor rejected).

If PEP 580 is accepted, \Cython can be updated to produce
functions which are equally fast as built-in functions
and which support all documentation
features that standard Python functions support.

\subsection{Parallel build of the reference manual as a single document}\label{pieces}\label{parallel}

The \Sage reference manual is currently not built as one monolithic document,
but in several pieces:
one for each mathematical topic such as combinatorics or matrices.
The original motivations were twofold: it makes it easy to build the various pieces
in parallel and it lowers the memory requirements.
Unfortunately, these pieces are not completely independent:
we still want to end up with one consistent document,
so things like the index and references need to merged.
This required a custom \Sage \texttt{multidocs} extension for \Sphinx.

In addition, the natural granularity of the pieces at the semantic
level never quite matched a proper granularity for parallelisation;
the existence of a few long chapters like on combinatorics led to
annoyingly long compilation.

Finally the parallelisation itself required a bespoke parallel
docbuilder, as \Sphinx did not have parallel build features at the
time. This is no more the case: \Sphinx now has native parallel
support, even within single documents.

Hence it is desirable to decouple parallelisation from documentation
splitting and work on memory requirements, to be able to build again
the reference manual as a single document.

At the OpenDreamKit organized Sage Days 77 workshop in April 2016
(see~\delivref{dissem}{workshops-3}), we had a coding sprint with with
Robert Lehmann, one of the main \Sphinx developers, where we
implemented a proof-of-concept showing the feasibility of this plan.
Once the memory requirements will be tackled (see \S\,\ref{resources})
it will be little work to finalize that plan.

\subsection{Custom syntax highlighting}

\Sage currently ships with a slightly patched version of \Pygments in
order to support highlighting of the \texttt{sage:} prompt in addition
to the standard Python \texttt{>>>} prompts. This is a burden for
packagers. We need to investigate whether we can configure \Pygments
without patching it, or push support for \Sage upstream to \Pygments.
Either way, this should be quick and straightforward.

\section{Upgrading \Sphinx}

In tandem with removing customisations, it is obvious that we should
use the most recent release of \Sphinx. This is not as trivial as it
sounds as we started with a very old version of \Sphinx in which the
custom code integrated tightly with \Sphinx. Thanks to an incremental
upgrade in a series of 6 tickets, \Sage is now using a reasonably
recent version of \Sphinx (1.7.6 as of now). Allowing for smoother
future upgrades is yet another reason for removing the customisations
mentioned in the previous sections.

\section{High resource usage}\label{resources}

Building the full \Sage documentation takes a lot of resources:
it needs about 3 gigabytes of memory and about 1.5 hours
on an averge personal computer to build the documentation from scratch.
This makes it the most resource-intensive part of
building the complete \Sage computational system.
Because of this, \Sage users who wish to build from source are often
suggested to skip building the documentation.
It is also a problem for automatic builds in continuous integration
setups, which typically only have limited resources.

During the aforementioned Sage Days 77 coding sprint with a \Sphinx
developer, we have investigated where the high memory usage comes from.
It turns out that it is a combination of factors
and that there is not a single obvious cause.
Some of it is caching of data which is no longer needed,
which should be easy to fix.
Some of it is data structures using basic Python primitives such as a \texttt{dict}
instead of more specialized (and space-efficient) structures.

The \Sage customizations to \Sphinx
(in particular what is mentioned in \S\,\ref{pieces} and \S\,\ref{parallel})
may cloud this analysis somewhat
as they might make things better or worse.
Therefore, it would make most sense to tackle this problem after
removing or at least reducing those customizations.

Fixing this will need to be tackled at the level
of \Docutils or \Sphinx.
Therefore, this will benefit all projects using \Sphinx.

\section{Related work}

It is beneficial for \Sage packages to use the same theme
and configuration for their documentation as \Sage.
We simplified the process by implementing the \verb/sage-package/ utility
(which contains among other things a copy of the Sphinx configuration for \Sage).
Its usage is illustrated in the \verb/sage_sample/
template for \Sage packages.
We are investigating how to avoid the aforementioned copy.
In any case, it is our goal to use as much as possible a standard \Sphinx setup,
so this should be limited to some basic configuration and style options.

In \longdelivref{UI}{ipython-kernels}, we report on
\href{https://github.com/minrk/thebelab}{ThebeLab}, a JupyterLab based
JavaScript library that turns static code examples in an HTML page
into live code cells that the user can edit and execute. We configured
\Sphinx to enable ThebeLab in the documentation of \Sage and \Sage
packages.

Finally, we mention two solutions to turn \Sphinx documentation into
\Jupyter notebooks: \ODK members, with some external contributions,
developed early on a stand-alone \verb/rst2ipynb/ program for
converting a single \verb/.rst/ file to a Jupyter notebook. More
recently a \Sphinx plugin \verb/sphinxcontrib-jupyter/ was developed
by the community, with some \ODK contributions. This plugin integrates
directly with \Sphinx, operating on a collection of documents instead
of a single document. As such, it supports crosslinks, which makes it
a lot more suitable for converting large documents such as books (see
\longdelivref{dissem}{ibook1}) or the Sage manuals. We are currently
in the process of adding support for this in the Sage docbuilder, see
\url{https://trac.sagemath.org/ticket/25909}

\appendix

\input{pep-0580.tex}

\end{document}
