\subsubsection{Work Package 1: Project Management}

%Explain, task per task, the work carried out in WP during the reporting period giving details of the work carried out by each beneficiary involved.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Overview}

\ednote{@nthiery, @IzabelaFaguet: update WP1 for RP3}

The general objectives of Work Package 1 are:

\begin{enumerate}
\item{Meeting the objectives of the project within the agreed budget and timeframe and carrying out control of the milestones and deliverables}
\item{Ensuring that all the risks jeopardising the success of the project are managed and that the final results are of high quality}
\item{Ensuring that the innovation process within the project is fully aligned with the objectives set up in the Grant agreement}
\end{enumerate}

\WPref{management} has been divided into three tasks. In the
following, we highlight the key achievement in this work package, when
necessary report specifically on these individual tasks.


\begin{itemize}
\item %\site{PS} led the 2nd, 3rd, 4th and 5th amendment to the consortium, in particular to cater for the moving to other institutions of permanent and/or non-permanent researchers who are key personnel for the success of \ODK. A collateral effect is the addition of the sites \site{Ghent}, \site{Leeds}, \site{FAU}, \site{XFEL}, and termination of the sites \site{UZH}, \site{USFD}, \site{JU}, \site{SOUTHAMPTON} since no relevant staff for \ODK remained at these institution. 

Since the beginning of the project, \site{PS} led  5 amendment processes to GA. 
During period 1 and 2, \site{PS} put forward 4 amendments; to manage several work 
plan revisions,  to reallocate staff and all remaining resources from the terminated 4 sites
 \site{UZH}, \site{USFD}, \site{JU}, \site{SOUTHAMPTON} to the added  sites \site{Ghent}, 
\site{Leeds}, \site{FAU}, \site{XFEL}.
\item During the third period, \site{PS} put forward a 5th amendment upon the request of \site{FAU}
to add a subcontractor to conduct a new task T6.11. Following the suggestions 
of the Project officer, \site{PS} formalized in this 5th amendment the budget transfers
between beneficiaries to optimize the use of remaining resources to achieve the project 
aims, this was made possible thantks to the significant unusued resources remaining at 
LEEDS due to the early departue of all the personnel. 

\item In earlier Reporting Periods, \site{PS} had organized the first
  project review in Brussels on April 2017, and steering committee
  meetings in Orsay (September 2015), St Andrews (January 2016),
  Edinburgh (January 2017), Brussels (March 2017), online (February
  2018), and at \site{XFEL} (June 2018).

  During Reporting Period 3, \site{PS} organized the second project
  review in Luxembourg on October 2018 and steering committee meetings
  in Luxembourg (October 2018) and Marseille (February 2019). It also
  organized a one week "report writing sprint" in Cernay (August 2019)
  to collectively write the project reports. The final review meeting
  will take place in Luxembourg on October 30, 2019, and will gather
  20 ODK participants to present the project final results.

\item \site{PS} made sure that all the milestones and deliverables of
  the third Reporting Period were achieved within its timeframe, and
  reported on in a timely manner.
\item Concerning communication, \site{PS} has been maintaining the
  internal communication tools that were described in
  \longdelivref{management}{infrastructure}. As for external
  communication the website for the project has been continuously
  updated with new content, and virtually all work in progress is
  openly accessible on the Internet to external experts and
  contributors (for example through open source software repositories
  on Github).
  % A new version of the website was released in June 2018.
  % Its end-user friendly interface and content makes it a tool not only
  % for internal communication but very much for dissemination and
  % progress tracking by the reviewers and the community.
\item Concerning the future of \ODK and of its infrastructure toolkit,
  the consortium kept accessing information and getting involved in
  the development of the European Open Science Cloud that is currently
  promoted by the European Commission. The project manager
  participated to \ednote{Izabela: a word on the two workshop you
    participated to}.

  In addition, two spin-off proposals were submitted on January 29th
  to the H2020 European E-Infrastructure call INFRAEOSC-02-2019:
  \begin{itemize}
  \item \href{https://github.com/bossee-project/proposal}{BOSSEE}:
    Building Open Science Services on European E-Infrastructure, with
    a focus on Jupyter and applications;
  \item \href{https://opendreamkit.org/2019/01/29/FAIRmat/}{FAIRMAT}:
    FAIR Mathematical Data for the European Open Science Cloud.
  \end{itemize}
  Both were prepared with the same open strategy as OpenDreamKit, and
  involved a combination of OpenDreamKit and external partners.

  % We took advantage of
  % the EOSC stakeholder forum on 28-29 November and the 2017 edition of
  % the DI4R (Digital Infrastructures for Research) in Brussels. During
  % these events we gathered information on the potential of EOSC and
  % how \ODK could fit in there. Furthermore we initiated a partnership
  % with EGI -- a key participant to EOSC -- to deploy \ODK based
  % infrastructure.
\item \site{PS} organized the continued interaction with the Advisory
  Board and the Quality Review Board to control the quality and the
  relevance of the software development relative to the end-user
  needs. See below.
\item \site{PS} produced a second version of the Innovation Management Plan
  %(\delivref{management}{data-plan2}).
 to assure that research activities meet the required milestones and produce outputs fully aligned with the project objectives.

\item As in the previous reporting periods we observed that the
  strategies we used (tailoring of the positions according to the
  known pool of potential candidates, in particular among previous
  related projects, strong advertisement, ...) have paid off, and we
  are really happy with the top notch quality of our recruits. In
  addition, we recruited three women near the beginning of period 3:
  Katja Bercic for a postdoc position at \site{FAU}, Odile Benassy for
  a research engineer position and Izabela Faguet who replaced Benoît
  Pilorget as project manager at \site{PS}. Even if very far from
  achieving gender equality, this is a step forward given the very
  small pool of experienced research software engineers for
  mathematics on temporary positions and its high gender imbalance.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph{Tasks}

\subparagraph{\longtaskref{management}{project-finance-management}}

As planned in WP1, \site{PS} has been coordinating \ODK and took care
of the budget management together with the administration body, the
D.A.R.I. (Direction des Activités de Recherche et de l'Innovation) and
its finance service. This included prefinancing and funds transfer to
cater for the moving of personnel across sites and from old sites to
new sites.


\subparagraph{\longtaskref{management}{project-quality-management}}


\begin{enumerate}
\item Continued success in the recruitment of highly qualified staff.
\end{enumerate}

The Quality Assurance Plan is described in detail in
\longdelivref{management}{ipr}. We will describe the main points
below. \site{PS} had launched during RP1 a Quality Review Board which
is chaired by Hans Fangohr. The four members of the board have a track
record of caring about the quality of software in computational
science. This board is responsible for ensuring key deliverables do
reach their original goal and that best practice is followed in the
writing process as well as in the innovation production process. The
board meets after the end of each Reporting Period (RP), and before
the Review following that RP. More details about the Quality Review
Board are given in Section~\ref{section.QAP}.

% More information on
% \longtaskref{management}{project-quality-management} can be found in
% Section 4 of this document: Quality assurance plan.

The other structure supporting \ODK to ensure the quality of the
infrastructure it produce is the Advisory Board. It is composed of
seven members, some of which specifically represent End Users:

\begin{itemize}
\item{Lorena Barba from the George Washington University}
\item{Jacques Carette from the McMaster University}
\item{Istvan Csabai from the Eötvös University Budapest}
\item{Françoise Genova from the Observatoire de Strasbourg}
\item{Konrad Hinsen from the Centre de Biophysique Moléculaire}
\item{William Stein, CEO of SageMath Inc.}
\item{Paul Zimmermann from the INRIA}
\end{itemize}

This Advisory Board is composed of academics and/or software
developers from different backgrounds, countries and communities. It
is a strong asset to understand the needs of a variety of end-user
profiles. After the first reporting period, \site{PS} led the review
by the Advisory Board of our Technical Report and later Work Plan
Revision Proposal. This feedback, complementing that of the reviewers,
was very helpful to guide the final work plan revisions.

\site{PS} has also been managing risks. Up to the Leeds situation, the
assessment we present in Section~\ref{} has only marginally deviated
from earlier assessments in at Month 12 in \delivref{management}{ipr}
and Month 36 in \delivref{management}{ipr2}.

\subparagraph{\longtaskref{management}{project-innovation-management}}

\longdelivref{management}{imp1} was produced at month 18, with a focus on:

\begin{itemize}
\item{The open source aspect of the innovation produced within \ODK}
\item{The various implementation processes the project is dealing with}
\item{The strategy to match end-users needs with the promoted VRE}.
\end{itemize}

The second version of the Innovation Management Plan in RP3 will add content
to explain all the innovations that the VRE is bringing to end-users.
However the open source approach and the "by users for users"
development process will not change.

One of the assessed risks for
\ODK is to have different groups not forming effective teams. Put in
other words, having developers of the different pieces of software
working solely for the benefit of the programme they were initially
working on and for. This risk was continuously tackled by the
Coordinator in order to reach the final goals of the VRE which are the
unification of open source tools with overlapping functionality, the
simplification of the tools for end-users without coding expertise,
and the development of user-friendly interfaces. For this, the
Scientific Coordinator is for example wilfully pushing for joint
actions and workshops.

The evolution of the social organization of software development
communities with a long history is by nature a slow process;
nevertheless, ever since the beginning of the project, collaborative
efforts across communities are progressively becoming a standard
practice, and not only for \ODK participants. More information on
joint workshops can be found in the next section.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:

%  LocalWords:  subsubsection longtaskref delivref organized Bougeret ipr Csabai
