\subsection{Generated System API Theories}\label{sec:sysapis:gentheories}

We have generated and cross-linked system API theories for the computer algebra systems
GAP, SageMath, and SINGUALAR, as well as the mathematical knowledge base LMFDB. They are
generated from pre-existing sources in the systems like type information, code
annotations, and systematic API documentation.  OpenDreamKit deliverable report, Section
5 of \textbf{D6.5}~\cite{ODK-D6.5} has the details.  


The OpenDreamKit System API Theories can be found at
\url{https://gl.mathhub.info/ODK/{GAP,SAGE,SINGULAR,LMFDB,knowls}}. These contain:
\begin{itemize}
\item \texttt{GAP}: (210 Theories, 8470 Symbols\footnote{Since the generated API theories
    do not come from \texttt{.mmt} source files, numbers of files and LoF are not
    informative metrics. Approximately, one \texttt{.mmt} file contains on average 2--3
    theories, and 1--2 LoF correspond to one symbol.}, 64 Commits) Generated from a JSON
  export from the typeo system and documentation of the GAP system (after adding
  constructor annotations throughout the GAP source code). A theory corresponds to a
  source file of a GAP package, a symbol represents a GAP method or operation.
\item \texttt{Sage}: (5431 Theories, 7279 Symbols, 73 Commits) Analogously generated from
  a JSON export. Theories correspond to Sage categories, symbols to methods.
\item \texttt{Singular}: (179 Theories, 4519 Constants, 6 Commits) generate by
  heuristically parsing definitions, call patterns, and comment strings in the Singular C
  code. 
\item \texttt{LMFDB}: (161 Commits) Consists of schema theories for LMFDB databases
  (currently 5 databases covered, 182LoF) and interface theories for non-mathematical or
  LMFDB-specific concepts (labels, descriptors) represented therein (3 Files, 123
  LoF). See Section 6 of \textbf{D6.5}~\cite{ODK-D6.5} for a discussion of schema theories
  and the content of this archive. 
\item \texttt{knowls}: (945 flexiformal Theories, ca. 900 Symbols; 10 Commits) generated
  from the LMFDB knowls system of interface documentations. The knowls resource is
  actively maintained and currently under intense review by the LMFDB community. 
\end{itemize}

\subsection{The MitM Alignments}

The MitM alignments are available from \url{https://gl.mathhub.info/alignments/Public} (6
files, 1040 LoF, 240 Commits). They are represented as text files containing pairs of MMT
URIs annotated with various semantic classification schemata;
see~\cite{MueGauKal:cacfms17} for details.\ednote{MK: what else can we write about them?}
\ednote{this figure does not fit in the text width}
\begin{figure}[ht]\centering
  \tikzinput[width=.98\textwidth]{../D6.5/alignmentimg}
  \caption{Alignments between the MitM Ontology and the \GAP API}\label{fig:cgtontology}
\end{figure}

As an example, Figure~\ref{fig:cgtontology} sketches the alignments between the
computational group theory ontology from above and the constructors and operations of
\GAP.

%%% Local Variables:
%%% mode: visual-line
%%% fill-column: 5000
%%% mode: latex
%%% TeX-master: "report"
%%% End:

%  LocalWords:  textbf texttt ednote knowls flexiformal MueGauKal:cacfms17 centering
%  LocalWords:  tikzinput textwidth fig:cgtontology
