\begin{event}{Sage Days 70}{sd70}{Berkeley (US California), 2015-11-08 to 2015-11-14}{PS}{16}{https://wiki.sagemath.org/days70}

\textbf{Main goals.} Gather developers from Sage, SageMathCloud and Jupyter together to learn the
inner machineries of the different projects and code together towards common goals.

\textbf{ODK implication.} ODK was not the main organizer of this event, it was used to fund
European participants which are either part of ODK or very closely related.

\textbf{Event summary.} The event featured many interesting talks on the inner mechanics of
both SageMathCloud and Jupyter, in particular:
\begin{itemize}
\item \href{https://youtu.be/GOuy07Kift4}{How to contribute to SageMathCloud} by William Stein
\item \emph{The PARI Jupyter kernel} by Jeroen Demeyer
\item \emph{Jupyter Notebook development} by Jason Grout.
\end{itemize}
Lots of time was devoted to projects and code such as: installing a development version of SageMathCloud,
following tutorials on SageMathCloud development, working toward the integration of the Jupyter notebook
in Sage.

Furthermore a Jupyter interface for HPC-GAP was developed, and the Jupyter
interface for GAP was improved. A talk \emph{The current status of (HPC-)GAP}
was contributed.

\textbf{Results and impact.} 
This workshop was essential to some ODK planned tasks. This was especially related to WP3 and WP4. Here are some tasks that
were started during the Sage Days:
\begin{itemize}
\item T3.6 Document and modularize SageMathCloud's codebase. This task was started during the workshop using the 
knowledge of the main developer of SageMathClod, William Stein.

\item T4.1 Uniform notebook interface for all interactive components. This is a major task of WP4. This workshop
was an occasion to share first hands information between Sage, GAP, and Jupyter developers.

\end{itemize}
The knowledge we gathered during presentations was relevant to all tasks including notebook interfaces and cloud
systems.

\end{event}