\begin{event}{Sage Days 73 May}{SD73}{Oaxaca, Mexico, 04 - 07 May 2016}{CNRS}{9}{https://wiki.sagemath.org/days73}

\textbf{Main goals.}
This Sage workshop was a satellite of the conference Flat Surfaces and
Dynamics of Moduli Space that happened in Oaxaca May 08-13. The aim was to
introduce participants to SageMath and share code and knowledge.

On the first day, we also had two participants from the University of 
Oaxaca.

\textbf{ODK implication.} ODK, via its Bordeaux node, supported the expenses
of participants. Vincent Delecroix made introductory and advanced talks about
SageMath and Python.
  
\textbf{Event summary.} The first day was dedicated to a SageMath introduction.
Each day in the afternoon, we had a demonstration from a participant. The rest
of the time was dedicated to programming.

\textbf{Demographic.}
11 persons took part in these Sage Days: 2 females and 9 males, originating
from the following countries: Mexico (3), Canada (3), France (3), USA (2)

\textbf{Results and impact.}
\begin{itemize}
\item A step toward the convergence of the IPython and SageMath notebooks with \url{https://trac.sagemath.org/ticket/20562} (REFERENCE TO WORKPACKAGE)
\item A fix in SageMath for a problem discovered Maxime Fortier-Bourque during the workshop (\url{https://trac.sagemath.org/ticket/20566})
\item Charles Fougeron's code about Lyapunov exponents gets integrated in \url{https://github.com/videlec/sage-flatsurf}.
\item New visualization tools for geometry of translation surfaces at \url{https://github.com/videlec/flatsurf-package}.
\item And several experimentations by the other participants.
\end{event}
