\begin{event}{13th Joint Magnetism and Magnetic Materials (MMM) -
    Intermag Conference}{mmm2016}{San Diego (CA, USA), 2016-01-11 to
    2016-01-15}{SOUTHAMPTON}{30}{https://magnetism.org}

  \textbf{Main goals.} The main goals of presenting the project at
  this important international Joint Magnetism and Magnetic Materials
  and Intermag conference was to introduce the Micromagnetic Virtual
  Reasearch Environment (VRE) to our target user audience - the
  micromagnetic scientific community.


\textbf{ODK implication.} The OpenDreamKit project has sent the
speaker (Hans Fangohr), and paid expenses for the trip.

\textbf{Event summary.} Hans Fangohr submitted a talk [1] that was peer
reviewed and accepted for presentation. In the talk, he outlined the
vision for the project and invited feedback from the community.

The ODK project for computational micromagnetics was discussed with
various attendees informally throughout the conference.

In addition, we organised a meeting with the main developers of the
OOMMF micromagnetic simulation code, Dr M. Donahue and
Dr D. Porter, in order to discuss the project's vision, our plans for
interfaces to get early feedback.

\textbf{Demographic.} About 30 people were present, but no demographic survey was done.

\textbf{Results and impact.} We announced the project and its website
to the community and encouraged input to extend our vision, to make
sure the tool we develop can be as practical and efficient for as
large parts of the community as possible.

[1]
\href{http://joommf.github.io/assets/2016-01-12-MMM2016-AD02-Jupyter-OOMMF.pdf}{pdf
of slides}
\end{event}
