\begin{event}{Conference Combinatoire, Algèbre et Théorie des Nombres}{2016CATN}{Monastir, Tunisia -- 24-28 Mars 2016}{UB}{30}{http://www.edsf.fss.rnu.tn/Colloque1/colloque3.html}

\textbf{Main goals.}
CTAN 2016 was a mathematical conference that took place in Tunisia.
The first goal of ODK was to make an initiation on SageMath. 
The second goal was to create a Sage developer team in Tunisia.

\textbf{ODK implication.} A. Boussicault was sent by ODK to deliver training sessions on
SageMath. ODK paid the travel and accommodation of A. Boussicault.

\textbf{Event summary.}
The event wasone-week long, with a 2-hour long intervention per day. 
Trainings were given on participants' computers.
We worked with Sage by using a Debian live USB stick.

A Linux installfest on participants' laptops took place every evening. 
The evening sessions were beginning at 20:30 and ended at midnight.

The session was conducted by A. Boussicault (University of Bordeaux) and 
Imad Eddine Bousbaa (University of USTHB of Algeria), who was met during a 
previous conference in Algeria (\ref{event-2015DIMACOS}).
The purpose was to present and make mathematical calculus with Sage.
We also gave classes on Python and on Combinatorics.

During that event, we discussed with CTAN organizers and some researchers 
to organize some important mathematical events in Tunisia, Algeria, Morocco and
Lebanon.
Many researchers were present in the discussion. For example, there were 
Professor H. Belbachir (University USTHB in Alger), 
Professor I. Boudabbous (University of Sfax in Tunisia)
Professor O. Khadir (University Hassan II of Casablanca),
Professor M. Pouzet (University of Lyon1),
professor H. Kheddouci, (University of Lyon1).
We decided to organize, in the next year, an event with the mathematical 
conference CTAN followed by mathematical schools and some Sage Days.
In the school, we could give some Sage classes and we could implement ideas 
and mathematical tools during the Sage Days.

\textbf{Demographic.} 30 participants. CTAN being an international conference, participants
 came from Lebanon, Algeria, Tunisia, Morocco, etc. 

\textbf{Results and impact.}
The will to organize an event with a mathematical conference 
followed by mathematical schools, and Sage Days in Morocco or in Algeria.

We made some Tunisian contacts to create a developer team in University of 
Sfax.

\end{event}
