\documentclass{deliverablereport}

\usepackage[style=alphabetic,backend=bibtex]{biblatex}
\addbibresource{report.bib}
\addbibresource{../../lib/publications.bib}

\usepackage{xparse}
\usepackage{etoolbox}
\usepackage{caption}
\graphicspath{{events/}}
 \ExplSyntaxOn

\newcounter{eventcounter}

\newenvironment{event}[7]{
\vspace{0.5cm}
\refstepcounter{eventcounter}
\label{event-#2}


\noindent\textbf{Event~\theeventcounter -~ #1}\newline % title

\noindent #3 \newline % location and date

\noindent ODK~partners~involved:~ \clist_map_inline:nn{#4}{\site{##1}~}\newline %partners

\ifx&#5&%
      % no participant #
\else
\noindent #5~participants~
\ifx&#6 &%
    % no odk participant
\else 
~(including~#6~from~within~ODK)\newline
\fi
\fi

\ifx&#7&%
      % no website
\else
\noindent \url{#7}\newline
\fi



}{\begin{center}\noindent\rule{4cm}{0.4pt}\end{center}}

 \ExplSyntaxOff


\deliverable{dissem}{workshops-3}
\duedate{31/08/2018 (M36)}
\deliverydate{31/08/2018}
\author{Viviane Pons et al.}

\begin{document}
\enlargethispage{4ex}
\maketitle
\githubissuedescription
\tableofcontents
\newpage


\section{Development workshops}

We call a development workshop an event with a restricted number of participants
who meet to work on a specific task. These workshops are an inherent part
of \ODK development process as described in \taskref{dissem}{devel-workshops}:
 they bring together
developers from within and outside of \ODK and allow effective work
and discussions on many technical aspects. They also participate in building
and maintaining a community of developers inside \ODK and within the
open-source communities we belong to.

Throughout years 2 and 3 of the project, we have had 15 workshops dedicated mostly
to development. Some of them also included a training approach. 

\subsection{Atelier PARI-GP}

The PARI/GP Ateliers were established in 2012 as a yearly meeting
between developers and users of the PARI/GP system.

The main goals are advertising new features and improvements,
discussing further developments, sharing best practices, and collaborative
code writing (hacking sessions, doc reviews, bug-squashing parties).

You can find the list of previous PARI Ateliers at
\url{http://pari.math.u-bordeaux.fr/ateliers.html}.

We have listed here the three Ateliers which focused more on development. 
Other Ateliers can be found under the Dissemination section.

\input{events/2017-01-Atelier-PARI-GP}

\input{events/2017-06-PARI-GP}

\input{events/2018-01-Atelier-PARI-GP}

\subsection{Linbox}

The LinBox developer meeting are meant to gather developers and users of the
LinBox ecosystem (composed of \texttt{givaro}, \texttt{fflas-ffpack} and
\texttt{LinBox}).

The main goals are to enable interactions between the several groups of
developers and users in order to share experiences, take collegiate design
decisions and practice collaborative code writing.

\input{events/2017-12-linbox}

\input{events/2018-06-linbox}

\subsection{Software interactions}

As the project evolves, we are prone to have more and more events which 
concern many software projects altogether. All the 10 events below
included members from various communities to work on interactions and
general cross-projects technical questions. This included two special
workshops at ICMS 2018 as well as events at CICM 2017 and 2018.

\input{events/2017-03-13-SD85}

\input{events/2017-07-17-cicm}

\input{events/2017-10-23-live-documents}

\input{events/2018-04-30-Cernay}

\input{events/2018-06-04-jupyter-gap}

\input{events/2018-07-13-tetrapod}

\input{events/2018-07-24-icms-mitm}

\input{events/2018-07-24-icms-partbacktrack}

\input{events/2018-08-13-cicm}

\input{events/2018-08-17-cicm-caat}

\subsection{Other}

\input{events/2018-06-PM}

\section{Dissemination and outreaching activities}

We describe here all activities related to \taskref{dissem}{dissemination}:
these are all events oriented towards dissemination, training, and outreach. This
includes events organized or co-organized by \ODK and also
participating in external events and many communication activities.

\subsection{Training workshops}

Training the community to use the tools developed by \ODK has become a big 
part of our work. By organizing specific workshops, or events inside bigger
conferences, \ODK members have been able to share \ODK technologies such as \Jupyter,
\Sage, \GAP, PARI-GP with more than 500 end users.

\input{events/2016-10-CoDiMa-training}

\input{events/2017-01-Edinburgh-Jupyter}

\input{events/2017-04-04-IOP-Magnetism}

\input{events/2017-04-24-Intermag}

\input{events/2017-11-06-MMM2017}

\input{events/2017-11-06-sheff-BAD}

\input{events/2018-02-03-AIM}

\input{events/2018-04-12-sheff-hpc}

\input{events/2018-04-Atelier-PARI-GP}

\input{events/2018-05-21-MathExp}

\input{events/2018-05-sheff-webdata}

\input{events/2018-06-14-leeds-training}

\input{events/2018-07-15-ICM}

\input{events/2018-07-PGTC}

\subsection{Organization of Sage Days in established mathematical communities}

One goal of \ODK is to support local communities of researchers
and developers who contribute to the open-source software related to
the project. For \Sage, this means supporting the organization of Sage-Days
workshops that arise from within all the different mathematical communities. The main 
goal of these workshops is mostly to improve the Sage coverage of some mathematical
area. They also play a major role in training and communication. The
impact for \ODK can be summarized this way:

\begin{itemize}
\item \textbf{Making \ODK known to the end users}: by supporting Sage Days,
\ODK makes itself known to the Sage community and can
thus share the many developments of the project.

\item \textbf{Improving the overall quality of Sage}: by fostering researchers
in specific areas, Sage Days help bring interesting mathematics into
the software, which is beneficial for Sage and so \ODK.

\item \textbf{Training, bringing more user}: Sage Days are the perfect place
for new comers, especially students, to get their first experience with the software.

\item \textbf{Fostering a community}: Sage Days are helping making Sage a vibrant
community, which is vital for the success of \ODK.
\end{itemize}


\input{events/2016-11-SD79}

\input{events/2017-02-27-SageDays84}

\input{events/2017-04-17-sd86}

\input{events/2018-02-19-SageDays93}

\input{events/2018-07-ICERM}



\subsection{Training activities in developing countries}

As open-source software developers, we wish our products
to be accessible to as many people as possible. Even though we offer
 a free access, there is still a technical gap in many 
developing countries that 
often prevents schools and researchers to benefit from our softwares.
This is why we believe the role of \ODK is to foster 
a wider community that does not leave a part of the world behind. In 
this section, we describe training activities that have been conducted 
through \ODK in this regard.

\input{events/2017-05-sage-days-89}

\input{events/2017-11-Atelier-PARI-GP}

\input{events/2018-01-22-STM-Morelia}

\input{events/2018-06-ECCO}

\subsection{Women in \ODK}

\ODK is aware of the gender gap that exists in science in general
and more specifically in software development. We have been organizing
events to support specifically women developers, engineers and scientists.

\input{events/2017-01-SD82}

\input{events/2017-sheff-codefirstgirls}

\input{events/2017-11-29-sheff-DISCU}

\input{events/2017-PyLadies}

\subsection{Communication and participation to external events}

Dissemination activities also include the participation of \ODK
members to many different conferences of various size and topics
in computer science, mathematics, physics, and more. The goal is
to reach potential end-users, build bridges between communities and stay aware 
of current development in the scientific community.

We list here major events and communication. 

\input{events/2017-07-27-PASCO}

\input{events/2017-09-EuroComb}

\input{events/2017-09-07-sheff-rseconf}

\input{events/2017-10-20-Netmath}

\input{events/2017-10-23-live-documents}

\input{events/2018-05-sheff-pycon}

\input{events/2018-06-05-SPLS}

\input{events/2018-06-15-leeds-rse}

\input{events/2018-06-EJC-TDN}

\input{events/2018-06-ZETAS}

\input{events/2018-07-01-AMI-CoSO}


\section{Upcoming events and plans for the future}

As we reach the final year of the project, we continue our many events
policy to keep engaging the scientific community. By doing so,
not only do we train new researchers but we defend the computational aspects
of mathematics and the open source spirit which is essential to \ODK. 
Some events to look forward to:
\begin{itemize}
\item \textbf{Software Tools for Mathematics} is held in Koper (Slovenia). Confirmed
instructors include Alexander Konovalov for \GAP and Samuel Lelièvre for \Sage. This 
school is a direct consequence of event~\ref{event-STM_2018_01_Morelia} in Morelia.
\item \textbf{PyConFR} is in Lille in October and Viviane Pons will be a keynote speaker.
\item The conference \textbf{Free Computational Mathematics} is to be a major dissemination
event to promote the use of open source software in mathematics. It is organized solely by \ODK
and will be held in CIRM (Marseille, France) from Feb. 11 to Feb. 15, 2019.
\item More \textbf{development workshops} will be organized as the project goes and we get close
to the finish line.
\item Another \textbf{Women in Sage} will be organized during the summer 2019.
\end{itemize}


\newpage\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
