\TOWRITE{ALL}{Proofread WP 2 Dissemination pass 2}

% This work package must start in month 1. Not sure what the
% appropriate load factor should be, and why that is important?
\begin{workpackage}[id=dissem,wphases=0-48!.5,
  short={Community Building/Dissemination},
  title={Community Building, Training, Dissemination, Exploitation, and Outreach},
  lead=PS,
  PSRM=9,
  PSRM=9,
  SARM=18,
  USORM=1, %16 in total
  XFELRM=21, % moved 6PM from WP7 (15 + 6 = 21)
  USHRM=31,  % moved 24 PM from WP7 (22+24 = 46)
  USRM=30,
  UVRM=2,
  UKRM=2,
  UBRM=19,
  SRRM=2,
  LLRM=6,
  UGRM=1,
  LEEDSRM=15 %real PM is 14.67
]

\begin{wpobjectives}
  The objective of this work package is to further develop the community at the
  European scale, foster cross team collaboration, spread the
  expertise, and engage the greater community to participate in the
  definition and refinement of the requirements, and the implementation and use of the
  produced solutions. This includes:
  \begin{compactitem}
  \item ensuring awareness of the results in the user community;
  \item engaging cross communities discussions to foster scientific collaboration and conjoint development;
  \item spreading the expertise through workshops and training sessions;
  \item providing training for partners inside the project, the
	community engaging with contributions to the project, and
	end-users of \TheProject
  \item reviewing emerging technologies,
  \item develop demonstrators,
  \item defining individual exploitation plans; and,
  \item managing existing and new intellectual property.
  \end{compactitem}
\end{wpobjectives}

\begin{wpdescription}
  We will organise regular open workshops (e.g. Sage Days, PARI Days,
  summer schools, etc.); some of them will be focused on development
  and coding sprints, and others on training. This is also an occasion
  to organise cross community workshops like Sage-Jupyter days.

  Some of the networking activities and internal training will come
  from short- to long-term mutual visits by participants, to
  collaborate on specific features. A typical visit would bring
  together a \Jupyter developer with a GAP developer for a couple of
  days to implement a first prototype of a notebook interface to GAP.

  A number of demonstrators will be developed in the project and
  disseminated in different ways.

  We will also participate in the concertation activities,
  consultations and other meetings and events of the European
  E-Infrastructure projects.

  All the code, documents, test and build infrastructure produced as
  part of the project will be made available as open source.
  Open access to all publications resulting from the project will be ensured.

  This work package will complement and lean on a parallel COST
  network which role is to build and engage a greater community.
\end{wpdescription}

\begin{tasklist}

\begin{task}[title=Dissemination and Communication activities, lead=PS, partners={SA}, id=dissemination-communication, PM=11, wphases=0-48, issue=24 ]
  This task comprises all forms of direct dissemination and public
  communication activities such as press releases, creation of the
  project web-site (\delivref{management}{infrastructure}) including visitor analysis and monitoring tools,
  scientific and technical publications, outreach activities
  (seminars, keynote talks, media interviews, press releases),
  promotion through social media (e.g. Twitter, Facebook, LinkedIn),
  creation of advertisement materials such as flyers, posters, and
  electronic feeds as well as their distribution. We will use standard
  community building technology such as mailing lists, Wikis and
  Forums, to ensure dissemination and engagement of the community to
  support this. We will also generate press releases at appropriate
  moments (\delivref{dissem}{press-release-1}, \delivref{dissem}{press-release-2}). %, making use of the public relation support services in the respective institutions.
  %At least two press releases will be
  %generated in the course of the project.


  % News articles will be produced by experienced professional staff
  % at relevant partners including ... and communicated to local,
  % national and international media, as appropriate.
\end{task}

\begin{task}[title=Training and training portal,
id=training-portal,lead=PS,PM=1,wphases={0-48!.1,0-1},issue=25]
Training is at the heart of this project: through our open source
approach, networking activities, workshops, demonstrators
(\localtaskref{index-librorum-salvificorum}), interactive books (such
as in \taskref{UI}{oommf-tutorial-and-documentation} and
\localtaskref{ibook}), and training for teachers and trainers
(\localtaskref{project-intro}), we have firmly integrated the training
aspect into the core of our project plan.

Each of these activities will create dedicated webpages hosting the
material to make it accessible to public as wide as possible ---
in line with our philosophy for software, we believe in the benefits of
sharing the material and maximising the value of the financial
investment into this project.

In this task, we create a central \TheProject training portal that
serves as an inclusive point of entry to explore available training
materials. This will be hosted on the projects website (\localtaskref{dissemination-communication}).
\end{task}

\begin{task}[title=Community Building: Development Workshops, lead=PS,PM=24, partners={UB,UK,SR,SA,LEEDS,UG}, id=devel-workshops, wphases=0-48, issue=26]
  We will organise development workshops all throughout the
  project. The aim of these workshops is to bring together developers
  from the different communities to design and implement some key
  aspects of \TheProject such as user interface, and documentation and
  to ensure cross compatibility. These meetings will gather not only
  participants of \TheProject but also members of the different
  communities involved. Bringing talented people together is the best
  way to make actual progress on the different aspects of \TheProject
  and to kick-start new challenges. It is also a way to work within
  the communities we're reaching in and include them in the discussions
  and development. It fosters collaboration between scientists and
  developers from different backgrounds to build tools that are needed
  by all.

  Each workshop will be aimed at a specific software component (\Sage,
  \GAP, \SMC, \IPython, \Singular, etc.) or be joint meeting between
  different communities to improve interoperability and joint
  developments. We are planning to have 4 or 5 such events per
  year. The currently anticipated list of workshops includes:

\begin{compactitem}
\item One \Sage workshop per year in Cernay France (near Orsay) where
  similar gatherings took place before. One of them will be a
  Sage-Sphinx day, dedicated to documentation.

\item One Atelier \Pari in Bordeaux per year. The team in Bordeaux has
  a great experience in organizing this kind of \Pari events.

\item Two \Singular workshops and two \GAP-\Singular workshops in Kaiserslautern
  over the four years.

\item Two workshops dedicated to high performance mathematical
  computing in relations with \WPref{hpc}.

\item Two Data Science workshops which use \TheProject to develop effective machine learning and data modelling practice organised by Sheffield.

\item A joint meeting on the topics of \SMC and \Jupyter in Simula in
  relation with \WPref{UI}.

\item A joint event between \GAP, \Sage, and \Singular in ICMS,
  Edinburgh.

\item A joint \Jupyter and \Sage event in Orsay.

\item A joint \LMFDB and \Sage event in Warwick to work towards
  \WPref{dksbases}.

\end{compactitem}

Yearly reports on the impact of these workshops on the community (\delivref{dissem}{workshops-1}, \delivref{dissem}{workshops-4}) will be delivered.
\end{task}


\begin{task}[title=Reviewing emerging technologies, id=tech-review, lead=PS, partners={SA,USO,LEEDS,US,UV,UB,SR,XFEL},PM=10, wphases=0-48,issue=27]
  In this task, we will produce periodic reviews (\delivref{dissem}{techno}) of emerging
  technologies and relevant developments elsewhere, and implications
  for our plans, taking into account input from the communities. This
  include the review of standard components and service for storage
  and sharing, computational resources, authentication, package
  management, etc. This may further include negotiating access or
  shared development when appropriate. This information will be fed to
  the other work packages, in particular Work
  Package~\WPref{component-architecture} Component Architecture.
\end{task}


\begin{task}[title=Dissemination: reaching towards users and fostering diversity, lead=PS,PM=12, partners={UB,LEEDS,SA}, id=dissemination, wphases=0-48, issue=28]
  As lead developers of \TheProject, most of us consider themselves as
  both scientists and developers. We have experience in reducing the
  gap between those two worlds. We organise training workshops such
  as Sage-days to promote our tools and bring more users and
  developers from the scientific world. On the other hand, we often
  attend and present to more development-oriented gatherings like
  PyCon and SciPy to interact with engineers and foster
  collaboration.

  The aim of this task is to exploit this winning strategy within
  \TheProject. Three events should be organised in the spirit of
  Sage-days to gather and train more users and foster scientific
  development around \TheProject. These conferences usually welcome
  around 50 participants and have a big impact on the scientific
  community. One of them will be at CIRM in Marseille, another one at
  ICMS in Edinburgh and a third one probably in Dagstuhl, Germany. In
  the same spirit, we will also have training sessions organised
  within the universities (Orsay and Grenoble). 
  We will also run a series of workshops and trainings in developed as well as 
  in developing countries, especially in Africa and
  South America. One of the aims of these events will be to train young researchers
  to tools developed by  \TheProject. The events in developping countries will
  be connected to CIMPA schools.  The CIMPA is an international organisation based in Nice
  (France) that promotes research in mathematics in developing
  countries. It organises each year around 20 schools.

  The under-representation of women in the scientific world is even
  more visible if we intersect science with software
  development. As we know, we have many talented women in our
  community, and we will organise some events targeted at women in the
  spirit of the "Women in Sage" days that happened many times in the
  US already. We are planning to have two of them in Orsay and at least
  one in Oxford where "Women in CS" days already take place on a regular basis.

  Apart from these different events, we will also be present at
  important events of both our scientific community (international
  mathematical conferences such as FPSAC for combinatorics) and the
  python / open-source software development community: PyCon, SciPy,
  EuroPython, etc. The material we develop for presentation at these
  events will be made publicly available.
\end{task}


%Mike Croucher and Neil Lawrence,Sheffield
\begin{task}[title=Introduce \TheProject to Researchers and Teachers, id=project-intro,lead=LEEDS,wphases=6-44,PM=20,partners={USO,XFEL}, issue=29]
 
In this task, we will develop and deliver materials that will introduce OpenDreamKit to potential users—both researchers
and teachers. Our initial focus will be on teachers,the materials will be deployed to researchers, both local to the
University of Sheffield and across the wider computational biology and machine learning fields.

  We will develop a `taster' seminar (1-2 hours) and follow-up short course
  (1-2 days) on \TheProject for researchers and lecturers in all
  disciplines \delivref{dissem}{short-course}. At Sheffield, this will
  be added to the set of courses that are offered as part of IT
  Services' research support department. As such, it could potentially
  reach all disciplines. It will also be made publicly available for
  widespread dissemination and collaborative modification.

  Elements of this work will also be integrated with the GP Summer
  Schools and Roadshows (\url{http://ml.dcs.shef.ac.uk/gpss/}). The
  Summer School is now in its fourth edition (over 140 students
  educated). The Roadshows have taken place in Uganda, Colombia and in
  2015 they are scheduled for Italy, Australia and Kenya. The Kenya
  school will be the first to have more of a `data science' focus that
  we think will be particularly appropriate for dissemination of
  \TheProject (\delivref{dissem}{IntroODK}).

  These seminars and short courses will also be used to identify
  potential collaborators who are interested in utilising \TheProject
  immediately. We will act as consultants to these collaborators in
  two ways:

  We will work with lecturers at Sheffield to introduce \TheProject to
  various disciplines via the production of interactive lecture notes
  (\delivref{dissem}{IntroODK}). The focus for the student here
  will not necessarily be on programming but rather on interaction
  with the subject matter via use of \TheProject. Interactive lecture
  notes are an area where commercial vendors such as MapleSoft and
  Wolfram Research are spending a lot of time and money developing
  material. We will provide technical and programming expertise to
  lecturers---helping them to develop the interactive part of notes
  while they provide the subject material.

We will work as consultants with researchers at Sheffield to
  introduce \TheProject to their workflow. Any projects that
  successfully do this will be promoted as case studies for
  \TheProject.

We have a collaboration with a lecturer from \href{https://en.wikipedia.org/wiki/Parthenope_University_of_Naples}{Partenope,
  University of Naples} who would like us to deliver our short course to lecturers there
  and to assist in the migration of some of their courses to \TheProject technologies.

We will use these experiences of migrating courses at Sheffield and Naples to \TheProject technologies
to develop a web template that allows the straightforward development of course websites from a
combination of GitHub and a folder-full of Jupyter notebooks.

The integration of Sun Grid Engine and Project Jupyter, done as part of \TheProject, has led to some educators considering using the notebook to introduce various aspects of High
Performance Computing. This includes GPU, OpenMP and MPI
  programming. \textbf{We have emerging collaborations with
	\href{https://notebooks.azure.com/}{Microsoft},
	\href{http://alces-flight.com/}{AlcesFlight}, and
	\href{http://gpucomputing.shef.ac.uk/}{Sheffield's GPU community}
	which will help further this aspect of \TheProject dissemination into the
	HPC community}.

The development of nbval and nbdime within \TheProject is
leading to some researchers switching to using the notebook for all
of their software documentation and tutorials. This is because
nbval makes it possible to ensure that such documentation will
always work as development of code progresses.
Using nbval allows the documentation to become part of
the formal testing framework.

Building on the work of nbval and nbdime,
we will deliver a workshop on reproducible science using \TheProject at
the UK national Research Software Engineering conference in 2017.

We will work with \href{http://rse.shef.ac.uk/blog/sheffield-code-firstgirls/}{Sheffield Code First: girls}, which introduces programming to women across the UK,
  to incorporate \TheProject into their curriculum.
This potentially has great impact for women in STEM subjects.

 By the end of the
  project we will have produced a series of interactive notebook
  demonstrators \delivref{dissem}{IntroODK} of \TheProject with a
  particular focus on computational biology, data science and machine
  learning. These notebooks will expand the use of VREs in these
  domains, appealing to researchers used to the domains of
  Bioconductor and MATLAB. We will make use of live notebook posters
  and commenting systems. These interactive notebooks
  will be provided in a public repository
  (\delivref{dissem}{IntroODK}).
\end{task}

\begin{task}[id=dissemination-of-oommf-nb-virtual-environment,
  title=Open source dissemination of micromagnetic VRE,
  lead=XFEL,PM=4,partners={SR,LEEDS,PS},wphases=28-32,issue=30]
  % 3 months person time + 1 months investigator time
  Tasks \taskref{UI}{oommf-py-ipython-attributes} and
  \taskref{UI}{oommf-tutorial-and-documentation} provide the
  micromagnetic VRE demonstrator
  (\ref{sec:introduction-micromagnetic-vre-demonstrator}) built on top
  of \TheProject.  In this task, we set up of the infrastructure
  to encourage
  and invite code contributions from the micromagnetic community to
  both code and created notebooks, while automating quality control
  and maintaining trust effectively (\delivref{dissem}{oommfnb-vre-deliver}).

  The source code of the micromagnetic VRE will be made available as
  open source on public repository hosting sites (such as
  GitHub/Bitbucket), and announced to the community via appropriate
  mailing lists and other means. We will set up a publicly accessible
  Jenkins/Travis continuous integration (CI) system to (i) run
  regression tests (from
  \taskref{component-architecture}{oommf-python-interface} and
  \taskref{UI}{oommf-py-ipython-attributes}) routinely when the
  micromagnetic VRE code or underlying OOMMF core code changes, (ii)
  re-execute notebooks (from
  \taskref{UI}{oommf-tutorial-and-documentation}) and use them as
  regression tests (using the outcome of task
  \taskref{UI}{notebook-verification}), and (iii) re-build
  downloadable installation files and virtual machine images.
  %This set
  %up will test user-contributions automatically.

  %versions (\delivref{dissem}{oommfnb-source-and-testing-setup}).

\end{task}

\begin{task}[title=Micromagnetic VRE dissemination workshops,
id=dissemination-of-oommf-nb-workshops,lead=XFEL,PM=6,partners={USO},wphases=18-44!0.46,issue=31] % funny ratio here, that's okay (HF, MB)

  % 3 months person time, 2 months investigator time

  We will run a series of workshops
  (\delivref{dissem}{oommfnb-vre-deliver}) during the evenings of 4
  major international meetings on magnetism
  research\footnote{Anticipated most significant international
	meetings in the appropriate time frame are 61st Conference on
	Magnetism and Magnetic Materials (MMM2016), October 31-November 4,
	2016, New Orleans, Louisiana; 62nd Conference on Magnetism and
	Magnetic Materials (MMM 2017), November 6-10, 2017, Pittsburgh,
	Pennsylvania; 21st International Conference on Magnetism (ICM
	2018), July 16–20, 2018, San Francisco, California; 14th Joint
	MMM-Intermag Conference (MMM2019), January 14-18, 2019,
	Washington, DC). Each of those meetings is one week long, and
	serves as a focal point of networking for the european and
	international research community. Other training events have been
	held in the past at these conferences and were well attended.} to
  disseminate the micromagnetic virtual research environment
  (Sect.~\ref{sec:introduction-micromagnetic-vre-demonstrator} and
  \localtaskref{dissemination-of-oommf-nb-virtual-environment}) in the
  micromagnetic community. Each conference attracts around 1500
  participants, and we expect at least 30 for our workshops at every
  event. Depending on demand, multiple workshops will be given per
  conference.

  The taught material will include (i) use of the \Jupyter-based
  micromagnetic VRE, and an (ii) introduction to the standard
  techniques for contributing to open source software (version
  control, pull requests, testing frameworks) to foster excellence in
  computational science and to make the micromagnetic VRE project
  self-sustaining as quickly as possible. In addition, all teaching
  materials, including videos, will be made available on a website.

  For each workshop, we request \euro{500} room hire at the magnetism
  conference location and the travel expenses for two teachers from
  Southampton to attend the one week international conference,
  totalling (\euro{500} + 2x\euro{2200}=\euro{4900}) per
  workshop. There are no other costs.

  We will use the micromagnetic VRE demonstrator
  (\taskref{UI}{oommf-tutorial-and-documentation}), its dissemination
  workshops (\taskref{dissem}{dissemination-of-oommf-nb-workshops})
  and interactions with its users and contributors in the
  micromagnetic community to evaluate, reflect and report on the
  project, taking into account technical and social aspects. A survey
  will be developed and used to gather user input and feedback on
  usefulness of the provided capabilities, with particular focus on
  the capabilities of the micromagnetic VRE to (i) enable new and
  better science, to (ii) allow to make progress effectively, to (iii)
  carry out computational science reproducibly, to (iv)
  collaboratively enable trust and to (v) become a self-sustained
  project from community contributions. Amongst other channels, we
  will target attendees of the micromagnetic VRE dissemination
  workshops (\taskref{dissem}{dissemination-of-oommf-nb-workshops}) to
  gather data.

  All results and insights will be summarised in a public document and
  reported at appropriate workshops and conferences to share the
  lessons learned from this \Jupyter-based VRE for micromagnetics. We
  will create a manuscript for journal publication, summarising the
  demonstrator project and this evaluation. An important point of this
  publication is to provide a reference that can be cited by
  publications making use of the new micromagnetic VRE, to allow
  tracking of uptake and development of this VRE beyond the life time
  of this H2020 project.
\end{task}

\begin{task}[title=Demonstrator: Interactive books,
id=ibook,lead=US,partners={USO,XFEL},PM=42,wphases={0-46,40-46},issue=32]
  % 2x12 _ 3x 3 months for students
  % 6 months Southampton, ibook4: maths for engineering, in months 40-46
One of the important elements of VREs is a common flexible writing format which
enables the creation of large structured documents. There are many
known solutions to that problem, but they usually compromise the
interactivity of the notebook interface and typesetting quality of desktop
publishing software like LaTeX.

Recently, a few approaches tried to bring both interactivity and the
typographic features. The modestly tagged markup language
\href{http://hplgit.github.io/doconce/doc/web/}{DocOnce}
targets the problem of reusability of the document source code for
producing traditional LaTeX-based printed books, IPython notebooks, Sphinx
documents (with Sage cells), and many other formats. MathBook XML
is a lightweight XML application for authors of scientific articles,
textbooks and monographs extensively using Sage cells for
interactive elements. The Sphinx documentation software has been
successfully applied for creation of interactive books containing Sage
cells. Additional interactivity is offered using the \href{http://runestoneinteractive.org}{Runestone tools}.

The technical aspects of format for interactive publications is a
subject of the task ``Structured documents'' in
\taskref{UI}{structdocs}. In this task we will demonstrate usability
of the results of \taskref{UI}{structdocs} in creation of scientific
textbooks. Three interactive books will be created:

\begin{compactitem}
\item Nonlinear Processes in Biology (\delivref{dissem}{ibook1})
\item Linear Algebra (\delivref{dissem}{ibook1})
\item Computational Mathematics for Engineering  (\delivref{dissem}{ibook3c})  % removed ibook4 to reduce number of deliverables as requested
																			   % point to ibook 3c instead.
\item Problems in Physics with Sage/Python (\delivref{dissem}{ibook3c})
\end{compactitem}

The choice of those particular topics has been made for the sake of
maximal diversity. The ``Nonlinear Processes in Biology'' will heavily
use numerical solution of ODEs and PDEs, the Linear Algebra book will
be a classical mathematical textbook, while the next book is
targetting the engineering community.  The last example will focus
mostly on collaborative editing and modularity of content which is
produced using VRE technologies. We will demonstrate the power of
symbolic algebra where appropriate throughout.  All books are natively
designed within the VRE, pushing the next generation of learning
methodology out into multiple communities.

The main research aspect for this task will be to integrate modern
computational tools in classical scientific topics and explore how
the VRE environment can accelerate the development and produce electronic
documents with significantly enhanced pedagogy and learning effectiveness.

In particular we will answer following questions:
\begin{compactitem}
\item When is a fully interactive worksheet required and when is
  a textbook with executable code cells sufficient?
\item How to assemble a classical monograph by reusing independently working
  building block of text and code?
\item What are best tools and practices for using a single source for
  producing printed and electronic (interactive) textbooks?
\item How to collaboratively write reusable course material?
\item How can we facilitate automatic testing of all code examples, plots, etc?
\item How can students can benefit from using VRE?
\end{compactitem}
\end{task}

\begin{task}[title=Demonstrator: Computational mathematics resources indexing service,
id=index-librorum-salvificorum,lead=UV,PM=2,partners={UB},wphases=20-25,issue=33] Beyond official documentation and
  tutorials, users of mathematical software and VREs learn from a wide
  array of sources: university courses, Q\&\ A sites, web searches,
  etc.  A simple web search on any major software component yields
  dozens of non-official tutorials and how-tos in many different
  languages. However, search engines mostly miss the relevant
  metadata: how does one find a tutorial on linear algebra in \PariGP,
  written at an undegraduate level, in French or Spanish?

This need has been felt by most communities at some point, and each
has come up with its own solution: most software components (e.g.,
\GAP, \PariGP, \Sage, \dots) simply link material from their official
page; \Sage has a wiki (\url{http://wiki.sagemath.org/}) referencing
additional resources, and used to host a large number of tutorial
worksheets on \url{http://sagenb.org/}; the recent introduction of
public projects in \SMC is sparking approximately the same phenomenon
that had previously happened with \url{http://sagenb.org/}; \IPython
host the Notebook Viewer service (\url{http://nbviewer.ipython.org/}),
which renders (without hosting) community-made notebooks; and teaching
institutions host or link their own collections of pedagogical
resources.

These collections are usually incomplete, limited in scope, hard to
search, and difficult to keep up-to-date.  What the community needs is a
community-curated, searchable, metadata-driven, multilingual, platform
agnostic indexing service whose goal is to reference and rank all the
community generated knowledge around a software component or VRE.

The goal of this task is to create the tool powering such service, and
to host a (free) community-curated index for \TheProject related
resources as a demonstrator (\delivref{dissem}{ils-tool}).

\end{task}




\end{tasklist}



%Raw material:
%\begin{compactitem}
%\item Documentation improvements: overview, cross links, overview of
%  recent improvements
%\item Thematic tutorials
%\item Collections of pedagogical documents\\
%  E.g. a complete collection of interactive class notes with computer
%  lab projects for the ``Algèbre et Calcul formel'' option of the
%  French math aggregation (starting from 2014-2015, only open-source
%  systems will be supported, and Sage is a major player).
%  % See http://nicolas.thiery.name/Enseignement/Agregation/ as a starter
%  % Math labs with Sage for first year students in France (L1): http://math.univ-lyon1.fr/~omarguin/
%\item Localization of the Sage user interface and key documents in
%  various European languages.
%\item Distribution of the documents either in the main distribution of
%  Sage or through the online repository (see collaborative tools).
%\item Massive online introduction course to Sage, drawing on the sage tutorial/notebooks.
%Could be ``First year Sage course in a box''.
%\item Taking the opportunity of Python courses to propose Sage as a natural extension
%for mathematics; an example is French's
%% The url macro eats the accented letters.
%% It doesn't just eat it, it pukes it back!
%``Classes pr\'eparatoires''
%%\footnote{\url{http://en.wikipedia.org/wiki/Classe_préparatoire_aux_grandes_écoles}},
%where Python has been recently selected as the language to learn programming\footnote{See
%the ``Annexe'' at
%\url{http://www.education.gouv.fr/pid25535/bulletin_officiel.html?cid_bo=71586}}.
%%\item \TODO{please expand!}
%\end{compactitem}

% Jeroen: About teaching: in Gent, Sage is already integrated in the
% courses (maybe you can add this, don't know if it's relevant)
% starting in the first year. It's good for the students because it
% helps in 2 ways: it helps them to understand the mathematics better
% and it helps them to learn basic down-to-earth programming (they
% also have a programming course in Java but that contains a lot of
% theory about complicated class structures)
% Same thing in Orsay
% More python centered but same in UZH
% We have also Sage @ Silesia from 1st semester (physics)

\begin{wpdelivs}
  \begin{wpdeliv}[due=6,miles=startup,id=press-release-1,dissem=PU,nature=DEC,lead=PS,issue=34, status=delivered]{Starting press release}\end{wpdeliv}
  \begin{wpdeliv}[due=12,miles=startup,id=workshops-1,dissem=PU,nature=R,lead=PS,issue=42, status=delivered]{Community building: Impact of development workshops, dissemination and training activities, year 1}\end{wpdeliv}
%  \begin{wpdeliv}[due=12,miles=startup,id=ibook3a,dissem=PU,nature=DEM,lead=US]{Demonstrator: Problems in Physics with Sage v1} \end{wpdeliv}
  \begin{wpdeliv}[due=12,miles=startup,id=techno,dissem=PU,nature=R,lead=PS,issue=43, status=delivered]{Review on emerging technologies} \end{wpdeliv}
 \begin{wpdeliv}[due=18,miles=proto1,id=short-course,dissem=PU,nature=DEC,lead=USH,issue=44, status=delivered]{A short course for lecturers on using \TheProject for delivering mathematical education.}\end{wpdeliv}
  \begin{wpdeliv}[due=24,miles=proto1,id=datascience-course,dissem=PU,nature=DEC,lead=USH,issue=45,status=canceled]{Course material on using \TheProject in data science and education}
  This deliverable was merged into \localdelivref{IntroODK}.
  \end{wpdeliv}
  \begin{wpdeliv}[due=24,miles=proto1,id=workshops-2,dissem=PU,nature=R,lead=PS,issue=46,status=canceled]
	{Community building: Impact of development workshops, dissemination and training activities, year 2}
	This deliverable was merged into~\localdelivref{workshops-3}.
  \end{wpdeliv}
  \begin{wpdeliv}[due=24,miles=proto1,id=ils-tool,dissem=PU,nature=DEM,lead=UV,issue=47,status=delivered]{Community-curated indexing tool (open source)} \end{wpdeliv}
% merged with ils-tool  \begin{wpdeliv}[due=24,miles=proto1,id=ils-service,dissem=PU,nature=DEM,lead=UV]{Community-curated indexing service for \TheProject} \end{wpdeliv}
  \begin{wpdeliv}[due=24,miles=UI-vre-prototype,id=ibook2,dissem=PU,nature=DEM,lead=US,issue=48,status=canceled]
	{Demonstrator: Linear Algebra - interactive book}
	This deliverable was merged into~\localdelivref{ibook1}.
  \end{wpdeliv}
% merged with workshop reports  \begin{wpdeliv}[due=24,miles=proto1,id=dissem-1,dissem=PU,nature=R,lead=PS]{Impact of dissemination and training activities, years 1 and 2}\end{wpdeliv}
%  \begin{wpdeliv}[due=30,miles=community,id=ibook3b,dissem=PU,nature=DEM,lead=US]{Demonstrator: Problems in Physics with Sage v2} \end{wpdeliv}
  \begin{wpdeliv}[due=36,id=ibook1,miles=UI-vre-prototype,dissem=PU,nature=DEM,lead=US,issue=49,status=delivered]
	{Demonstrator: interactive books on Linear Algebra and Nonlinear Processes in Biology}
  \end{wpdeliv}
  \begin{wpdeliv}[due=36,id=lecture-notes,miles=UI-vre,dissem=PU,nature=DEM,lead=USH,issue=35,status=canceled]
	{Demonstrator: Interactive lecture notes and marking systems based on \TheProject}
	This deliverable was merged into \localdelivref{IntroODK}.
  \end{wpdeliv}
  \begin{wpdeliv}[due=36,id=workshops-3,dissem=PU,miles=eval,nature=R,lead=PS,issue=36,status=delivered]
	{Community building: Impact of development workshops, dissemination and training activities, year 2 and 3}
  \end{wpdeliv}
  \begin{wpdeliv}[due=44,miles=eval,id=notebook-repo,dissem=PU,nature=DEM,lead=USH,issue=37,status=canceled]
	{Demonstrator: Repository of interactive Notebooks in Machine Learning and Computational Biology based on \TheProject.}
	This deliverable was merged into \localdelivref{IntroODK}.
  \end{wpdeliv}
  \begin{wpdeliv}[due=48,miles=UI-vre,id=oommfnb-vre-deliver,dissem=PU,nature=OTHER,lead=XFEL,issue=38]{Micromagnetic VRE completed and online} \end{wpdeliv}
%  \begin{wpdeliv}[due=47,miles=eval,id=ibook3c,dissem=PU,nature=DEM,lead=US]{Demonstrator: Problems in Physics with Sage v3} \end{wpdeliv}
%  \begin{wpdeliv}[due=47,miles=eval,id=ibook4,dissem=PU,nature=DEM,lead=USO]{Demonstrator: Computational Mathematics for Engineering} \end{wpdeliv}
% Combine the above demonstrators into one Deliverable
  \begin{wpdeliv}[due=47,miles=eval,id=ibook3c,dissem=PU,nature=DEM,lead=US,issue=39]{Demonstrators: Problems in Physics with Sage, Computational Mathematics for Engineering}
  \end{wpdeliv}

  \begin{wpdeliv}[due=48,miles=eval,id=workshops-4,dissem=PU,nature=R,lead=PS,issue=40,status=delivered]{Community building: Impact of development workshops, dissemination and training activities, year 4}\end{wpdeliv}
% merged with workshop reports  \begin{wpdeliv}[due=48,id=dissem-2,dissem=PU,nature=R,lead=PS]{Impact of dissemination and training activities, years 3 and 4}\end{wpdeliv}
  \begin{wpdeliv}[due=48,miles=eval,id=press-release-2,dissem=PU,nature=DEC,lead=PS,issue=41]{Ending press release}\end{wpdeliv}
% The deliverable below is a result of merging datascience-course, lecture-notes and notebook-repo
\begin{wpdeliv}[due=48,miles=UI-vre,id=IntroODK,dissem=PU,nature=DEC,lead=LEEDS,issue=252]{Introduce OpenDreamKit to Researchers and Teachers as laid out in Task 2.6}\end{wpdeliv}
\end{wpdelivs}
\end{workpackage}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  workpackage dissem wphases wpobjectives wpdescription tasklist WPref nmag
%  LocalWords:  delivref linkedin organisation finalpressrelease organise wpdelivs github
%  LocalWords:  wpdeliv dissemination-of-oommf-nb-virtual-environment OOMMFNB taskref
%  LocalWords:  oommf-python-interface oommf-tutorial-and-documentation mumag magpar
%  LocalWords:  mumax micromagnum micromagnetic oommf-py-ipython-attributes summarising
%  LocalWords:  dissemination-of-oommf-nb-workshops localtaskref MMM-Intermag Fangohr
%  LocalWords:  maximise sagecell structdocs Algèbre Calcul formel eparatoires Annexe
%  LocalWords:  Jeroen
