\TOWRITE{UM}{Proofread WP 7 Social aspects pass 2}
\begin{draft}
\begin{verbatim}
- [X] have all the tasks in this Work Package a lead institution?
- [X] have all deliverables in the WP a lead institution?
- [X] do all tasks list all sites involved in them?
- [X] does the table of sites and their PM efforts match lists of sites for each task?
      (each site from the table is listed in all relevant tasks, and no site is listed
      only in the table or only at some task)
\end{verbatim}
\end{draft}


\begin{workpackage}[id=social-aspects,wphases=0-48,
  title=Social Aspects (canceled),
  lead=UO,
  UORM=6,status=canceled] % deleted Sheffield participation of 24PM and deleted 17 PM from Oxford, XFEL deleted its 6 PM

\color{gray}
\begin{wpobjectives}
\color{gray}
The processes by which mathematical knowledge and mathematical
software are developed, validated and applied are quite
distinctive. In other sciences, the universe provides ``ground truth''
and the scientific texts or theories can be validated against that by
experiment. In mathematics the text itself is the ground truth. The
traditional model of mathematical research is a mathematician, or a
small group of mathematicians, standing around a blackboard, producing
a proof they would ``clean up'': remove all traces of the process
that led to its discovery and then submit the ``clean'' text to
their peers for review.

Mathematicians have adopted new technology in a variety of ways: email
and shared documents are used to collaborate on problem-solving and
writing; larger ``crowdsourcing'' \cite{polymath_SIAM, PolymathBlog},
arrangements pull together diverse experts; symbolic computation
tackles huge routine calculations; and computers check proofs that are
too long and complicated for a human to comprehend. These
technologies reveal (since email messages, version control
systems and bulletin boards can be analysed) and alter the ways in
which mathematicians collaborate.

In an EPSRC funded project ``The Social Machine of Mathematics''
Martin and others are bringing together rigorous methods from the social
sciences to study these collaborative processes. Combining this
research with the algorithmic game theory expertise of Elkind and Pasechnik,
in this work package we intend to pursue the following objectives:

\begin{compactitem}
\item incorporate the insights from this and similar projects into the
  design of \TheProject VRE, ensuring that it supports the ways in
  which mathematicians really work, rather than the way software
  developers---or indeed mathematicians---think they do;
\item extend this work to study the collaborative processes of free
  open source (mathematical) software development so as to produce
  guidelines for best practice as well as to develop ideas for extending existing processes
  to a ``system of systems''.
\end{compactitem}
\end{wpobjectives}

\begin{wpdescription}
\color{gray}
``Crowdsourcing''---fine-grained collaborative development of ideas,
  proofs or software---is a common theme to both objectives. The purpose
  of a VRE is to allow effective crowdsourcing of computationally
  supported mathematical results (theorems, proofs, etc.),
  while free software development is inherently a collaborative process,
  and we wish to study the best ways of allowing it to scale.

  In a sense, mathematics has been a crowdsourced endeavour, dating as
  far back as the foundation of the Royal Society (UK) in the seventeenth
  century.  The first scientific journals were published collections
  of letters received, posing questions and observations and offering
  solutions.  Although limited by the speed of physical post, this
  model had much in common with the public email lists that
  underpinned collaborative software development in the 1990s.

  In recent years, the internet and critical tools such as distributed
  version control have supported much more widespread and finer-grained
  forms of crowdsourcing, first in software development, and, more recently,
  in mathematics: examples are provided by online mathematics communities, such as Math-overflow
  \cite{MathOverflow} and Polymath Projects \cite{polymath_SIAM,
  PolymathBlog}.  Supporting and encouraging ``Mutual
  crowdsourcing'' is the main driving force for developing and
  maintaining any large-scale open-source virtual research
  environment.

  %deletion:In this work package we will build on the work of Prof. Martin and her collaborators
  %on the EPSRC project, and, in particular, their study of crowdsourcing, and
  %integrate their findings with tools provided by the burgeoning field
  %of algorithmic mechanism design in order to to
  %optimise crowdsourcing workflows in open-source VREs.
  This work package was initially supposed to build on the work of Prof. Martin and her collaborators on the EPSRC
  project. However due to her retirement and to change of PI
  at USFD it became difficult for the project to accomplish objectives.
  As a result tasks T7.1, T7.2 and T7.3 were cancelled. The task T7.4 was merged with T2.8.

\end{wpdescription}

\begin{tasklist}
\begin{task}[title=Social Science Input to
    Design,id=social-input,lead=UO,PM=18, partners={UO,PS},wphases={0-3!.5,12-42!.5},issue=144,
    status=canceled]
%The purpose of this task is to ensure that the design of \TheProject
%VRE reflects the lessons learned by social scientists studying the
%ways in which mathematicians actually collaborate and work. Since UO and
%Martin in particular are already central in the community
%working in this area, we are well placed to ensure that this happens.

%As soon as the project begins, team members at UO will combine their
%own work with a review of the published literature, and identify and
%meet with key research groups in this area, in order to distill relevant
%current knowledge for use in the design phases of other parts of the
%project. They will  present the lessons
%learned at project meetings and workshops and deliver it as a report
%\localdelivref{social-report-three} (part I) in month 3.

%After that, they will monitor the further development of this area and
%ensure that any new insights are communicated promptly to the rest of
%the project. This will be synthesised for archival purposes into two
%further reports
%\localdelivref{social-report-three} in months 24 (part II) and 42 (part III).

%We will survey the data needed to assess development models of
%large-scale academic open-source projects, such as the probable
%correlation between the size of the atomic contribution vs. the speed
%of the contribution making it into the code, and collect appropriate
%statistical data, to be published as a report (and possibly a conference
%publication) \localdelivref{social-datareport}.
%The latter will require non-trivial amount of
%programming work, even only for the test system, \Sage.
This task was cancelled. Only D7.1 could be delivered.
\end{task}

\begin{task}[title=Implications of VREs for Publication,id=social-output,lead=USH,PM=12,wphases=12-42,partners={UO},issue=145,status=canceled]
  %A key aspect of the \TheProject VRE is support for the full
  %life-cycle of mathematical research, up to, and after publication of
  %results. While it is necessary to support established models of
  %publication, which are central to mathematical practice and academic
  %life, it is also appropriate to explore whether VRE technologies may enable
  %novel models for the distribution of scientific output
  %that are more effective for new forms of mathematical results.

  %The current model for dissemination of scientific output stems from
  %an era when the printing press was dominant. The process has become
  %formalised through peer review and publication of journals. The PDF
  %format widely used for distribution of documents reflects the
  %\textit{status quo} that a scientific paper is written as if for
  %printing and remains an unchanging document. In scientific blogging
  %we are seeing that more rapid propagation of ideas can occur when
  %the constraints of the printed format are relaxed; however, these
  %dissemination routes lack the formalisation that ensures (usually) fair
  %attribution of ideas and commentary.

  %We will prototype and evaluate tools and ideas for dissemination of
  %scientific knowledge that do not rely on a static format and allow
  %for the full spectrum of scientific debate.  The tools will
  %enable proper credit allocation by encouraging shared
  %attribution of ideas, software and data. This will interact with
  %work in \WPref{dksbases} concerning attribution and citeability for
  %mathematical databases.

  %Tools to be prototyped include live posters for distribution of
  %knowledge, designed for integration with either large touch screens
  %or smaller tablets \localdelivref{social-poster} as well as extensions
  %of the Jupyter project that would provide facilities for commenting on
  %notebooks, which we expect to encourage debate on mathematical and computational
  %ideas \localdelivref{jupyter-comment}.
  This task was cancelled.
\end{task}


\begin{task}[title=Mechanism Design for Free Software Development,PM=15,lead=UO,
  wphases=6-48!.5,id=isocial-decisionmaking,issue=146,status=canceled]
%While crowdsourced open-source software development has become an
%incredibly powerful force in recent years, it still has limitations.
%Open source projects tend to be fragile, in the community sense, and
%suffer from disagreements that ultimately result in ``forks'' and the
%resulting duplication of effort. We will analyse this phenomenon in the framework
%of algorithmic game theory, and try to design finely tuned systems of
%%ncentives and rewards for contribution so as to increase the stability of
%the community and its useful output.

%We will focus on three areas:
%(1) prioritisation of bug fixes and feature requests to achieve reliable and useful systems;
%(2) effective cooperation among multiple collaborative projects; and
%(3) making decisions about the strategic direction of the system.

%We will use prioritisation as a testbed for designing incentives that encourage all
%participants to contribute towards sustained development
%of the most important parts of the system.
%To this end, we will use
%ideas from the burgeoning field of mechanism design \cite{AGTbook} and
%in particular recent research on crowdsourcing in algorithmic
%mechanism design \cite{crowds}.  While doing so, we will apply
%outcomes to a case study system---\Sage.

%The reason why prioritisation poses a challenge in the
%development of open-source academic software is that this process is task-driven:
%typically, tasks (also known as tickets) are posted on a website, and their
%priorities are set in an ad hoc manner.  This model is usually
%good enough for simple bug fixing, but for more elaborate tasks it often
%leads to unacceptable delays. We will apply preference
%and opinion aggregation techniques \cite{pref-aggr} to develop a
%community prioritisation scheme for bugs and feature requests (which may rely on a reputation scores
%technique, such as one used on MathOverflow),
%and implement this scheme as a TRAC \cite{Trac} add-on
%\localdelivref{social-tracaddon}.
%As \Sage is using the TRAC server \cite{trac-sagemath},
%this will be easy to test on our testbed system.

%Trusting results of computer calculations is crucial for
%usability; channels for communicating bug reports and fixes need to be
%carefully analysed from social point of view.  Commercial
%closed-source computer algebra and other computational systems often fail to
%react to bug reports in a timely manner, and sometimes fall
%into the short-sighted trap of hiding bugs from potential and current
%users \cite{misfort}. Open source systems are only marginally better
%in this respect, as indicated by recent computer security scares, such as the one
%around Bash \cite{shellshock}.  A game-theoretic analysis of
%this situation will be attempted.

%A key strength of free and open-source software models is the ability
%to build upon pre-existing software. \GAP, \PariGP, \Singular and
%especially \Sage have made heavy use of this ability. Problems arise over
%time, however, as priorities of the system developers
%diverge. For instance, bugs reported by so-called ``downstream'' systems may not be
%given the same priority as bugs reported by direct
%users of the ``upstream'' system, or ignored altogether;
%similarly, incompatible changes can be made as long as they are acceptable
%to the direct user community, even if they cause problems
%for a dependent system. We will explore how sociological and game-theoretic
%insights can be used to reduce these problems.

%The results of this task will be summarised in \localdelivref{social-gametheoretic}
%and reported at relevant AI workshops and conferences.
This task was cancelled.
\end{task}

\begin{task}[title=Evaluation of Micromagnetic VRE,lead=XFEL,PM=6,
  id=oommf-nb-evaluation,partners={UO,PS},wphases=32-44!0,issue=147,status=canceled]
  % 4 person months, 1 person month investigator time
  This task has been moved into \taskref{dissem}{dissemination-of-oommf-nb-workshops}.
\end{task}



\end{tasklist}

% Things to investigate?
% - User surveys. Cf. https://groups.google.com/d/msg/sage-devel/v8Kfky4p6D4/_xRM0bggCo8J
% - The discussion about Code of Conducts and the like

\begin{wpdelivs}
\begin{wpdeliv}[due=18,miles=proto1,id=social-datareport,dissem=PU,nature=R,lead=UO,issue=148, status=delivered]
{The flow of code and patches in open source projects}
\end{wpdeliv}

\begin{wpdeliv}[due=24,miles=proto1,id=social-tracaddon,dissem=PU,nature=OTHER,lead=UO,issue=149,
status=canceled]
{TRAC add-on to manage ticket prioritisation}
\end{wpdeliv}

\begin{wpdeliv}[due=24,miles=proto1,id=jupyter-comment,dissem=PU,nature=DEM,lead=USH,issue=150,status=canceled]
   {Demo: Mechanism for comments on posted Jupyter notebooks}
\end{wpdeliv}

 \begin{wpdeliv}[due=36,miles=eval,id=social-poster,dissem=PU,nature=DEM,lead=USH,issue=151,status=canceled]
   {Demo: Jupyter Notebook Live Poster}
\end{wpdeliv}

\begin{wpdeliv}[due=42,miles=eval,id=social-report-three,dissem=PU,nature=R,lead=UO,issue=152,
status=canceled]
 {Report on relevant research in sociology of mathematics and lessons
   for design of \TheProject VRE, parts I-III, with part I (resp. II) due at month 3 (resp. 24)}
\end{wpdeliv}
\begin{wpdeliv}[due=42,miles=eval,id=social-publishing-report,dissem=PU,nature=R,lead=USH,issue=153,status=canceled]
{Review of new publication mechanisms, including evaluation of
  demonstrator projects}
\end{wpdeliv}

\begin{wpdeliv}[due=42,miles=eval,id=social-gametheoretic,dissem=PU,nature=R,lead=UO,issue=154,
status=canceled]
{Game-theoretic analysis of development practices in open-source VREs}
\end{wpdeliv}

\begin{wpdeliv}[due=48,id=oommf-nb-evaluation,dissem=PU,nature=R,lead=XFEL,issue=155,status=canceled]
     {Micromagnetic VRE environment evaluation report}
     This deliverable was merged into \delivref{dissem}{oommfnb-vre-deliver}.
\end{wpdeliv}
\end{wpdelivs}
\end{workpackage}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

%  LocalWords:  workpackage wphases TOWRITE wpobjectives analyse wpdescription AGTbook
%  LocalWords:  mathoverflow Sagemath pref-aggr prioritisation trac-sagemath Trac misfort
%  LocalWords:  analysed shellshock tasklist datacollection decisionmaking incentivised
%  LocalWords:  OOMMFNB taskref oommf-python-interface oommf-tutorial-and-documentation
%  LocalWords:  micromagnetic dissem dissemination-of-oommf-nb-virtual-environment texttt
%  LocalWords:  dissemination-of-oommf-nb-workshops summarised delivref wpdelivs wpdeliv
%  LocalWords:  oommf-nb-evaluation compactitem recomputation-style phenomenom endeavour
%  LocalWords:  localdelivref synthesised social-datareport textit WPref dksbases Elkind
%  LocalWords:  citeability isocial-decisionmaking social-tracaddon social-gametheoretic
%  LocalWords:  linebreak micromagnetics summarising Pasechnik
