\section{Conclusion}\label{sec:concl}
In this paper we have presented the \ODK project and the ``Math-in-the-Middle'' approach
it explores for mitigating the system integration problems inherent in combining an ecosystem
of open source software systems into a coherent mathematical virtual research environment.
The MitM approach relies on a central, curated, flexiformal ontology of the mathematical
domains to be covered by the VRE together with system-near interface theories and
inter-views to the core ontology that liaise with the respective systems. We have reported
on two case studies that were used to evaluate the approach: an interface for the \LMFDB,
and a more semantic handle interface between \GAP and \Sage.

Even though the development of the MitM is still at a formative stage, these case studies
show the potential of the approach. We hope that the nontrivial cost of curating an
ontology of mathematical knowledge and inter-views to the interface theories will be offset
by its utility as a resource, which we are currently exploring; the unification of the
knowledge representation components
\begin{compactenum}
\item enables VRE-wide domain-centered (rather than system-centered) documentation;
\item can be leveraged for distributed computation via uniform protocols like the
  SCSCP~\cite{HorRoz:ossp09} and MONET-style service
  matching~\cite{CaprottiEtAl:MathServiceMatching04:tr} (the absence of content
  dictionaries -- MitM theories -- was the main hurdle that kept these from gaining more
  traction);
\item will lead to the wider adoption of best practices in
  mathematical knowledge management in the systems involved; in
  fact, this is already happening.
\end{compactenum}
Whether in the end the investment into the MitM will pay off also depends on the quality
and usability of the tools for mathematical knowledge management. Therefore we invite the
CICM community to interact with and contribute to the \ODK project, on this work package
and the others. Possible contributions include
\begin{compactenum}
\item interfacing another system to the MitM architecture via interface theories
\item contributing to the MitM core ontology 
\item MitM-refactoring existing integrations of mathematical software systems.
\end{compactenum}

\subsubsection*{Acknowledgements}

The authors gratefully acknowledge the other participants of the St~Andrews workshop, in
particular John Cremona, Luca de Feo, Mihnea Iancu, Steve Linton, Dennis M\"uller, Viviane
Pons, Florian Rabe, and Tom Wiesing, for discussions and experimentation which clarified
the ideas behind the math-in-the-middle approach.

We acknowledge financial support from the OpenDreamKit Horizon 2020 European Research
Infrastructures project (\#676541), from the EPSRC Collaborative Computational Project
CoDiMa (EP/M022641/1) and from the Swiss National Science Foundation grant PP00P2\_138906.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

%  LocalWords:  ednote subsubsection Dehaye Mihnea Iancu Konovalov Leli evre Wiesing
%  LocalWords:  concl flexiformal compactitem ossp09 compactenum domain-centered
%  LocalWords:  MitM-refactoring
