\section{Introduction}

\ednote{NT: wondering if the next three paragraph would not fit better
  in the odk section; the reader more interested by the MitM part may
  be put off by this long discussion}

From their earliest days, computers have been used in pure mathematics, either to make
tables, to prove theorems (famously the four colour theorem) or, as with the astronomer's
telescope, to explore new theories. Computer-aided experiments, and the use of databases
relying on computer calculations such as the Small Groups Library in GAP, the Modular
Atlas in group and representation theory, or the $L$-functions and Modular Forms Database (\LMFDB, see later), are part of the standard
toolbox of the pure mathematician, and certain areas of mathematics completely depend on
it. Computers are also increasingly used to support collaborative work and education.

The last decades witnessed the emergence of a wide ecosystem of open-source tools to
support research in pure mathematics. This ranges from specialized to general purpose
computational tools such as \GAP, \PariGP, \Linbox, \MPIR, \Sage, or \Singular, via online
databases like the \LMFDB and does not count online services like Wikipedia,
\Arxiv, or MathOverflow. A great opportunity is the rapid emergence of key technologies,
and in particular the \Jupyter (previously \IPython) platform for interactive and
exploratory computing which targets all areas of science.
\ednote{R3: "does not count online services like Wikipedia" - why not, specifically?}

This has proven the viability and power of collaborative open-source development models,
by users and for users, even for delivering general purpose systems targeting a large
public (researchers, teachers, engineers, amateurs, \ldots). Yet some critical long term
investments, in particular on the technical side, are in order to boost the productivity
and lower the entry barrier:
\begin{compactitem}
\item Streamlining access, distribution, portability on a wide range of platforms, including
  High Performance Computers or cloud services.
\item Improving user interfaces, in particular in the promising area of collaborative
  workspaces as those provided by \SMC.
\item Lowering barriers between research communities and promoting dissemination. For example
  make it easy for a specialist of scientific computing to use tools from pure
  mathematics, and reciprocally.
\item Bringing together the developer communities to promote tighter collaboration and
  symbiosis, accelerate joint development, and share best practices.
\item Outsourcing as much of the development as possible to larger communities to focus
  the work forces on their core specialty: the implementation of mathematical algorithms
  and databases.
\ednote{R3: "Outsourcing as much of the development as possible" - is this a "critical 
long term investment"?  Seems more like a "goal".  How will it work?}
\item And last but not least: Promoting collaborations at all scales to further improve
  the productivity of researchers in pure mathematics and applications.
\end{compactitem}
These can be subsumed by the goal of \emph{Virtual Research Environments} (VRE), that is
online services enabling groups of researchers, typically widely dispersed, to work
collaboratively on a per project basis. This is exactly where the \ODK project kicks in. 
\ednote{R3: "These can be subsumed by the goal of Virtual Research Environments" - I'm not 
sure I totally believe this claim, vis a vis the long-term/large-scale outreach items.}

We will introduce the \ODK project Section~\ref{sec:odk} to establish the context for the
``Math-in-the-Middle'' (MitM) integration approach described in
Section~\ref{sec:mitm}. The remaining sections then elucidate the approach by presenting
first experiments and refinements of the chosen integration paradigm:
Section~\ref{sec:lmfdb} details how existing knowledge representation and data structures
can be represented as MitM interface theories with a case study of equipping the \LMFDB
with a MitM-based programming interface.  Section~\ref{sec:gapsage} discusses system
integration between \GAP and \Sage and how this can be routed through a MitM
ontology. Section~\ref{sec:concl} concludes the paper and discusses future work.
\ednote{R1: Unfortunately there is no related work section included in the description 
in order to place the proposal in context. Are there no existing computational math 
interface services or attempts for at least a couple of resources? Surely there are 
at least some primitive expression-based services for connecting systems via MathML 
or OpenMath. Furthermore, the basic idea of providing integrated services has been 
adopted in many fields, and the fundamental architecture based on "mediators" 
together with their associated "wrappers" or "adapters" per service is well-understood 
and should be referenced [Gio Wiederhold: Mediators in the Architecture of Future 
Information Systems. IEEE Computer 25(3): 38-49 (1992).].}
\ednote{R1: Around p. 2 (i.e. close to the end of Introduction), it would be very 
informative to include a worked-out example of the sort of facility that the system 
will eventually provide, e.g., to "make it easy for a specialist of scientific 
computing to use tools from pure mathematics, and reciprocally." This should not 
include how the conversions are accomplished, but rather serve as a feasible 
example of how the integrated service could be used by a researcher in mathematics.}
\ednote{R1: Throughout the paper is would be better to reduce the use of acronyms. 
For example, the paper is unnecessarily difficult to read because of the extensive 
use of "VRE" and "COMs" as well as the many other acronyms. AK: I suggest to use VRE,
but revise the others.}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

%  LocalWords:  specialized Arxiv Jupyter IPython ldots compactitem emph mitm lfmdb concl
%  LocalWords:  gaptypes
