The mathematical software systems to be integrated via the MitM approach have so far been computation-oriented, e.g., computer algebra systems.
Their API theories typically declare types and functions on these types (the latter including constants seen as nullary functions).
Even though database systems differ drastically from these in many respects, they are very similar at the MitM level: a mathematical database defines
\begin{compactitem}
 \item some types: each table's schema is essentially one type definition,
 \item many constants: each table entry is one constant of the corresponding type.
\end{compactitem}
Thus, we can reuse many of the same concepts.
In particular, the API theories must contain definitions of the database schemas. 

Apart from standard software engineering tasks, this leaves three conceptual problems we had to solve:
\begin{compactenum}[\bf P1]
\item Turn the database schemas and tables into \ommt theories and declarations. 
\item Lift data in \emph{physical} representation (as records of the
  underlying database) to \ommt object in \emph{semantic} representation.
\item Translate semantic queries to queries about physical representations so
  that they can be executed directly on the database without loading the entire theory into
  \mmt.
\end{compactenum}
We deal with \textbf{P1} in Section~\ref{sec:vt}, with \textbf{P2} in Section~\ref{sec:vt:codec}, and with \textbf{P3} in Section~\ref{sec:qmt}. 

In this paper, we focus on \lmfdb as an example, of which we give an overview in Section~\ref{sec:lmfdb}.
However, our methods are general enough to apply to many other mathematical databases such as OEIS, or findstat.

 \subsection{LMFDB Overview}\label{sec:lmfdb}
  \input{sota}
 \subsection[Virtual Theories]{\lmfdb as a Set of Virtual Theories}\label{sec:vt}
  \input{vt}
 \subsection{Ascribing Encodings in Schema Theories}\label{sec:vt:codec}
  \input{access}
 \subsection{Translating Queries}\label{sec:qmt}
  \input{qmt}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:

%  LocalWords:  nullary compactitem compactenum ommt emph textbf textbf sec:qmt lmfdb qmt
%  LocalWords:  sec:lmfdb
